---
title: "Section 5: Simulation Study — Rule B"
subtitle: "Composed Pipeline Operating Characteristics with Blinded Nuisance-Parameter SSR"
author: "Lu Qian · Zetyra | Evidence in the Wild"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-float: true
    code-fold: true
    code-tools: true
    theme:
      - cosmo
      - custom.scss
    fig-width: 9
    fig-height: 5.5
    df-print: kable
    embed-resources: true
  pdf:
    toc: true
    number-sections: true
    fig-width: 7
    fig-height: 4.5
    geometry: margin=1in
    fontsize: 11pt
execute:
  echo: true
  warning: false
  message: false
  cache: true
  cache-path: "_cache_rule_b/"
bibliography: references.bib
---

```{r setup}
#| cache: false
library(RBesT)
library(ggplot2)
library(dplyr)
library(tibble)
library(purrr)
library(tidyr)
library(knitr)
library(patchwork)
library(scales)

# ── Color palette ─────────────────────────────────────────────────────────────
# Ink/green retained as primary accent; backgrounds switched to clean white/slate
ZET <- list(
  ink      = "#1A1A2E",   # deep navy-black — titles, axis text
  green    = "#1B3A2E",   # forest green  — primary series, headers
  green2   = "#2D6A4F",   # mid green     — secondary series
  teal     = "#40916C",   # light green   — tertiary / timing sensitivity
  slate    = "#4B5563",   # dark slate    — axis text, subtitles, legends
  gridline = "#E5E7EB",   # cool light gray — major gridlines
  red      = "#B91C1C",   # accessible red — reference lines, FAIL
  amber    = "#D97706",   # amber         — warning / border
  bg       = "#FFFFFF",   # pure white    — plot background
  panel    = "#FFFFFF",   # white         — panel background
  muted    = "#545C6B",   # dark grey     — captions, reference annotations (6.7:1)
  neutralL = "#46505F",   # darker grey   — grey DATA series (8.0:1 on white)
  neutralF = "#8B95A3"    # mid grey fill — baseline/neutral BARS (3.0:1, WCAG 1.4.11)
)

# ── Cache invalidation on palette change ──────────────────────────────────────
# knitr keys a chunk's cache on that chunk's own code text. The palette lives
# here, in an uncached setup chunk, so editing a colour does NOT invalidate the
# cached plotting chunks — a palette-only edit would silently leave most figures
# at the OLD colours while the few uncached ones update, producing a deck with
# mismatched greys. Making the palette part of every chunk's cache key fixes
# that: change a colour and every figure regenerates; leave it alone and the
# expensive simulations stay cached.
# autodep keys each chunk on the objects it USES, not only on its own source
# text. Without it, editing simulate_one_trial() or an SSR rule leaves every
# expensive downstream chunk cached and silently stale -- the same class of
# failure the palette cache.extra was added to fix.
knitr::opts_chunk$set(cache.extra = list(zet_palette = ZET), autodep = TRUE)
knitr::dep_auto()

# ── Modern publication theme ──────────────────────────────────────────────────
theme_zetyra <- function(base_size = 12) {
  theme_minimal(base_size = base_size, base_family = "sans") +
    theme(
      # Backgrounds — clean white, no colored fills
      plot.background   = element_rect(fill = ZET$bg,    color = NA),
      panel.background  = element_rect(fill = ZET$panel, color = NA),

      # Grid — thin horizontal only; vertical removed for clarity
      panel.grid.major.x = element_blank(),
      panel.grid.major.y = element_line(color = ZET$gridline, linewidth = 0.45),
      panel.grid.minor   = element_blank(),

      # Axes — subtle bottom/left rule; no full box
      axis.line.x  = element_line(color = "#D1D5DB", linewidth = 0.5),
      axis.ticks.x = element_line(color = "#D1D5DB", linewidth = 0.4),
      axis.ticks.length = unit(3, "pt"),
      axis.text    = element_text(color = ZET$slate,  size = base_size - 1),
      axis.title   = element_text(color = ZET$ink,    size = base_size,
                                   face = "bold", margin = margin(t = 4, r = 4)),

      # Titles — left-aligned, strong hierarchy
      plot.title    = element_text(color = ZET$ink,   size = base_size + 3,
                                   face = "bold",     hjust = 0,
                                   margin = margin(b = 4)),
      plot.subtitle = element_text(color = ZET$slate, size = base_size - 1,
                                   hjust = 0,         margin = margin(b = 10)),
      plot.caption  = element_text(color = ZET$muted, size = base_size - 2,
                                   hjust = 0,         margin = margin(t = 8)),
      plot.margin   = margin(12, 16, 10, 12),

      # Legend — top-right, no box
      legend.position    = "top",
      legend.justification = "left",
      legend.background  = element_rect(fill = NA, color = NA),
      legend.key         = element_rect(fill = NA, color = NA),
      legend.text        = element_text(color = ZET$slate, size = base_size - 1),
      legend.title       = element_text(color = ZET$ink,   size = base_size - 1,
                                         face = "bold"),
      legend.spacing.x   = unit(6, "pt"),

      # Facet strips — light gray pill-style
      strip.background = element_rect(fill = ZET$gridline, color = NA,
                                       linewidth = 0),
      strip.text       = element_text(color = ZET$ink, face = "bold",
                                       size = base_size - 1, margin = margin(4,6,4,6))
    )
}
```

---

## Overview

This document implements the simulation study for a two-arm Phase II oncology
design with historical borrowing on the control arm, Bayesian posterior
monitoring, a blinded nuisance-parameter sample size re-estimation (SSR) rule,
and response-adaptive randomization (RAR).

**Rule B.** At the SSR look, the pooled response rate is used only as a blinded
nuisance-parameter estimate. The planned treatment effect remains fixed at
$0.45-0.25=0.20$, and the two-arm sample size is recomputed under a prespecified
working 1:1 allocation. The rule permits increases only, from $N=90$ up to
the prespecified cap $N=150$; with these planning inputs its actual mapping at
the $n=45$ look ranges only from $N=90$ to $N=100$. Because the operational
design uses Thompson-sampling RAR and a Bayesian final decision, this is a
conventional-style working nuisance rule, not an exact power calculation for
the full adaptive design.

**Primary claim — calibration is local.** At the calibration anchor
$p=0.25$ with no trend, the full design's estimated Type I error is 0.0236
(MCSE 0.0021), consistent with the nominal 0.025 level. Holding
$\delta=0$, changing only the common baseline rate raises Type I error to
0.0810 at $p=0.30$, 0.1854 at $p=0.36$ and 0.2786 at $p=0.45$, the largest
rate evaluated. The curve is still climbing at the grid boundary — the last
three increments are $+0.0540$, $+0.0504$ and $+0.0428$.

0.2786 is the largest *estimated* value on the evaluated grid, not a
deterministic lower bound on the supremum: it carries Monte Carlo error
(MCSE 0.0063) and may lie either side of the true rejection probability at that
rate. A one-sided 95% simulation limit puts that rate's Type I error at or
above 0.2682. Because the supremum is at least the value at every evaluated
point, that pointwise limit is also a conservative lower confidence bound for
the supremum. It does not locate the supremum or provide an upper bound. The
superiority null is $H_0:p_e\le p_c$. This grid evaluates only its constant-rate
equality boundary, $p_e=p_c=p$. Finding an error rate above $\alpha$ anywhere on
that boundary is sufficient to refute uniform control, but the study does not
establish that the boundary is least favorable or characterize the interior
$p_e<p_c$. The supremum over the whole composite null is not estimated, since
the grid is finite and the curve has not flattened. Thus
calibration succeeds at
the anchor but is not portable across the evaluated constant-rate equality
boundary; this
failure does not require a calendar trend or a misspecified time-constant
analysis. Fixed-threshold ablations that disable SSR and RAR do not restore
Type I error control across the constant-rate grid either.

**The trend's contribution is not monotone in the departure.** Comparing the
two grids at matched baseline rates, the trend-active point estimate is higher
at every evaluated rate. The difference rises from $+0.0034$ at $p=0.20$ to
$+0.0396$ at $p=0.30$, dips slightly to $+0.0356$ at $p=0.33$, reaches
$+0.0440$ at $p=0.40$, and falls to $+0.0088$ at $p=0.45$. The $p=0.20$ gap
is borderline (2.0 Monte Carlo standard errors), and the $p=0.45$ gap is
indistinguishable from zero (1.0 standard error), so no claim of uniform
separation is made. The fall from $p=0.40$ to $p=0.45$ is $-0.0352$, about
2.8 standard errors, and rests on a single point at the grid boundary.

We record that as an observation about the boundary point and offer no
mechanism for it. An earlier draft of this document described the two curves as
narrowing monotonically and attributed the narrowing to the departure
saturating whatever the trend would otherwise exploit; the full nine-point grid
does not support either statement, which was drawn from the two endpoints
alone. What can be said is narrower: the trend-active point estimates are
higher throughout the evaluated range, with borderline separation at the
lowest rate and no distinguishable separation at the largest.

**Secondary finding — composition.** At the frozen full-design threshold, the
combined SSR/RAR contrast in the trend-active sweep changes sign across the
evaluated range: negative at the lower baseline rates, positive at the upper
ones. The crossing is not localised
— rows in the middle of the grid are within one Monte Carlo standard error of
zero, so the sweep establishes the sign change over the range but not the point
at which it occurs. The factorial interaction estimates are imprecise and are
not interpreted as evidence of a stable interaction in either direction. The
two estimates differ materially in magnitude; both are negative, and only the
lower-precision estimate excludes zero. Their inverse-variance pooled estimate
is -0.0077 (95% simulation interval: -0.0157 to +0.0002).

**Scope.** On the constant-rate grid, the largest estimated Type I error is
0.2786 (MCSE 0.0063) at $p=0.45$, about 11.1 times nominal; on the lower
evaluated part of the grid ($p = 0.27$ to $0.36$, i.e. departures of two to
eleven percentage points from the anchor) it runs from 0.0380 to 0.1854, up to
7.4 times nominal. The two largest rates are included to characterise the
shape of the curve; this study does not assert that they are plausible control
rates for the indication. The value $p=0.45$ additionally coincides with the
experimental-arm rate assumed under the planning alternative. That row remains
a **null**: both arms sit at 0.45, so there is no treatment effect, and the
equality is arithmetic coincidence rather than an alternative-hypothesis
evaluation. The largest trend-active
rejection probability is 0.2874 (MCSE 0.0064), but that stress scenario has a
different estimand as described above. Monte Carlo error means neither point
estimate is a deterministic lower bound on the underlying supremum. Each value
is reported with its Monte Carlo standard error, and neither finite grid locates
the true continuum supremum. Scenarios are
departures of the baseline rate from the calibration anchor $p_0 = 0.25$, not
quantified prior–data conflict. For context on the five-percentage-point
scenario, @tang_2010 treat a shift of that size as modest in an oncology
simulation of historical-control designs; their study does not validate the
design used here. @viele_2014 review conditions under which historical and
concurrent control rates diverge. Allocation targets follow the
optimal-allocation literature [@rosenberger_optimal; @rosenberger_lachin].

The Zetyra platform and its broader benchmark framework are documented in its
white paper [@zetyra_whitepaper]; related Bayesian calculators are documented
separately [@zetyra_bayesian_toolkit].

Simulation code:
`github.com/evidenceinthewild/Composed-Adaptive-Trial-Validation`

---

## 5.1 Design Parameters

```{r design-params}
# ── Trial parameters ──────────────────────────────────────────────────────────
p0       <- 0.25   # null / historical control response rate
p_alt    <- 0.45   # alternative response rate
alpha    <- 0.025  # one-sided Type I error target

# ── Sample size & looks ───────────────────────────────────────────────────────
N_base   <- 90L
N_max    <- 150L
n_looks  <- c(30L, 45L, 60L, 90L)
n_ssr    <- 45L    # SSR trigger at 50% of N_base

# ── MAP prior ────────────────────────────────────────────────────────────────
# The prior is FITTED from historical data in 5.2.1 and stored in `prior_mix`,
# a matrix with rows (w, a, b), one column per mixture component.
#
# NOTE (2026-07 correction): earlier versions declared scalar defaults
#   a_inf = 6.5; b_inf = 19.5; w_inf = 0.70; a_vag = 0.5; b_vag = 0.5
# both here AND as default arguments of simulate_one_trial(). Because R
# evaluates default arguments as literals in the function frame, the fitted
# values assigned here never reached the simulation: every result was produced
# by the hand-specified 0.70*Beta(6.5,19.5) + 0.30*Beta(0.5,0.5) mixture
# (mean 0.325), not by the RBesT MAP fit. The prior is now passed explicitly
# as a single object so the two cannot diverge.
prior_mix <- NULL   # assigned in 5.2.1; simulations require it

# ── Departures from the calibration anchor & time trend ──────────────────────
# NAMING NOTE: the variables below are still called `p0_conflict` for historical
# reasons. They are NOT quantified prior-data conflict, which would have to be
# defined against the prior predictive distribution. They are departures of the
# common baseline rate from the calibration anchor p0 = 0.25. The prose and all
# reported labels use the accurate term; only the identifiers are legacy.
#
# Departure scenarios are defined relative to the DESIGN NULL p0 = 0.25 -- the
# historical control rate the efficacy threshold is calibrated at -- NOT
# relative to the mean of the robustified MAP mixture.
#
# NOTE (2026-07 correction): these were labelled "2pp / 5pp above the prior
# mean". That is false for every candidate prior: the fitted robustified
# mixture has mean 0.3132, so p = 0.27 and p = 0.30 both lie BELOW it. The
# reference point is now named explicitly wherever these are described.
# For orientation, the relevant means are:
#   design null / historical rate       0.2500
#   dominant (informative) component    0.2568
#   full robustified mixture            0.3132
p0_conflict      <- 0.30   # +5 pp above the design null p0 = 0.25 (stress test)
p0_conflict_mild <- 0.27   # +2 pp above the design null p0 = 0.25 (headline)
delta_trend <- 0.05   # linear drift over full enrollment
# 0.40 and 0.45 are APPENDED, never inserted. Both sweeps seed per grid position
# (seed_base + 1000*i), so appending leaves i = 1..7 untouched and the seven
# previously reported points reproduce bit-identically; only the two new columns
# are new computation. Inserting a point anywhere earlier would silently reseed
# every point after it and change numbers already reviewed.
#
# Seed-block headroom: i now runs 1..9, so the blocks occupy 601000-609000,
# 701000-709000, 901000-909000 and 1001000-1009000. Still disjoint, and disjoint
# from the factorial (300000), crossing (400000) and fresh-seed (800000) blocks.
#
# Interpretive caveat: p = 0.45 is 20 percentage points above the calibration
# anchor (0.25) and about 14 above the prior mixture mean (0.3132) -- the two
# reference points are different and the distance should be named against
# whichever one is meant. It also coincides with the planning alternative
# p_alt = 0.45 -- a valid null (p_e = p_c) but not a clinically motivated one. The two added points are for locating the SHAPE of the curve
# (does Type I error plateau or keep climbing?); claims about practically
# relevant departures should stay anchored on 0.27-0.36.
p_null_grid <- c(0.20, 0.22, 0.25, 0.27, 0.30, 0.33, 0.36, 0.40, 0.45)

# ── Monte Carlo & simulation counts ──────────────────────────────────────────
# (N_mc removed: posterior probabilities are now computed by deterministic
#  quadrature rather than Monte Carlo -- see map_posterior_prob().)
N_sim_calib  <- 20000L
N_sim_main   <- 10000L
N_sim_sens   <-  5000L

# ── Monte Carlo SE helper ───────────────────────────────────────────────────
# Binomial MCSE for a proportion estimated from N_sim replications
mcse <- function(p_hat, n_sim) sqrt(p_hat * (1 - p_hat) / n_sim)

# 95% simulation interval
mcse_ci <- function(p_hat, n_sim) {
  se <- mcse(p_hat, n_sim)
  c(lower = p_hat - 1.96 * se, upper = p_hat + 1.96 * se)
}
```

---

## 5.2 Component-Level Validation

### 5.2.1 MAP Prior — Fitted-Object Internal Consistency Check [@weber_rbest]

```{r map-prior-fit}
#| warning: true
#| fig-cap: "MAP prior: fitted mixture components and the resulting density. The robustified mixture is the weighted sum of the two fitted Beta components and a flat robustify component. The maximum absolute deviation between the analytic mixture density and RBesT's dmix is printed below the figure; it is at floating-point precision and is an internal consistency check, not an external benchmark."

hist_data <- data.frame(
  study = c("H1", "H2", "H3"),
  n     = c( 50,   40,   60),
  r     = c( 13,   10,   15)
)

set.seed(42)
map_fit <- gMAP(
  cbind(r, n - r) ~ 1 | study,
  data       = hist_data,
  family     = binomial,
  tau.dist   = "HalfNormal",
  tau.prior  = 1,
  beta.prior = 2,
  chains     = 4,
  iter       = 4000,
  warmup     = 1000
)

map_diag <- posterior::summarise_draws(map_fit$draws)
map_rhat_max <- max(map_diag$rhat, na.rm = TRUE)
map_ess_bulk_min <- min(map_diag$ess_bulk, na.rm = TRUE)
map_ess_tail_min <- min(map_diag$ess_tail, na.rm = TRUE)

cat(sprintf(
  "MCMC diagnostics: max R-hat = %.4f; min bulk ESS = %.0f; min tail ESS = %.0f\n",
  map_rhat_max, map_ess_bulk_min, map_ess_tail_min))
stopifnot(map_rhat_max < 1.01,
          map_ess_bulk_min > 400,
          map_ess_tail_min > 400)

map_mix <- automixfit(map_fit, Nc = 2)
rob_mix <- robustify(map_mix, weight = 0.20, mean = 0.5)
print(rob_mix)
```

```{r map-prior-validation, fig.cap="MAP prior density: components and mixture from the fitted rob_mix object. Analytic density reproduces dmix() to floating-point precision."}
#| cache.extra: !expr list(zet_palette = ZET, frozen_md5 = tools::md5sum("map_prior_frozen.R"))
# cache.extra carries the frozen file's checksum because this chunk now READS
# that file to set prior_mix, and autodep only tracks objects passing between
# cached chunks -- it cannot see a file read. Without this key, editing
# map_prior_frozen.R would leave the cache valid and the edit would silently
# not take. zet_palette is repeated from the global cache.extra because a
# chunk-level value replaces the global one rather than extending it, and this
# chunk draws a figure.

p_seq <- seq(0.01, 0.99, length.out = 500)

# Parameters from fitted rob_mix (matrix with rows "w","a","b"; cols = components)
w_fit   <- rob_mix["w", ]   # mixture weights
a_fit   <- rob_mix["a", ]   # Beta alpha parameters
b_fit   <- rob_mix["b", ]   # Beta beta  parameters

cat("Fitted mixture components:\n")
print(rbind(w = round(w_fit, 4), a = round(a_fit, 3), b = round(b_fit, 3)))

# Analytic density using the EXTRACTED parameters
rbest_dens    <- dmix(rob_mix, p_seq)          # RBesT reference
analytic_dens <- sapply(p_seq, function(p)
  sum(w_fit * dbeta(p, a_fit, b_fit)))         # our formula with fitted params

max_dev_map <- max(abs(rbest_dens - analytic_dens))

# Approximation check against the predictive draws used by automixfit(). This is
# an in-sample fit diagnostic, distinct from the algebraic dmix() reproduction.
pred_draws <- posterior::as_draws_matrix(map_fit$draws)[, "theta_resp_pred"]
mixfit_cdf_dev <- max(abs(ecdf(pred_draws)(p_seq) - pmix(map_mix, p_seq)))

# ── Carry the FULL fitted mixture into the simulation ────────────────────────
# automixfit(Nc = 2) + robustify() returns THREE components. Earlier versions
# assumed two: they took the most informative component, treated the *first*
# leftover as "vague", and dropped the Beta(1,1) robustify() component
# entirely. The full mixture is now passed through intact.
prior_mix <- rbind(w = w_fit, a = a_fit, b = b_fit)

cat(sprintf("\nPrior carried into simulation (%d components):\n", ncol(prior_mix)))
print(round(prior_mix, 4))
cat(sprintf("  mixture mean = %.4f\n", sum(w_fit * a_fit / (a_fit + b_fit))))
stopifnot(abs(sum(w_fit) - 1) < 1e-8)

# ── The refit is a validation calculation; the frozen file is the input ──────
# The stopifnot() above is the only assertion this block used to carry, and it
# cannot fail: mixture weights sum to one by construction. Every operating
# characteristic in this study is computed against this prior, and the fit runs
# automixfit() -> robustify() on top of gMAP()/rstan, so an RBesT or Stan
# upgrade can move it -- and would move every number in the paper with no check
# going red.
#
# So the simulation is driven from a committed constant rather than from a live
# package call, and the refit above becomes the thing being validated. The
# checks themselves live in the uncached preflight chunk immediately below.
#
# Canonicalise by descending `a` first, and drop dimnames, so neither the
# comparison nor the input depends on the order or the component names that
# automixfit() and robustify() happen to return.
map_canonical <- function(m) {
  m <- m[, order(m["a", ], decreasing = TRUE), drop = FALSE]
  unname(m[c("w", "a", "b"), , drop = FALSE])
}
prior_fitted_canon <- map_canonical(prior_mix)

# Required, never silently recreated. Regenerating on absence would let a
# changed fit quietly install itself as the new reference, which is the failure
# this whole block exists to prevent.
MAP_FROZEN_PATH <- "map_prior_frozen.R"
if (!file.exists(MAP_FROZEN_PATH)) {
  stop(MAP_FROZEN_PATH, " is missing. It is a required release artifact and is ",
       "not created automatically. To create it deliberately, run once:\n",
       "    dput(prior_fitted_canon, \"", MAP_FROZEN_PATH,
       "\", control = c(\"all\", \"hexNumeric\"))\n",
       "then review the diff and commit it.")
}
prior_mix <- unname(dget(MAP_FROZEN_PATH))
dimnames(prior_mix) <- list(c("w", "a", "b"), NULL)
stopifnot(identical(dim(prior_mix), dim(prior_fitted_canon)))

# This assignment is deliberately in a CACHED chunk. autodep keys cached chunks
# on the objects they use, but its graph covers cached chunks only -- an
# uncached chunk is not a node in it. Setting prior_mix from an uncached chunk
# would leave every downstream cached result restored from the previous prior,
# with nothing to signal it.
cat(sprintf("  simulation input: %s (%d components)\n",
            MAP_FROZEN_PATH, ncol(prior_mix)))

# Component densities for plot
df_comps <- tibble(
  p     = rep(p_seq, length(w_fit) + 1),
  dens  = c(sapply(seq_along(w_fit), function(j)
               w_fit[j] * dbeta(p_seq, a_fit[j], b_fit[j])),
             rbest_dens),
  label = rep(c(sprintf("Component %d (w=%.2f)", seq_along(w_fit), w_fit),
                "Mixture (RBesT)"), each = 500)
)

comp_colors <- c("#2166AC", "#4393C3", "#D6604D", "#B2182B")
comp_labels <- levels(factor(df_comps$label))
names(comp_colors) <- comp_labels[seq_along(comp_colors)]

ggplot(df_comps, aes(p, dens, color = label, linetype = label)) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = comp_colors) +
  labs(x = bquote("Control response rate" ~ p[c]), y = "Density",
       color = NULL, linetype = NULL) +
  theme_zetyra() + theme(legend.position = "bottom")

cat(sprintf("\nMAP PRIOR:  max dev = %.2e  [floating-point precision]  %s\n",
            max_dev_map, ifelse(max_dev_map < 1e-8, "PASS", "FAIL")))
cat(sprintf("MIXTURE APPROXIMATION: max empirical-CDF dev = %.4f (%d predictive draws)\n",
            mixfit_cdf_dev, length(pred_draws)))
stopifnot(max_dev_map < 1e-8, mixfit_cdf_dev < 0.05)
```

```{r map-prior-preflight}
#| cache: false
# ── Preflight: the prior invariant, before anything expensive ────────────────
# Uncached, so a cached render cannot skip it. Placed here rather than at the
# end of the document so that a failure costs seconds instead of a 30-60 minute
# render. It reads prior_fitted_canon and prior_mix and writes nothing any later
# chunk consumes, so it does not enter the autodep graph.

# LAYER 1 -- the published representation.
# Not a tolerance. The question is precisely whether this fit still rounds to
# the four-decimal table in rerun/VERIFIED_NUMBERS.md Sec. 5, because that
# rounded matrix is what both manuscripts print. A tolerance cannot express
# that: two values can sit well within any epsilon of each other and still
# round to different fourth decimals.
MAP_REFERENCE_4DP <- unname(rbind(
  w = c(0.5887,  0.2113, 0.2000),
  a = c(18.4026, 1.9564, 1.0000),
  b = c(53.2729, 4.7033, 1.0000)
))
map_fitted_4dp <- round(prior_fitted_canon, 4)
if (!identical(map_fitted_4dp, MAP_REFERENCE_4DP)) {
  cat("refit, rounded to 4 dp:\n"); print(map_fitted_4dp)
  cat("recorded reference:\n");     print(MAP_REFERENCE_4DP)
  stop("The refitted MAP mixture no longer rounds to the published ",
       "four-decimal reference. VERIFIED_NUMBERS.md Sec. 5 and both ",
       "manuscripts print that matrix: reconcile them before rendering.")
}

# LAYER 2 -- the computational object. A DIAGNOSTIC, NOT A GATE.
#
# map_prior_frozen.R is authoritative: the simulations run on it, so a refit
# that differs cannot move a single operating characteristic. What a deviation
# means is narrower and worth stating exactly -- this environment does not
# reproduce the recorded fit. That is worth knowing and worth recording. It is
# not a reason to stop a replicator from reproducing results that do not depend
# on it: a different BLAS could shift the gMAP fit at 1e-11 and abort an
# hour-long render over a quantity with no influence on the answer.
#
# The strict requirement lives in the author-release path instead. This chunk
# writes map_prior_status.txt on every render, and tools/release/
# check_release.py asserts on it -- see the `map_prior` check. Certification
# happens on the producing machine; reproduction is not gated by it.
#
# The comparison is an explicit maximum absolute difference, described as one.
# Not identical(): the file is a dput() serialisation, committed with
# control = c("all", "hexNumeric"), which round-trips the generating doubles
# bitwise. "digits17" was tried and rounded two entries by one ulp.
MAP_FROZEN_TOL <- 1e-12
# unname() on both sides: prior_mix carries rownames because downstream code
# indexes it as prior_mix["a", ], while prior_fitted_canon is bare. identical()
# compares attributes, so without this the bitwise report would always be FALSE
# for a reason that has nothing to do with the numbers.
map_frozen_dev     <- max(abs(unname(prior_mix) - prior_fitted_canon))
map_frozen_bitwise <- identical(unname(prior_mix), prior_fitted_canon)
map_frozen_ok      <- map_frozen_dev <= MAP_FROZEN_TOL

cat("MAP prior preflight\n")
cat("  layer 1  refit rounds to the published 4 dp reference: PASS  [hard gate]\n")
cat(sprintf("  layer 2  max |frozen - refit| = %.3e  (bound %.0e)  %s  [diagnostic]\n",
            map_frozen_dev, MAP_FROZEN_TOL,
            if (map_frozen_ok) "within bound" else "OUT OF BOUND"))
cat(sprintf("  bitwise identical:  %s\n", if (map_frozen_bitwise) "yes" else "no"))
cat(sprintf("  simulation input:   %s\n", MAP_FROZEN_PATH))
cat("  results depend on the frozen file, not on the refit\n")

if (!map_frozen_ok) {
  bar <- strrep("!", 72)
  cat("\n", bar, "\n", sep = "")
  cat("MAP PRIOR REFIT DOES NOT REPRODUCE THE FROZEN MIXTURE\n")
  cat(sprintf("  max |frozen - refit| = %.3e, above the %.0e bound.\n",
              map_frozen_dev, MAP_FROZEN_TOL))
  cat("  Results in this render are UNAFFECTED: the simulations run on\n")
  cat(sprintf("  %s, not on the refit.\n", MAP_FROZEN_PATH))
  cat("  What it means is that this environment does not reproduce the\n")
  cat("  recorded fit. See SOFTWARE_ENVIRONMENT.md for the producing versions.\n")
  cat("  A release build must not be cut from this render.\n")
  cat(bar, "\n", sep = "")
  cat("frozen (simulation input):\n"); print(unname(prior_mix))
  cat("refit:\n");                     print(prior_fitted_canon)
  warning("MAP prior refit does not reproduce ", MAP_FROZEN_PATH,
          " (max dev ", signif(map_frozen_dev, 3), " > ", MAP_FROZEN_TOL,
          "). Simulation results are unaffected -- the frozen file is the ",
          "input -- but this render is not release-certifiable.",
          call. = FALSE, immediate. = TRUE)
}

# Machine-readable status for the author-release gate. Written every render, so
# the release checker asserts on a fact the render itself produced rather than
# re-deriving it without R.
writeLines(c(
  "map_prior_status v1",
  sprintf("frozen_path: %s", MAP_FROZEN_PATH),
  sprintf("bound: %.0e", MAP_FROZEN_TOL),
  sprintf("max_abs_deviation: %.17e", map_frozen_dev),
  sprintf("within_bound: %s", if (map_frozen_ok) "TRUE" else "FALSE"),
  sprintf("bitwise_identical: %s", if (map_frozen_bitwise) "TRUE" else "FALSE"),
  "rounds_to_published_4dp: TRUE",
  sprintf("rendered_utc: %s",
          format(Sys.time(), "%Y-%m-%dT%H:%M:%SZ", tz = "UTC")),
  sprintf("r_version: %s", R.version.string)
), "map_prior_status.txt")
```

### 5.2.2 Posterior Probability Helper

The monitoring rule is a posterior-probability threshold applied at each
interim look; see @zhou_ji_2024 for the design family and the question of
whether such rules require adjustment for interim analyses.

```{r posterior-prob-fn}
# ── Conjugate update of an arbitrary Beta-mixture prior ──────────────────────
# mix: matrix with rows (w, a, b), one column per component.
mixture_posterior <- function(mix, x_c, n_c) {
  w <- mix["w", ]; a <- mix["a", ]; b <- mix["b", ]
  a_post <- a + x_c
  b_post <- b + (n_c - x_c)
  lml <- lbeta(a_post, b_post) - lbeta(a, b)      # log marginal likelihood
  lw  <- log(w) + lml
  lw  <- lw - max(lw)                              # numerically stable
  w_post <- exp(lw); w_post <- w_post / sum(w_post)
  rbind(w = w_post, a = a_post, b = b_post)
}

# ── Gauss-Legendre nodes on [0,1] for deterministic integration ──────────────
# Golub-Welsch: nodes/weights are the eigenvalues/first-components of the
# symmetric tridiagonal Jacobi matrix. Base R only -- no extra dependency.
gauss_legendre_01 <- function(n) {
  k <- 1:(n - 1)
  b <- k / sqrt(4 * k^2 - 1)
  J <- diag(0, n)
  J[cbind(k, k + 1)] <- b
  J[cbind(k + 1, k)] <- b
  e <- eigen(J, symmetric = TRUE)
  ord <- order(e$values)
  x <- e$values[ord]                  # nodes on [-1, 1]
  w <- 2 * (e$vectors[1, ord])^2      # weights on [-1, 1]
  list(x = 0.5 * (x + 1), w = 0.5 * w)   # mapped to [0, 1]
}
.gl <- gauss_legendre_01(400)
.QX <- .gl$x
.QW <- .gl$w

# Sanity checks on the rule actually used below.
#
# Normalisation alone is too weak to certify a quadrature rule: a set of
# positive weights summing to 1 passes it whether or not the nodes are right.
# A standalone diagnostic written for this project built the Jacobi matrix with
# the off-diagonal terms on the MAIN diagonal, still integrated the constant 1
# to 1.0, still produced nodes inside (0, 1) -- and integrated x to 0.7887
# instead of 0.5. Moment exactness to degree 2n-1 is the property that actually
# distinguishes a Gauss-Legendre rule, so it is checked here, at render time,
# on the rule the simulation uses.
.g8 <- gauss_legendre_01(8L)
stopifnot(
  abs(sum(.QW) - 1)                     < 1e-12,   # int 1    = 1
  abs(sum(.QW * .QX) - 0.5)             < 1e-12,   # int x    = 1/2
  abs(sum(.QW * .QX^2) - 1 / 3)         < 1e-12,   # int x^2  = 1/3
  abs(sum(.g8$w * .g8$x^15) - 1 / 16)   < 1e-10,   # exact to degree 2n-1
  abs(sum(.QW * dbeta(.QX, 3, 7)) - 1)  < 1e-10)   # a smooth Beta density
rm(.g8)

# ── P(p_e > p_c | data), by deterministic quadrature ─────────────────────────
# = int f_e(x) F_c(x) dx, where p_e ~ Beta(.5+x_e, .5+n_e-x_e) (Jeffreys, no
# borrowing on the experimental arm) and p_c ~ posterior mixture.
#
# NOTE (2026-07 correction): this was previously a 2,000-draw Monte Carlo
# estimate. Because the monitoring rule takes a maximum over interim looks,
# Monte Carlo noise in the estimate can upward-bias that maximum, and identical
# trial data can yield different decisions. Deterministic 400-node Gauss-Legendre
# quadrature removes that randomness.
#
# It is DETERMINISTIC, not exact. The sanity checks above verify the
# Gauss-Legendre weights and one smooth Beta integral; they do not bound the
# approximation error for every state the adaptive design can visit. Extreme
# all-success or all-failure states can concentrate posterior mass near a
# boundary and are the most demanding cases for a fixed-node rule.
#
# All reported simulations use the specified 400-node rule. This supplement
# does not claim machine-precision agreement with a Monte Carlo reference:
# a 4e5-draw reference has Monte Carlo error of order 1e-3. Nor does it claim
# invariance of gamma* or the T1E cells to the node count; establishing that
# would require a paired rerun of calibration and each reported experiment
# under both the 400-node rule and a higher-order reference.
map_posterior_prob <- function(x_e, n_e, x_c, n_c, mix) {
  post <- mixture_posterior(mix, x_c, n_c)
  f_e <- dbeta(.QX, 0.5 + x_e, 0.5 + (n_e - x_e))
  F_c <- rowSums(vapply(seq_len(ncol(post)),
                        function(j) post["w", j] *
                          pbeta(.QX, post["a", j], post["b", j]),
                        numeric(length(.QX))))
  # Clamp to [0,1]. The integrand is a probability, but a fixed-node rule has
  # no such guarantee: at perfect separation (x_e = n_e) the Beta(n+1/2, 1/2)
  # density has an integrable singularity at 1 that 400 Gauss-Legendre nodes
  # resolve only approximately, and near-separation states can overshoot (e.g.
  # x_e = 44, n_e = 45 returns 1.0000005). The error is ~0.3-1.1% on those
  # states. The adaptive-integration reference values for those states remain
  # above gamma*, so their individual boundary decisions do not change;
  # returning a "probability" above 1 is nonetheless indefensible on inspection.
  min(max(sum(.QW * f_e * F_c), 0), 1)
}
```

### 5.2.3 Rule B — Blinded Nuisance-Parameter Re-estimation

Rule B holds the planned treatment effect fixed and re-estimates only the
nuisance event-rate level from the pooled interim responses. The reference
calculation uses a prespecified working 1:1 allocation. Its target is the
fixed-sample Wald power implied by the original $N=90$ design at
$(p_c,p_e)=(0.25,0.45)$; it is not presented as the actual power of the
Thompson-sampling RAR design or of the Bayesian monitoring rule.

The SSR decision consumes only $(x_{\mathrm{pool}},n_{\mathrm{int}})$, permits
increases only, and is capped at $N=150$. Arm labels, arm-specific outcomes,
posterior probabilities, and the realised allocation ratio are not inputs.

```{r ssr-helpers}
# Fixed planning quantities. The power target is derived from, rather than
# imposed on, the original N = 90 design.
delta_plan <- p_alt - p0
z_alpha_ssr <- qnorm(1 - alpha)
var_plan <- p0 * (1 - p0) + p_alt * (1 - p_alt)
n_arm_plan <- N_base / 2
z_power_plan <- delta_plan * sqrt(n_arm_plan / var_plan) - z_alpha_ssr
power_plan <- pnorm(z_power_plan)

# Core Rule B calculation from a blinded pooled rate. Under working 1:1
# allocation, the fixed planned effect is imposed symmetrically around p_pool.
ssr_nuisance_rate <- function(p_pool, n_int, n_base = N_base, n_max = N_max,
                              delta = delta_plan,
                              target_power = power_plan) {
  stopifnot(length(p_pool) == 1L, is.finite(p_pool),
            p_pool >= 0, p_pool <= 1,
            n_int >= 1L, n_base >= n_int, n_max >= n_base,
            delta > 0, delta < 1,
            target_power > 0, target_power < 1)

  # Truncation keeps both working rates inside (0,1) at extreme pooled counts.
  p_c_work <- min(max(p_pool - delta / 2, 1e-6), 1 - delta - 1e-6)
  p_e_work <- p_c_work + delta
  var_work <- p_c_work * (1 - p_c_work) +
              p_e_work * (1 - p_e_work)

  n_arm <- (z_alpha_ssr + qnorm(target_power))^2 * var_work / delta^2
  # Round before ceiling(). At the exact planning rates n_arm evaluates to
  # 44.999999999999986 -- 1.4e-14 below the integer -- so a one-ulp difference
  # in the qnorm/pnorm round trip across platforms or BLAS builds would flip
  # ceiling() to 46 and break the N = 90 parity assertion mid-render. Adjacent
  # attainable n_arm values differ by >= 1.4, so rounding at 1e-9 cannot change
  # any decision this rule actually makes.
  n_new <- 2L * as.integer(ceiling(round(n_arm, 9)))
  min(max(n_new, as.integer(n_base), as.integer(n_int)), as.integer(n_max))
}

# Operational interface: the only outcome input is the pooled response count.
ssr_nuisance <- function(x_int, n_int, n_base = N_base, n_max = N_max,
                         delta = delta_plan,
                         target_power = power_plan) {
  stopifnot(length(x_int) == 1L, is.finite(x_int),
            x_int >= 0, x_int <= n_int)
  ssr_nuisance_rate(x_int / n_int, n_int, n_base, n_max,
                    delta, target_power)
}

cat(sprintf("Planned effect held fixed: %.2f\n", delta_plan))
cat(sprintf("Working allocation: 1:1 (%d per arm at N = %d)\n",
            as.integer(n_arm_plan), N_base))
cat(sprintf("Reference Wald power preserved by Rule B: %.4f\n", power_plan))
```

```{r ssr-validation}
# Exhaust the complete support of the blinded count at the planned SSR look.
ssr_lookup <- tibble(
  x_pool = 0:n_ssr,
  n_int = n_ssr,
  p_pool = x_pool / n_int,
  final_N = vapply(x_pool, function(x)
    ssr_nuisance(x, n_ssr, n_base = N_base, n_max = N_max), integer(1))
)

# Independent algebra checks: exact planning rates reproduce N = 90; every
# operational decision is even, increase-only, and inside the prespecified cap.
n_at_planning_rates <- ssr_nuisance_rate(
  (p0 + p_alt) / 2, n_ssr, n_base = N_base, n_max = N_max)
stopifnot(
  n_at_planning_rates == N_base,
  all(ssr_lookup$final_N >= N_base),
  all(ssr_lookup$final_N <= N_max),
  all(ssr_lookup$final_N %% 2L == 0L)
)

# Selected blinded states make the adaptation's magnitude transparent.
ssr_demo <- ssr_lookup |>
  filter(x_pool %in% c(0L, 9L, 11L, 14L, 16L, 18L, 22L, 27L, 36L, 45L)) |>
  mutate(
    p_c_work = pmin(pmax(p_pool - delta_plan / 2, 1e-6),
                     1 - delta_plan - 1e-6),
    p_e_work = p_c_work + delta_plan
  ) |>
  select(x_pool, n_int, p_pool, p_c_work, p_e_work, final_N)

kable(ssr_demo, digits = 3, row.names = FALSE,
      caption = paste(
        "Rule B decisions at selected pooled interim counts.",
        "The planned effect remains 0.20; only the blinded nuisance level changes."),
      col.names = c("Pooled responses", "Interim N", "Pooled rate",
                    "Working p_c", "Working p_e", "Final N"))

# Exact Poisson-binomial distribution of the pooled count under each composite-
# null grid point with the active time trend. Under p_e(t) = p_c(t), this pooled
# distribution is independent of the realised treatment-allocation path.
poisson_binomial_pmf <- function(prob) {
  out <- c(1, rep(0, length(prob)))
  for (j in seq_along(prob)) {
    old <- out
    out[1] <- old[1] * (1 - prob[j])
    for (k in 2:(j + 1L))
      out[k] <- old[k] * (1 - prob[j]) + old[k - 1L] * prob[j]
  }
  out
}

ssr_size_summary <- bind_rows(lapply(p_null_grid, function(p) {
  prob <- pmin(pmax(p + delta_trend * (0:(n_ssr - 1L)) /
                      (N_base - 1L), 0), 1)
  pmf <- poisson_binomial_pmf(prob)
  final_n <- ssr_lookup$final_N
  tibble(
    p = p,
    `Pr(N > 90)` = sum(pmf[final_n > N_base]),
    `Expected N` = sum(pmf * final_n),
    `Maximum possible N` = max(final_n[pmf > 0])
  )
}))

kable(ssr_size_summary, digits = 4, row.names = FALSE,
      caption = paste(
        "Exact sample-size behavior at the n = 45 SSR look under the evaluated",
        "composite-null scenarios with the active time trend."))

cat(sprintf("Rule B validation: PASS\n"))
cat(sprintf("  N at exact planning nuisance rate: %d\n", n_at_planning_rates))
cat(sprintf("  Operational final-N range over x = 0,...,%d: %d to %d\n",
            n_ssr, min(ssr_lookup$final_N), max(ssr_lookup$final_N)))
```

### 5.2.4 RAR — Stand-Alone One-Step Check Against Analytic Allocation [@rosenberger_optimal]

This diagnostic reimplements a single Thompson-sampling allocation draw from
a fixed posterior state. It does not invoke the allocation path inside
`simulate_one_trial()` and therefore should not be read as an end-to-end check
of the production simulator.

```{r rar-validation}
#| cache: true
thompson_alloc <- function(n, p_e, p_c, a0 = 0.5, b0 = 0.5, n_burn = 10L) {
  ae <- a0; be <- b0; ac <- a0; bc <- b0; n_e <- 0L
  for (i in seq_len(n)) {
    arm <- if (i <= n_burn) (i - 1L) %% 2L else
           as.integer(rbeta(1, ae, be) >= rbeta(1, ac, bc))
    if (arm == 1L) {
      y <- rbinom(1, 1, p_e); ae <- ae + y; be <- be + (1 - y); n_e <- n_e + 1L
    } else {
      y <- rbinom(1, 1, p_c); ac <- ac + y; bc <- bc + (1 - y)
    }
  }
  n_e / n
}


# P(X > Y) where X ~ Beta(a1,b1), Y ~ Beta(a2,b2) via numerical integration
prob_beta_gt <- function(a1, b1, a2, b2) {
  integrate(function(x) dbeta(x, a1, b1) * pbeta(x, a2, b2),
            lower = 0, upper = 1, rel.tol = 1e-8)$value
}

# Validate: after observing fixed outcomes, does one-step TS allocation
# proportion match P(Beta(ae,be) > Beta(ac,bc))?
# We do n_reps single-step allocations from a known posterior state.

single_step_ts <- function(ae, be, ac, bc, n_reps = 100000L) {
  mean(rbeta(n_reps, ae, be) >= rbeta(n_reps, ac, bc))
}

set.seed(42)
# Test scenarios: known posterior states after observing specific data
# ae = 0.5 + x_e, be = 0.5 + (n_e - x_e); similarly for control
rar_scen <- tribble(
  ~x_e, ~n_e, ~x_c, ~n_c,
    5,   10,    2,   10,   # clearly favors experimental
    3,   10,    5,   10,   # clearly favors control
    5,   10,    5,   10,   # symmetric (should give ~0.5)
    8,   20,    4,   20,   # moderate asymmetry
   12,   30,    8,   30    # larger sample
)

rar_res <- rar_scen |>
  mutate(
    ae = 0.5 + x_e, be = 0.5 + (n_e - x_e),
    ac = 0.5 + x_c, bc = 0.5 + (n_c - x_c),
    # Analytical reference: P(Beta(ae,be) > Beta(ac,bc))
    ref_alloc = pmap_dbl(list(ae, be, ac, bc),
                         ~ prob_beta_gt(..1, ..2, ..3, ..4)),
    # Empirical: 100k single-step TS draws
    emp_alloc = pmap_dbl(list(ae, be, ac, bc),
                         ~ single_step_ts(..1, ..2, ..3, ..4)),
    deviation = abs(emp_alloc - ref_alloc),
    pass      = deviation < 0.005   # Monte Carlo error at n=100k is ~0.001
  )

kable(
  rar_res |> select(x_e, n_e, x_c, n_c, ref_alloc, emp_alloc, deviation, pass),
  digits = 4,
  caption = paste(
    "RAR validation: Thompson Sampling one-step allocation probability vs.",
    "analytical P(Beta(ae,be) > Beta(ac,bc)) via numerical integration.",
    "n=100k MC draws; expected Monte Carlo SE ≈ 0.001."),
  col.names = c("x<sub>e</sub>","n<sub>e</sub>","x<sub>c</sub>","n<sub>c</sub>","Analytical P(θ<sub>e</sub>>θ<sub>c</sub>)",
                "TS proportion","|dev|","Pass"),
  escape = FALSE)

cat(sprintf("RAR:  max dev = %.5f  [threshold 0.005]  %s\n",
            max(rar_res$deviation), ifelse(all(rar_res$pass), "PASS", "FAIL")))
```

---

## 5.3 Composed Pipeline Simulation

### Core rejection-probability simulation

The implementation is efficacy-only. It completes enrollment to the total
selected by Rule B, evaluates the data available at each scheduled look, and
returns the maximum posterior probability. Because allocation, outcomes, and
SSR do not depend on the efficacy boundary, thresholding this maximum gives the
same reject/no-reject indicator as prospective efficacy-only monitoring with
immediate outcomes. It does **not** reproduce operational early stopping and
therefore does not estimate efficacy stopping times, sample sizes under early
stopping, or allocations after a stopping boundary would first be crossed.

```{r sim-fn}
# simulate_one_trial() -----------------------------------------------------------
#
# Returns the maximum posterior probability across interim looks; see the note
# below. Convert to a rejection decision with t1e_at(pp, gamma).
#
# Two-arm Phase II design with historical borrowing on the control arm via a
# MAP prior, three co-occurring adaptive mechanisms:
#   MAP Bayesian monitoring, Rule B blinded nuisance SSR (if use_ssr),
#   Thompson Sampling RAR (if use_rar)
#
# Earlier-SSR timing sensitivity (NOT causal isolation -- moving SSR also
#   changes the realised final-N distribution):
#   Same monitoring schedule as the trend-active stress scenario
#   (looks = c(30,45,60,90)).
#   Same use_ssr=TRUE, use_rar=TRUE.
#   ONLY difference: n_ssr_trig=30 instead of 45.
#   => SSR co-occurs with the n=30 look (not n=45), breaking co-occurrence
#      at the richer n=45 interim where conflict and trend have accumulated.
#
#   Note this is NOT a clean isolation: moving the trigger changes the
#   information entering the rule AND the realised final-N distribution
#   together, so the monitoring looks are not taken at identical sample sizes
#   across the two arms of the comparison. See section 5.3.
#
# Factorial ablation: use_ssr x use_rar gives four conditions.
# Factor crossing: vary p_c_data (baseline-rate departure) and delta (trend).

# Returns the MAXIMUM posterior probability across scheduled looks.
#
# Data generation does not depend on gamma. Enrollment is simulated first and
# the efficacy rule is evaluated retrospectively, so rejection at threshold
# gamma is exactly {max_pp > gamma}. This preserves the efficacy-only rejection
# indicator but not operational stopping-time or post-stop allocation summaries.
# Returning max_pp lets one simulated sample yield the rejection probability at
# every threshold, which makes calibration cheap.
#
# NOTE (2026-07 correction): `mix` has no default. It previously did
# (w_inf = 0.70, a_inf = 6.5, b_inf = 19.5), and because every caller omitted
# the prior arguments, R used those literals and the fitted MAP prior was
# never actually used. Making the argument mandatory makes that failure
# impossible.
simulate_one_trial <- function(
    mix,                  # REQUIRED: fitted prior, rows (w, a, b)
    p_true_e   = 0.25,
    p_c_data   = 0.25,    # control baseline rate (departure from anchor if != p0)
    delta      = 0,       # time trend magnitude
    n_ssr_trig = 45L,     # when SSR fires (reference=45, sensitivity=30)
    use_ssr    = TRUE,    # FALSE = no sample size adaptation
    use_rar    = TRUE,    # FALSE = fixed 1:1 alternating allocation
    n_base     = 90L,
    n_cap      = 150L,
    looks      = c(30L, 45L, 60L, 90L),
    n_burn_rar = 10L) {

  stopifnot(!missing(mix), is.matrix(mix), all(c("w","a","b") %in% rownames(mix)))

  n_total  <- n_base
  ssr_done <- FALSE

  alloc_v <- integer(n_cap)
  resp_v  <- numeric(n_cap)

  ae <- 0.5; be <- 0.5   # TS posteriors (experimental)
  ac <- 0.5; bc <- 0.5   # TS posteriors (control)

  i <- 1L
  while (i <= n_total) {

    # Calendar drift on a FIXED per-patient scale.
    #
    # NOTE (2026-07 correction): this was (i - 1L) / n_total, and n_total is
    # mutated when SSR expands the trial. Normalized time therefore jumped
    # BACKWARD at the moment SSR fired (e.g. 44/90 = 0.489 -> 45/150 = 0.300),
    # so the response probability fell. The artifact was induced by SSR, which
    # contaminated the estimated SSR effect and the conflict x trend
    # interaction. The denominator is now fixed at the planned horizon: drift
    # accrues at a constant rate per patient and is invariant to adaptation.
    # Note t_frac may exceed 1 when SSR extends enrollment past n_base: drift
    # keeps accruing at the same per-patient rate, which is the intended
    # behaviour (a longer trial spans more calendar time). Clamp to [0,1] so
    # rbinom() can never receive an out-of-range probability under large delta.
    t_frac  <- (i - 1L) / (n_base - 1L)
    p_e_now <- min(max(p_true_e + delta * t_frac, 0), 1)
    p_c_now <- min(max(p_c_data + delta * t_frac, 0), 1)

    # Allocation: Thompson Sampling or fixed 1:1
    arm <- if (!use_rar || i <= n_burn_rar)
      (i - 1L) %% 2L
    else
      as.integer(rbeta(1, ae, be) >= rbeta(1, ac, bc))

    # Outcome
    if (arm == 1L) {
      y <- rbinom(1, 1, p_e_now)
      ae <- ae + y; be <- be + (1L - y)
    } else {
      y <- rbinom(1, 1, p_c_now)
      ac <- ac + y; bc <- bc + (1L - y)
    }
    alloc_v[i] <- arm
    resp_v[i]  <- y

    # SSR (fires at n_ssr_trig, uses all data up to that point)
    if (use_ssr && i == n_ssr_trig && !ssr_done) {
      ssr_done <- TRUE
      x_pool <- sum(resp_v[1:i])
      n_new  <- ssr_nuisance(x_pool, i, n_base = n_base, n_max = n_cap)
      if (n_new > n_total) n_total <- n_new
    }

    i <- i + 1L
  }

  enrolled <- i - 1L
  alloc_v  <- alloc_v[1:enrolled]
  resp_v   <- resp_v[1:enrolled]

  # Retrospective scheduled-look evaluation: same schedule regardless of
  # use_ssr / use_rar. This preserves the efficacy-only rejection indicator but
  # does not simulate operational stopping.
  actual_looks <- sort(unique(c(looks[looks <= n_base], n_total)))
  actual_looks <- actual_looks[actual_looks <= enrolled]

  cum_ne <- cumsum(alloc_v == 1L)
  cum_xe <- cumsum(resp_v  * (alloc_v == 1L))
  cum_xc <- cumsum(resp_v  * (alloc_v == 0L))

  max_pp <- 0

  for (lk in actual_looks) {
    ne <- cum_ne[lk];  xe <- cum_xe[lk]
    nc <- lk - ne;     xc <- cum_xc[lk]
    if (ne == 0L || nc == 0L) next

    # Posterior always uses all data up to this look (no stagger)
    pp <- map_posterior_prob(xe, ne, xc, nc, mix)
    if (pp > max_pp) max_pp <- pp
  }

  max_pp
}

# ── Helpers built on the max-pp representation ───────────────────────────────
# T1E at a given threshold, from a vector of per-trial max posterior probs.
t1e_at <- function(max_pp, gamma) mean(max_pp > gamma)

# Exact threshold: smallest gamma whose T1E does not exceed alpha.
#
# NOTE (2026-07 correction): gamma* was previously chosen from a 0.005-step
# grid, which left calibration T1E at 0.0290 against a 0.025 target. Reporting
# inflation relative to 0.025 then charged that calibration error to the
# scenario. Binary search removes it.
calibrate_gamma <- function(max_pp, alpha = 0.025,
                            lo = 0.80, hi = 0.9999, tol = 1e-6) {
  # The bisection assumes the target is bracketed. Unchecked it fails silently:
  # too much mass above `hi` returns hi with T1E = 1; no mass above `lo`
  # returns lo with T1E = 0. Neither looks wrong downstream.
  stopifnot(mean(max_pp > hi) <= alpha, mean(max_pp > lo) > alpha)
  for (k in seq_len(200)) {
    mid <- 0.5 * (lo + hi)
    if (mean(max_pp > mid) > alpha) lo <- mid else hi <- mid
    if (hi - lo < tol) break
  }
  hi
}

# Replicate a scenario, returning the vector of per-trial max-pp values.
#
# NOTE: do NOT write replicate(n_sim, simulate_one_trial(...)). replicate()
# is defined as sapply(integer(n), eval.parent(substitute(function(...) expr))),
# so `...` inside the expression binds to that anonymous function's dots --
# which receive the sapply loop index -- rather than to this function's dots.
# The loop index would be passed positionally as `mix`. Capture the arguments
# first and use do.call().
max_pp_sample <- function(n_sim, ...) {
  args <- list(...)
  vapply(seq_len(n_sim),
         function(i) do.call(simulate_one_trial, args),
         numeric(1))
}

# ── Re-weight the vague component of a fitted mixture ────────────────────────
# Used by the prior-weight sensitivity sweep. The vague component is the flat
# one added by robustify(); its weight is set to `w_vague` and the informative
# components share the remainder in their original proportions.
#
# The vague component is identified as the one with the smallest effective
# sample size (a + b) rather than by matching exact parameters: robustify()
# parameterises the added component by ESS, so it is Beta(1,1) or Beta(.5,.5)
# depending on the `n` argument and RBesT version.
#
# NOTE (2026-07 correction): the sweep previously passed a scalar `w_inf`
# straight into simulate_one_trial(), which -- given the default-argument
# defect -- meant it swept the weight of a Beta(6.5,19.5) that had nothing to
# do with the fitted prior.
set_vague_weight <- function(mix, w_vague) {
  ess <- mix["a", ] + mix["b", ]
  if (ncol(mix) < 2L)
    stop("mixture has a single component; nothing to re-weight")
  idx_vague <- which.min(ess)
  is_vague <- seq_len(ncol(mix)) == idx_vague
  w <- mix["w", ]
  w_new <- numeric(length(w))
  w_new[is_vague]  <- w_vague
  inf_w <- w[!is_vague]
  w_new[!is_vague] <- (1 - w_vague) * inf_w / sum(inf_w)
  stopifnot(abs(sum(w_new) - 1) < 1e-8)
  rbind(w = w_new, a = mix["a", ], b = mix["b", ])
}
```

### Threshold calibration

$\gamma^*$ is recalibrated with Rule B active under $H_0$ at the anchor
$p_0 = 0.25$ with no trend.

```{r calibration}
#| cache: true
#| fig-cap: "Monitoring threshold calibration. Simulated Type I error as a function of the efficacy threshold gamma, at the calibration anchor p = 0.25 with no time trend, targeting a nominal one-sided alpha = 0.025. The selected gamma* is marked. This is an in-sample curve: gamma* is chosen on the same sample, so the achieved value at gamma* is near-deterministic and carries no meaningful Monte Carlo interval."

# One sample of max-pp values yields T1E at every threshold, so the whole
# calibration curve and the exact gamma* come from a single simulation run.
set.seed(1L)
calib_pp <- max_pp_sample(N_sim_calib,
                          mix      = prior_mix,
                          p_true_e = p0,
                          p_c_data = p0,      # no conflict
                          delta    = 0,       # no trend
                          use_ssr  = TRUE,    # calibrate with mechanisms on
                          use_rar  = TRUE)

gamma_star  <- calibrate_gamma(calib_pp, alpha)
t1e_at_star <- t1e_at(calib_pp, gamma_star)

# Nonparametric bootstrap of the finite calibration sample. This describes the
# Monte Carlo variability of the selected threshold; downstream operating-
# characteristic intervals still condition on the realised gamma_star.
set.seed(2L)
N_calib_boot <- 500L
gamma_boot <- replicate(
  N_calib_boot,
  calibrate_gamma(sample(calib_pp, replace = TRUE), alpha))
gamma_boot_ci <- quantile(gamma_boot, c(0.025, 0.975), names = FALSE)

gamma_grid <- seq(0.90, 0.99, by = 0.005)
t1e_calib  <- sapply(gamma_grid, function(g) t1e_at(calib_pp, g))

cat(sprintf("gamma* = %.5f  (achieved alpha = %.4f, target %.4f)\n",
            gamma_star, t1e_at_star, alpha))
# No MCSE is printed here on purpose: gamma* is chosen by bisection on calib_pp
# itself, so t1e_at_star is the largest attainable value <= alpha on this very
# sample and is near-deterministic. The out-of-sample anchor estimate, which
# does carry a meaningful MCSE, is computed separately under seed 101.
cat("  (in-sample by construction; see the independent anchor validation",
    "for an out-of-sample estimate)\n")
cat(sprintf(
  "  bootstrap gamma*: SD = %.5f; 95%% interval [%.5f, %.5f] (%d resamples)\n",
  sd(gamma_boot), gamma_boot_ci[1], gamma_boot_ci[2], N_calib_boot))

tibble(gamma = gamma_grid, t1e = t1e_calib) |>
  ggplot(aes(gamma, t1e)) +
  geom_line(color = ZET$ink, linewidth = 1) +
  geom_point(size = 2, color = ZET$ink) +
  geom_hline(yintercept = alpha,      linetype = "dashed", color = ZET$red) +
  geom_vline(xintercept = gamma_star, linetype = "dashed", color = ZET$green) +
  annotate("text", x = gamma_star + 0.002, y = 0.075,
           label = sprintf("γ* = %.3f", gamma_star),
           color = ZET$green2, hjust = 0, size = 3.5, fontface = "bold") +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(x = "Efficacy threshold γ", y = "Simulated Type I error") +
  theme_zetyra()
```

```{r posterior-decision-validation}
#| cache: true
# Check representative reachable states near the calibrated boundary against
# adaptive integration. Allocation fractions of approximately 30%, 50%, and 70%
# cover balanced and materially unbalanced states at every scheduled look.
candidate_states <- bind_rows(lapply(c(30L, 45L, 60L, 90L), function(n_total) {
  ne_values <- unique(pmax(1L, pmin(n_total - 1L,
                                    as.integer(round(n_total * c(0.30, 0.50, 0.70))))))
  bind_rows(lapply(ne_values, function(ne) {
    tidyr::expand_grid(
      n_total = n_total,
      n_e = ne,
      n_c = n_total - ne,
      x_e = 0:ne,
      x_c = 0:(n_total - ne))
  }))
}))

candidate_states$q_quad <- mapply(
  function(x_e, n_e, x_c, n_c)
    map_posterior_prob(x_e, n_e, x_c, n_c, prior_mix),
  candidate_states$x_e, candidate_states$n_e,
  candidate_states$x_c, candidate_states$n_c)

boundary_states <- candidate_states |>
  mutate(distance = abs(q_quad - gamma_star)) |>
  slice_min(distance, n = 12L, with_ties = FALSE)

map_posterior_prob_ref <- function(x_e, n_e, x_c, n_c, mix) {
  post <- mixture_posterior(mix, x_c, n_c)
  integrate(
    function(p_e) {
      f_e <- dbeta(p_e, 0.5 + x_e, 0.5 + n_e - x_e)
      F_c <- rowSums(vapply(
        seq_len(ncol(post)),
        function(j) post["w", j] *
          pbeta(p_e, post["a", j], post["b", j]),
        numeric(length(p_e))))
      f_e * F_c
    },
    lower = 0, upper = 1, rel.tol = 1e-11, abs.tol = 1e-12,
    subdivisions = 1000L)$value
}

boundary_states$q_ref <- mapply(
  function(x_e, n_e, x_c, n_c)
    map_posterior_prob_ref(x_e, n_e, x_c, n_c, prior_mix),
  boundary_states$x_e, boundary_states$n_e,
  boundary_states$x_c, boundary_states$n_c)
boundary_states$abs_dev <- abs(boundary_states$q_quad - boundary_states$q_ref)
boundary_states$decision_flip <-
  (boundary_states$q_quad > gamma_star) !=
  (boundary_states$q_ref > gamma_star)

cat(sprintf(
  "Near-boundary quadrature check: max |deviation| = %.2e; decision flips = %d/%d\n",
  max(boundary_states$abs_dev), sum(boundary_states$decision_flip),
  nrow(boundary_states)))
stopifnot(max(boundary_states$abs_dev) < 1e-8,
          !any(boundary_states$decision_flip))

boundary_states |>
  transmute(
    Look = n_total, n_e, x_e, n_c, x_c,
    Quadrature = sprintf("%.8f", q_quad),
    `Adaptive integration` = sprintf("%.8f", q_ref),
    `Absolute deviation` = sprintf("%.2e", abs_dev)) |>
  kable(caption = paste(
    "Posterior-probability validation for 12 representative states nearest",
    "the calibrated boundary among balanced and 30:70/70:30 allocation grids."))
```

This is a targeted decision-boundary check over the specified allocation grids,
not an exhaustive enumeration of every state reachable under RAR. It checks
selected numerical decisions near $\gamma^*$, but it does not establish
invariance for all reachable states or under a different integration rule. All
reported operating characteristics are conditional on the specified 400-node
implementation.

```{r anchor-validation}
#| cache: true
# Fresh-seed validation is separate from the sample used to choose gamma*.
set.seed(101L)
anchor_validation_pp <- max_pp_sample(
  N_sim_sens,
  mix      = prior_mix,
  p_true_e = p0,
  p_c_data = p0,
  delta    = 0,
  use_ssr  = TRUE,
  use_rar  = TRUE)
anchor_validation_t1e <- t1e_at(anchor_validation_pp, gamma_star)
anchor_validation_ci <- mcse_ci(anchor_validation_t1e, N_sim_sens)

cat(sprintf(
  "Independent anchor validation: T1E = %.4f, MCSE = %.4f, 95%% SI [%.4f, %.4f] (N = %d)\n",
  anchor_validation_t1e,
  mcse(anchor_validation_t1e, N_sim_sens),
  anchor_validation_ci["lower"], anchor_validation_ci["upper"],
  N_sim_sens))
```

All downstream Monte Carlo standard errors and simulation intervals condition
on the selected value of $\gamma^*$. The bootstrap above quantifies threshold
selection variability from the calibration sample; it is not propagated through
each downstream simulation interval.

### Trend-active stress scenario: shared $\mathcal{F}_t$

```{r primary-sim}
#| cache: true
# TREND-ACTIVE STRESS: SSR fires at n=45 (same look as monitoring)
# Factors: baseline-rate departure (p = 0.30 vs. calibration anchor 0.25) + trend
#
# H0 SPECIFICATION: Both arms generated from p0_conflict (= 0.30).
# gamma* was calibrated at 0.25 -- this is the departure from that anchor.
# H0 (no treatment effect) is preserved: p_true_e = p_c_data = 0.30.
set.seed(2026L)
pp_primary <- replicate(
  N_sim_main,
  simulate_one_trial(mix      = prior_mix,
                     p_true_e   = p0_conflict,  # H0: same rate as control
                     p_c_data   = p0_conflict,
                     delta      = delta_trend,
                     n_ssr_trig = 45L,    # SSR fires at n=45 monitoring look
                     use_ssr    = TRUE,
                     use_rar    = TRUE))

t1e_primary   <- t1e_at(pp_primary, gamma_star)
inflation_abs <- t1e_primary - alpha
inflation_pct <- inflation_abs / alpha * 100

cat(sprintf("TREND-STRESS rejection probability: %.4f  (nominal reference %.3f)\n",
            t1e_primary, alpha))
cat(sprintf("Excess over nominal:           +%.4f  (+%.1f%%)\n",
            inflation_abs, inflation_pct))
```

```{r mild-conflict-sim}
#| cache: true

# +2 PP FROM ANCHOR: p_e = p_c = 0.27, relative to p0 = 0.25
# (the rate at which gamma* was calibrated). Not defined relative to the
# robustified mixture mean, which is 0.3132 -- above both departure scenarios.
# These rows share the active trend and all design settings with the
# trend-active stress simulation; only the common baseline rate differs.
set.seed(2028L)
pp_mild <- replicate(
  N_sim_main,
  simulate_one_trial(mix      = prior_mix,
                     p_true_e   = p0_conflict_mild,
                     p_c_data   = p0_conflict_mild,
                     delta      = delta_trend,
                     n_ssr_trig = 45L,
                     use_ssr    = TRUE,
                     use_rar    = TRUE))

t1e_mild      <- t1e_at(pp_mild, gamma_star)
inflation_mild <- t1e_mild - alpha
inflation_pct_mild <- inflation_mild / alpha * 100

cat(sprintf("+2 PP FROM ANCHOR rejection probability: %.4f  (nominal reference %.3f)\n",
            t1e_mild, alpha))
cat(sprintf("Excess over nominal:                    +%.4f  (+%.1f%%)\n",
            inflation_mild, inflation_pct_mild))
cat(sprintf("vs. +5 pp from anchor: %.4f  (+5 pp = %.4f)\n", t1e_mild, t1e_primary))

# Comparison table
tribble(
  ~Scenario, ~`Baseline p (both arms)`, ~Trend, ~`Rejection probability`,
  ~`vs. nominal`,
  "At anchor (independent validation)", "0.25", "0", anchor_validation_t1e,
    sprintf("%+.1f%%", (anchor_validation_t1e - alpha) / alpha * 100),
  "+2 pp from anchor (p=0.27)", "0.27", "0.05", t1e_mild,
    sprintf("%+.1f%%", inflation_pct_mild),
  "+5 pp from anchor (p=0.30)", "0.30", "0.05", t1e_primary,
    sprintf("%+.1f%%", inflation_pct)
) |>
  mutate(`Rejection probability` = sprintf("%.4f", `Rejection probability`)) |>
  kable(caption = paste(
    "The anchor row is an independent no-trend validation.",
    "The two departure rows share delta = 0.05 and differ only in baseline rate;",
    "comparisons with the anchor combine baseline departure and trend."),
    align = c("l","c","c","r","r"))
```

### Earlier-SSR timing sensitivity (same monitoring schedule)

This sensitivity varies *when* Rule B fires while holding the monitoring
schedule fixed.

**Timing-sensitivity design:**

| Look | Reference timing | Earlier-SSR sensitivity |
|------|-----------------|-------------------|
| n=30 | monitor | **monitor + SSR** |
| n=45 | **monitor + SSR** | monitor |
| n=60 | monitor | monitor |
| final | monitoring | monitoring |

The scheduled monitoring looks are the same in both designs; only the interim
at which SSR fires differs.

**This is a sensitivity analysis, not a clean negative control.** Moving SSR
from $n=45$ to $n=30$ also changes the data on which re-estimation is computed,
and therefore the distribution of the realised final sample size — and the final
look is taken at `n_total`, which that distribution determines. The comparison
confounds SSR timing with its consequences for final $N$. Read it as evidence on
SSR timing; isolating a co-occurrence effect would require reporting and
controlling the final-$N$ distributions.

```{r negative-control}
#| cache: true

# Earlier-SSR sensitivity: n_ssr_trig=30, not n=45
# All other parameters identical to the trend-active stress scenario
set.seed(2027L)
pp_neg <- replicate(
  N_sim_main,
  simulate_one_trial(mix      = prior_mix,
                     p_true_e   = p0_conflict,  # H0: same rate as control
                     p_c_data   = p0_conflict,
                     delta      = delta_trend,
                     n_ssr_trig = 30L,    # SSR fires at n=30 look (not n=45)
                     use_ssr    = TRUE,
                     use_rar    = TRUE))

t1e_neg       <- t1e_at(pp_neg, gamma_star)
timing_change <- t1e_neg - t1e_primary
timing_change_pct_excess <- timing_change / inflation_abs * 100

cat(sprintf("Earlier-SSR rejection probability: %.4f  (nominal reference %.3f)\n",
            t1e_neg, alpha))
cat(sprintf("Reference-timing rejection probability: %.4f\n", t1e_primary))
cat(sprintf("Timing change: %+.4f (%+.1f%% of the stress-scenario excess over nominal)\n",
            timing_change, timing_change_pct_excess))
```

```{r results-plot, fig.cap="Pipeline-level rejection probability under Rule B, for a two-arm design with the MAP prior at a baseline departure of p_c = 0.30 against the 0.25 anchor, with a shared calendar trend of 0.05: no-trend calibration target, reference SSR timing (n=45), and earlier SSR timing (n=30). Dashed red line = nominal alpha = 0.025. This is a timing sensitivity, not a clean negative control: moving SSR also changes the realised final-N distribution. It does not apportion the excess among components."}
tibble(
  Scenario = factor(
    c("Calibration\n(at anchor,\nno trend)",
      "Trend stress\n(SSR at n = 45)",
      "Earlier SSR\n(n = 30)"),
    levels = c("Calibration\n(at anchor,\nno trend)",
               "Trend stress\n(SSR at n = 45)",
               "Earlier SSR\n(n = 30)")),
  T1E  = c(t1e_at_star, t1e_primary, t1e_neg),
  fill = c("a", "b", "c")
) |>
  ggplot(aes(Scenario, T1E, fill = fill)) +
  geom_col(width = 0.50) +
  geom_hline(yintercept = alpha, linetype = "dashed",
             color = ZET$red, linewidth = 0.8) +
  # Value labels clear the nominal-alpha line. A bar whose top sits within a
  # label height of alpha would otherwise have the dashed line struck through
  # its own label -- which happened to the 0.0232 and 0.0248 bars, both of which
  # land within 0.002 of 0.025. y is mapped explicitly (vjust = 0, bottom-
  # aligned) rather than nudged, so the offset is in data units and survives a
  # change of y-limit.
  geom_text(aes(y = T1E + ifelse(abs(T1E - alpha) < 0.006, 0.0080, 0.0020),
                label = sprintf("%.4f", T1E)),
            vjust = 0, size = 3.8, fontface = "bold", color = ZET$ink) +
  scale_fill_manual(values = c(a = ZET$neutralF, b = ZET$green, c = ZET$teal)) +
  scale_y_continuous(limits = c(0, max(t1e_primary, t1e_neg, 0.04) * 1.30),
                     labels = percent_format(accuracy = 0.1)) +
  labs(x = NULL, y = "Rejection probability") +
  theme_zetyra() + theme(legend.position = "none")
```

```{r results-table}
tribble(
  ~Scenario, ~`Rejection probability`, ~`vs. nominal`, ~Note,
  "Calibration (at anchor, no trend)", t1e_at_star, "—", "threshold target",
  "Trend stress: SSR at n = 45", t1e_primary,
    sprintf("%+.1f%%", inflation_pct), "SSR at the n=45 look",
  "Earlier SSR: n = 30", t1e_neg,
    sprintf("%+.1f%%", (t1e_neg - alpha)/alpha*100), "SSR at the n=30 look"
) |>
  mutate(`Rejection probability` =
           sprintf("%.4f", `Rejection probability`)) |>
  kable(caption = paste("Rejection-probability summary under Rule B.",
                        "The trend-active rows are contemporaneous no-effect",
                        "scenarios under a time-constant analysis; the n = 30",
                        "row is an SSR-timing sensitivity, not a causal isolation."),
        align = c("l","r","r","l"))
```

---

## 5.4 Fixed-Threshold Factorial Ablation

Running four designs (SSR × RAR) shows what happens when each adaptive component
is switched off.

**These are ablations at a fixed threshold, not a decomposition.** $\gamma^*$ was
calibrated with SSR and RAR *active*. The ablation cells retain that threshold,
so they are not calibrated to $\alpha$. Do not read the baseline cell as an
attributable share.

**The SSR-enabled cells can accrue an additional efficacy analysis.** The
monitoring schedule is $\{30, 45, 60, 90\}$ together with the *realised* final
sample size, so a trial that SSR extends beyond $N = 90$ is analysed five times
where a fixed-$N$ trial is analysed four. Because the test statistic is a
maximum over looks, the SSR-enabled cells get one further opportunity to cross
the threshold in the $\Pr(N > 90)$ fraction of replicates — 0.3157 at
$p = 0.30$, and 0.9403 at $p = 0.45$ (§5.2.3). This is correct trial conduct: a
trial that genuinely enrolls more patients does analyse at its realised final
$N$. But it means every full-design-versus-ablation contrast in this document
combines the adaptive mechanisms with a difference in multiplicity, and no such
contrast isolates the mechanisms alone.

All four conditions use the same monitoring threshold $\gamma^*$ and the same
baseline-rate departure and time trend as the trend-active stress scenario.

```{r factorial}
#| cache: true
#| fig-cap: "Fixed-threshold factorial ablation of SSR by RAR under Rule B, at p = 0.30 with the shared trend active. The two simple contrasts and the interaction point estimate are printed immediately below the figure. The displayed interaction is a point estimate; read it against the propagated 95% Monte Carlo simulation interval and the independent fresh-seed replicate reported below, both of which are given immediately after this figure. These are fixed-threshold ablations and cannot apportion error to a component."

cat("Fixed-threshold factorial ablation: use_ssr x use_rar\n")
# Four conditions: 2x2 SSR x RAR
factorial_grid <- expand.grid(use_ssr = c(FALSE, TRUE),
                               use_rar = c(FALSE, TRUE))

# Per-cell seeds. A single seed before apply() draws all four cells from one
# sequential stream, and the SSR-on cells consume a design-dependent number of
# draws, so each cell shifts the stream for the next. The interaction is a
# four-cell contrast, so that offset enters it directly.
t1e_factorial <- vapply(seq_len(nrow(factorial_grid)), function(k) {
  row <- factorial_grid[k, ]
  set.seed(300000L + 1000L * k)
  t1e_at(replicate(N_sim_sens,
    simulate_one_trial(mix    = prior_mix,
                       p_true_e = p0_conflict,  # H0: p_e = p_c = p0_conflict
                       p_c_data = p0_conflict,
                       delta    = delta_trend,
                       use_ssr  = as.logical(row[["use_ssr"]]),
                       use_rar  = as.logical(row[["use_rar"]]))), gamma_star)
}, numeric(1))

factorial_res <- factorial_grid |>
  mutate(T1E = t1e_factorial,
         label = case_when(
           !use_ssr & !use_rar ~ "Baseline\n(no SSR, no RAR)",
           use_ssr  & !use_rar ~ "SSR only\n(fixed allocation)",
           !use_ssr & use_rar  ~ "RAR only\n(no SSR)",
           # Was "(trend stress)", which read as though the other three cells
           # were not trend-active. All four are p = 0.30 with delta = 0.05;
           # the shared scenario belongs in the subtitle, not on one bar.
           use_ssr  & use_rar  ~ "SSR + RAR\n(both active)"
         ))

# Factorial effects — prefixed _f_ to avoid collision with sensitivity t1e_ssr
t1e_f_base <- factorial_res$T1E[!factorial_res$use_ssr & !factorial_res$use_rar]
t1e_f_ssr  <- factorial_res$T1E[ factorial_res$use_ssr & !factorial_res$use_rar]
t1e_f_rar  <- factorial_res$T1E[!factorial_res$use_ssr &  factorial_res$use_rar]
t1e_f_both <- factorial_res$T1E[ factorial_res$use_ssr &  factorial_res$use_rar]

# Compute deltas directly from factorial_res — no intermediate variables
# that could be shadowed by later chunks
delta_f_ssr  <- t1e_f_ssr  - t1e_f_base
delta_f_rar  <- t1e_f_rar  - t1e_f_base
delta_f_int  <- t1e_f_both - t1e_f_ssr - t1e_f_rar + t1e_f_base

# Report the CELL VALUES first, then the contrasts.
#
# NOTE (2026-07 correction): earlier versions printed only the three contrasts.
# That concealed the single most important number in the study -- the baseline
# cell, in which Bayesian monitoring runs with NO adaptive component. If the
# baseline is already far above alpha, then switching the adaptive components
# off does not return the rejection probability to the nominal level. These
# cells are trend active, so the elevated quantity is a rejection probability
# rather than a Type I error. This cell also keeps gamma*, which was calibrated
# with SSR+RAR active, so it is an ablation, not an attributable share.
cat("Cell values (rejection probability):\n")
cat(sprintf("  SSR + RAR disabled     = %.4f  (MCSE %.4f)\n",
            t1e_f_base, mcse(t1e_f_base, N_sim_sens)))
cat(sprintf("  + SSR only             = %.4f  (MCSE %.4f)\n",
            t1e_f_ssr,  mcse(t1e_f_ssr,  N_sim_sens)))
cat(sprintf("  + RAR only             = %.4f  (MCSE %.4f)\n",
            t1e_f_rar,  mcse(t1e_f_rar,  N_sim_sens)))
cat(sprintf("  + SSR and RAR          = %.4f  (MCSE %.4f)\n",
            t1e_f_both, mcse(t1e_f_both, N_sim_sens)))
cat(sprintf("\n  NOTE: ablation cells retain the full-design threshold and are\n"))
cat(sprintf("        not separately calibrated; do not read as attributable shares.\n"))
cat("\nContrasts:\n")
cat(sprintf("  Δ_SSR         = %+.4f\n", delta_f_ssr))
cat(sprintf("  Δ_RAR         = %+.4f\n", delta_f_rar))
cat(sprintf("  Δ_interaction = %+.4f  (point estimate; MC uncertainty below)\n",
            delta_f_int))

# Bar chart
factorial_res |>
  mutate(label = factor(label, levels = label)) |>
  ggplot(aes(label, T1E, fill = use_ssr & use_rar)) +
  geom_col(width = 0.52) +
  geom_hline(yintercept = alpha, linetype = "dashed",
             color = ZET$red, linewidth = 0.8) +
  # Value labels clear the nominal-alpha line. A bar whose top sits within a
  # label height of alpha would otherwise have the dashed line struck through
  # its own label -- which happened to the 0.0232 and 0.0248 bars, both of which
  # land within 0.002 of 0.025. y is mapped explicitly (vjust = 0, bottom-
  # aligned) rather than nudged, so the offset is in data units and survives a
  # change of y-limit.
  geom_text(aes(y = T1E + ifelse(abs(T1E - alpha) < 0.006, 0.0080, 0.0020),
                label = sprintf("%.4f", T1E)),
            vjust = 0, size = 3.8, fontface = "bold", color = ZET$ink) +
  scale_fill_manual(values = c(`FALSE` = ZET$neutralF, `TRUE` = ZET$green)) +
  scale_y_continuous(limits = c(0, max(t1e_factorial) * 1.30),
                     labels = percent_format(accuracy = 0.1)) +
  labs(x = NULL, y = "Rejection probability",
       subtitle = expression(paste(
         "All four cells: ", italic(p), " = 0.30, ", delta,
         " = 0.05, fixed full-design threshold ", gamma, "*"))) +
  theme_zetyra() + theme(legend.position = "none")
```

```{r factorial-table}
tribble(
  ~Condition, ~`Rejection probability`, ~`vs. baseline`,
  "Baseline (no SSR, no RAR)", t1e_f_base, "—",
  "SSR only (fixed allocation)", t1e_f_ssr, sprintf("%+.4f", delta_f_ssr),
  "RAR only (no SSR)",          t1e_f_rar, sprintf("%+.4f", delta_f_rar),
  "SSR + RAR (both active)", t1e_f_both,
    sprintf("%+.4f", t1e_f_both - t1e_f_base)
) |>
  mutate(`Rejection probability` =
           sprintf("%.4f", `Rejection probability`)) |>
  kable(caption = paste("Fixed-threshold factorial ablation. Interaction point estimate =",
                         sprintf("%.4f", delta_f_int),
                         "; uncertainty and a fresh-seed check are reported below."),
        align = c("l","r","r"))
```

The baseline-referenced contrasts are descriptive point estimates. Their
approximate Monte Carlo standard errors are 0.0064--0.0065, and none is used as
evidence of a directional component effect.

---

### Fresh-seed check on the SSR × RAR interaction

The interaction is the one composition result whose magnitude matters, so it is
re-estimated here on an independent seed with an independent replication count.
This is a repeat within the same R implementation, not an independent
reimplementation: it guards against a seed artefact, not against a shared
modelling error.

```{r factorial-fresh-seed}
#| cache: true
# This chunk carried the same defect the initial factorial run was fixed for: one
# set.seed() before a four-cell loop, with the SSR-on cells consuming a
# design-dependent number of draws, so each subsequent cell started at a
# different stream offset. That made it useless as a check on the initial run --
# it cannot adjudicate a discrepancy while sharing the suspected cause. Per-cell
# seeds now, on a block (800000) disjoint from the initial factorial (300000),
# the crossing (400000) and the sweep (600000/700000).
#
# N raised from 3,000. At 3,000 the interaction MCSE is 0.0118, which cannot
# resolve an effect of the size the initial run reports (-0.0240); 20,000 gives
# roughly 0.0045. Cost is 80,000 trials -- lower it if that is too slow, but
# anything under ~10,000 per cell cannot settle the question it is being asked.
N_fresh <- 20000L

fresh <- vapply(seq_len(4), function(k) {
  us <- k %in% c(2, 4)          # SSR on for cells 2 and 4
  ur <- k %in% c(3, 4)          # RAR on for cells 3 and 4
  set.seed(800000L + 1000L * k)
  t1e_at(replicate(N_fresh,
    simulate_one_trial(mix = prior_mix,
                       p_true_e = p0_conflict, p_c_data = p0_conflict,
                       delta = delta_trend,
                       use_ssr = us, use_rar = ur)), gamma_star)
}, numeric(1))

names(fresh) <- c("base", "ssr", "rar", "both")
int_fresh <- fresh["both"] - fresh["ssr"] - fresh["rar"] + fresh["base"]
se_fresh  <- sqrt(sum(mcse(fresh, N_fresh)^2))

cat(sprintf(
  "Fresh-run per-cell seeds: 800000 + 1000*k, k = 1,...,4; N = %d per cell\n",
  N_fresh))
for (nm in names(fresh))
  cat(sprintf("  %-5s = %.4f  (MCSE %.4f)\n", nm, fresh[nm], mcse(fresh[nm], N_fresh)))
cat(sprintf("\n  interaction = %+.4f  (MCSE %.4f)  95%% SI [%+.4f, %+.4f]\n",
            int_fresh, se_fresh, int_fresh - 1.96*se_fresh, int_fresh + 1.96*se_fresh))
cat(sprintf("  initial-run interaction for comparison: %+.4f\n", delta_f_int))
```

The two runs do **not** agree, and the disagreement is the point of reporting
both. The reconciliation below is computed rather than asserted.

```{r factorial-reconcile}
#| code-fold: false
# Inverse-variance combination of the two independent estimates of the same
# quantity, plus a one-degree-of-freedom heterogeneity test. Neither run is
# discarded; the weights follow from their precisions.
# The initial-run SE is rebuilt here from its own four cells because knitr
# evaluates in document order. This is the same construction used for se_fresh.
se_initial <- sqrt(sum(mcse(c(t1e_f_base, t1e_f_ssr, t1e_f_rar, t1e_f_both),
                           N_sim_sens)^2))

# unname() is load-bearing: int_fresh inherits the name "both" from
# fresh["both"], so c(replicate = int_fresh) yields the element name
# "replicate.both", which then fails to index ses and silently prints NA.
ests <- c(initial = unname(delta_f_int), replicate = unname(int_fresh))
ses  <- c(initial = unname(se_initial),  replicate = unname(se_fresh))
stopifnot(identical(names(ests), names(ses)), !anyNA(c(ests, ses)))
w    <- 1 / ses^2
pooled <- sum(w * ests) / sum(w)
se_p   <- 1 / sqrt(sum(w))
Q      <- sum(w * (ests - pooled)^2)

for (nm in names(ests))
  cat(sprintf("  %-10s = %+.4f  (MCSE %.4f)  95%% SI [%+.4f, %+.4f]\n",
              nm, ests[nm], ses[nm],
              ests[nm] - 1.96 * ses[nm], ests[nm] + 1.96 * ses[nm]))
cat(sprintf("\n  pooled     = %+.4f  (SE %.4f)  95%% SI [%+.4f, %+.4f]\n",
            pooled, se_p, pooled - 1.96 * se_p, pooled + 1.96 * se_p))
cat(sprintf("  separation between the two runs: %.2f SE\n",
            abs(diff(ests)) / sqrt(sum(ses^2))))
cat(sprintf("  heterogeneity Q = %.2f on 1 df, p = %.3f\n",
            Q, pchisq(Q, 1, lower.tail = FALSE)))
```

The replicate carries four times the information of the initial run and is
compatible with zero; the initial run, at $N=5{,}000$ per cell, is not. The two
sit about two standard errors apart, which for a single pair of estimates is
near the edge of what independent sampling produces on its own and is not by
itself evidence of a defect. The pooled estimate is also compatible with zero.

The operative conclusion is unchanged and now rests on the higher-precision
evidence: **Rule B does not support a factorial SSR-by-RAR interaction claim in
either direction.** The initial run's interval excluding zero is a
low-precision result that the better-powered replicate does not reproduce, and
it should not be quoted on its own. Because the reported cells are
fixed-threshold ablations, none of these quantities apportions rejection
probability to a component in any case.

---

## 5.5 Baseline-Rate Departure × Time Trend

The trend-active stress scenario activates both factors simultaneously. Here we
*cross* them in a 2×2 design (baseline-rate departure × time trend) to see how each
changes rejection probability, and whether their joint contrast differs from the
additive prediction.

This is a factor-crossing experiment, **not** mechanism isolation. Both factors feed
the monitoring rule, the SSR trigger and the RAR allocation rule simultaneously, and
the design does not separate those channels. We also describe the departures as
departures from the calibration anchor $p_0 = 0.25$, not as quantified prior–data
conflict, which would have to be defined against the prior predictive distribution.

```{r departure-trend-crossing}
#| cache: true
#| fig-cap: "Rejection probability in the baseline-rate-departure by time-trend crossing under Rule B. Trend-active cells impose contemporaneous equality between arms but use a time-constant analysis. The annotated interaction is a point estimate; its propagated 95% Monte Carlo simulation interval includes zero."

cat("Factor crossing: 2x2 baseline-rate departure x trend\n")
mech_grid <- expand.grid(
  conflict = c(FALSE, TRUE),
  trend    = c(FALSE, TRUE)
)

# Per-cell seeds, for the same reason as the factorial: the four cells differ
# in how often SSR fires, so a shared sequential stream leaves each cell at a
# different offset and contaminates the four-cell interaction contrast.
# Indexing with [[ ]] on the data frame row also avoids apply()'s coercion of
# the logical grid to character.
t1e_mech <- vapply(seq_len(nrow(mech_grid)), function(k) {
  row <- mech_grid[k, ]
  set.seed(400000L + 1000L * k)
  p_c <- if (row[["conflict"]]) p0_conflict else p0   # 0.30 vs 0.25
  d   <- if (row[["trend"]])    delta_trend  else 0    # 0.05 vs 0
  # H0: p_true_e = p_c_data (both arms at same rate, MAP prior is what conflicts)
  t1e_at(replicate(N_sim_sens,
    simulate_one_trial(mix    = prior_mix,
                       p_true_e = p_c,   # H0 preserved: same rate as control
                       p_c_data = p_c,
                       delta    = d,
                       use_ssr  = TRUE,
                       use_rar  = TRUE)), gamma_star)
}, numeric(1))

mech_res <- mech_grid |>
  mutate(T1E   = t1e_mech,
         label = case_when(
           !conflict & !trend ~ "Anchor, no trend\n(independent run)",
           conflict  & !trend ~ "Departure only\n(no trend)",
           !conflict & trend  ~ "Trend only\n(at anchor)",
           conflict  & trend  ~ "Both\n(trend stress)"
         ))

# Additive prediction vs. actual
t1e_neither  <- mech_res$T1E[!mech_res$conflict & !mech_res$trend]
t1e_conf_only<- mech_res$T1E[ mech_res$conflict & !mech_res$trend]
t1e_trend_only<-mech_res$T1E[!mech_res$conflict &  mech_res$trend]
t1e_both_mech<- mech_res$T1E[ mech_res$conflict &  mech_res$trend]

additive_pred <- t1e_conf_only + t1e_trend_only - t1e_neither
interaction_term <- t1e_both_mech - additive_pred

cat(sprintf("Departure only:   %.4f  (+%.4f vs. anchor run)\n",
            t1e_conf_only,  t1e_conf_only  - t1e_neither))
cat(sprintf("Trend only:       %.4f  (+%.4f vs. anchor run)\n",
            t1e_trend_only, t1e_trend_only - t1e_neither))
cat(sprintf("Additive prediction: %.4f\n", additive_pred))
cat(sprintf("Observed (both):     %.4f\n", t1e_both_mech))
cat(sprintf("Interaction term:    %+.4f  (point estimate; MC uncertainty below)\n",
            interaction_term))

mech_res |>
  mutate(label = factor(label, levels = label)) |>
  ggplot(aes(label, T1E,
             fill = factor(conflict + trend * 2,
                           labels = c("Neither","Departure","Trend","Both")))) +
  geom_col(width = 0.52) +
  geom_hline(yintercept = alpha, linetype = "dashed",
             color = ZET$red, linewidth = 0.8) +
  geom_hline(yintercept = additive_pred, linetype = "dotted",
             color = ZET$slate, linewidth = 1.0) +
  # ── in-panel annotation ─────────────────────────────────────────────────────
  # This figure carries no title or subtitle: the manuscript prints a caption
  # beneath every figure, so a title only repeats it. What the subtitle used to
  # carry is annotated inside the panel instead.
  #
  # The three positions below were checked against the bar tops, their value
  # labels and the legend box at the rendered 9 x 5.5 in size; each has under
  # one line height of slack in at least one direction. Re-check before nudging.
  #
  # Additive-prediction label: left-hand end, sitting just above the dotted line
  # it names rather than floating mid-gap, so the label and its line read as one
  # object. Offset is in data units off additive_pred, not a fixed y.
  annotate("text", x = 0.55, y = additive_pred + 0.0035,
           label = sprintf("Additive prediction: %.4f", additive_pred),
           size = 4.1, fontface = "bold", color = ZET$ink, hjust = 0) +
  # The arc runs from the additive-prediction label to the observed value on the
  # "Both" bar. Its two endpoints are the two numbers whose difference IS the
  # interaction, so the span carries the meaning and the label above it only has
  # to name the quantity. Both endpoints sit above every bar, so the path stays
  # clear however the bar heights move; offsets are taken off the data rather
  # than hardcoded for the same reason.
  # Both endpoints are referenced to additive_pred rather than to one endpoint
  # each. The two quantities differ by the interaction, which is ~0.0006 here, so
  # anchoring the arc to t1e_both_mech at one end made it visually level anyway
  # while leaving it free to collide with the dotted line if the gap ever widens.
  # Riding above additive_pred keeps a fixed clearance over the line, the "Both"
  # bar and that bar's value label at the same time.
  # The arc runs number to number: it starts just right of the "0.1146" inside
  # the additive-prediction label and ends just left of the "0.1140" above the
  # Both bar. Those two numbers are what the interaction is the difference of,
  # and they happen to sit at almost the same height (0.1181 vs 0.1180), so the
  # span reads as a comparison rather than as a rise. The endpoints track the
  # label heights, not the bar tops, so they stay attached if the values move.
  annotate("curve", x = 1.80, y = additive_pred + 0.0035,
           xend = 3.68, yend = t1e_both_mech + 0.0040,
           curvature = -0.16, angle = 90, ncp = 12,
           arrow = arrow(length = unit(0.18, "cm"), type = "closed"),
           color = ZET$green, linewidth = 0.6) +
  annotate("text", x = 2.74, y = additive_pred + 0.0270,
           label = sprintf("Interaction point estimate = %+.4f", interaction_term),
           size = 4.1, fontface = "bold", color = ZET$green, hjust = 0.5) +
  # Short form for the nominal-alpha line: the caption spells it out in full,
  # and only the gap between the first two bars is wide enough for a label.
  # x = 1.68, not 1.5: the first bar's value label is centred at x = 1 and sits
  # at almost exactly this height, so 1.5 left the two nearly touching.
  annotate("text", x = 1.68, y = alpha, label = "α = 0.025",
           hjust = 0.5, vjust = -0.45, size = 3.5,
           fontface = "bold", color = ZET$red) +
  # Value labels clear the nominal-alpha line. A bar whose top sits within a
  # label height of alpha would otherwise have the dashed line struck through
  # its own label -- which happened to the 0.0232 and 0.0248 bars, both of which
  # land within 0.002 of 0.025. y is mapped explicitly (vjust = 0, bottom-
  # aligned) rather than nudged, so the offset is in data units and survives a
  # change of y-limit.
  geom_text(aes(y = T1E + ifelse(abs(T1E - alpha) < 0.006, 0.0080, 0.0020),
                label = sprintf("%.4f", T1E)),
            vjust = 0, size = 3.8, fontface = "bold", color = ZET$ink) +
  # No key. The fill maps to the same four categories the x axis already names,
  # in the same order, so it encodes nothing position does not -- the legend
  # entries were just abbreviated tick labels. The fills stay: they separate the
  # reference condition from the composed one at a glance. Contrast Figure 2,
  # where two series share one pair of axes and colour is the only thing telling
  # them apart; that legend is load-bearing and stays.
  scale_fill_manual(values = c(Neither  = ZET$neutralF,
                                Departure = "#2563EB",
                                Trend    = ZET$teal,
                                Both     = ZET$green),
                    guide = "none") +
  # 1.20 was sized for the older, lower arc. The arc now rides above the dotted
  # additive line and the interaction label sits above the arc, so the label
  # needs about 1.27; 1.30 leaves it roughly a line of headroom without opening
  # a band of empty panel above it. Checked in the matplotlib preview.
  scale_y_continuous(limits = c(0, max(t1e_mech) * 1.34),
                     labels = percent_format(accuracy = 0.1)) +
  labs(x = NULL, y = "Rejection probability", fill = NULL) +
  theme_zetyra() +
  theme(axis.title = element_text(size = 15, face = "bold", color = ZET$ink,
                                  margin = margin(t = 4, r = 4)),
        axis.text  = element_text(size = 13, color = ZET$slate))
```

```{r departure-trend-table}
tribble(
  ~Condition, ~`Rejection probability`, ~`vs. anchor run`, ~Note,
  "Anchor, no trend (independent run)", t1e_neither, "—", "independent reference cell",
  "Departure only (no trend)", t1e_conf_only,
    sprintf("%+.4f", t1e_conf_only - t1e_neither),  "baseline away from calibration anchor",
  "Trend only (at anchor)", t1e_trend_only,
    sprintf("%+.4f", t1e_trend_only - t1e_neither), "rate drifts from anchor",
  "Both (trend stress)",      t1e_both_mech,
    sprintf("%+.4f", t1e_both_mech - t1e_neither),  sprintf("interaction = %+.4f", interaction_term)
) |>
  mutate(`Rejection probability` =
           sprintf("%.4f", `Rejection probability`)) |>
  kable(caption = paste("2x2 crossing of baseline-rate departure and time trend.",
                        "Both factors feed the full pipeline, so this",
                        "crosses factors rather than isolating mechanisms."),
        align = c("l","r","r","l"))
```

---

## 5.6 Rejection Probability Across the Common-Rate Grid

For superiority, $H_0:p_e\le p_c$; the common-rate family $p_e=p_c=p$ is its
equality boundary. The first sweep sets $\delta=0$ and estimates pointwise Type
I error along that boundary. A violation at any evaluated point is enough to
disprove uniform control, but this sweep neither establishes the least-favorable
configuration nor evaluates the interior $p_e<p_c$. A second, explicitly
separate sweep retains the shared
calendar trend used in the trend-active stress scenario. In that DGM,
$p_e(i)=p_c(i)$ for every enrollment index $i$, but the arms need not sample
the same calendar periods. With RAR disabled, the implemented rule
`arm = (i - 1) %% 2` gives deterministic strict alternation: control occupies
odd enrollment positions and experimental treatment occupies even positions.
By construction, this balances arm counts at even looks and tightly interleaves
enrollment time, but it is not exact time balance under the coded patient-level
trend. At every
even look, experimental patients are systematically one position later within
each pair, so their average generating rate exceeds control by
$\delta/(N_{\mathrm{base}}-1)=0.05/89\approx0.00056$; at the odd $n=45$ look,
the arm counts differ by one. With RAR active, allocation additionally depends
on accumulating data. The analysis remains time-constant in both cases. We
therefore report these values as rejection probabilities under contemporaneous
no treatment effect and analysis-model misspecification, not as Type I error
for the paper's constant marginal-rate null.

Both sweeps hold $\gamma^*$ at the value calibrated for the full design at
$p=0.25$ and $\delta=0$. The comparison series merely disables SSR and RAR; it
is a fixed-threshold ablation, not a separately calibrated monitoring-only
design.

Three caveats govern how the sweeps should be read.

**Finite-grid contrasts are not sharply identified.** Each row is a Monte Carlo
estimate and the grid is discrete. Locating a crossing in a fixed-$\gamma^*$
contrast would require a denser grid and preferably paired common-random-number
contrasts; the tables support statements only about the evaluated points.

**The SSR rule shares a threshold with the sweeps.** SSR is a pooled blinded
nuisance-parameter rule under a working 1:1 Wald calculation, while the
operational allocation is response-adaptive and the final estimand is the
Bayesian two-arm comparison $p_e>p_c$. Any contrast involving SSR therefore
belongs to this prespecified working rule and should not be generalized to all
blinded SSR procedures.

**Extended trials accrue more drift in the stress sweep.** With drift anchored
to the planned horizon $n_{\text{base}}=90$, Rule B's largest attainable final
sample size of $n=100$ experiences terminal drift
$0.05\times99/89\approx0.0556$, rather than $0.05$ at the planned terminal
position. A longer trial therefore spans more calendar time. This is part of
the stated trend-active DGM, but it is absent from the constant-rate
equality-boundary sweep.

```{r composite-null-sweep}
#| cache: true

# Each series uses disjoint per-cell seeds. The trend-active seed blocks preserve
# the prior analysis; the static-null blocks are new and non-overlapping.
run_null_sweep <- function(delta, seed_disabled, seed_full) {
  bind_rows(lapply(seq_along(p_null_grid), function(i) {
    pn <- p_null_grid[i]
    set.seed(seed_disabled + 1000L * i)
    disabled <- t1e_at(replicate(
      N_sim_sens,
      simulate_one_trial(
        mix = prior_mix, p_true_e = pn, p_c_data = pn, delta = delta,
        use_ssr = FALSE, use_rar = FALSE)), gamma_star)

    set.seed(seed_full + 1000L * i)
    full <- t1e_at(replicate(
      N_sim_sens,
      simulate_one_trial(
        mix = prior_mix, p_true_e = pn, p_c_data = pn, delta = delta,
        use_ssr = TRUE, use_rar = TRUE)), gamma_star)

    tibble(
      p = pn,
      `SSR + RAR disabled` = disabled,
      `All mechanisms` = full,
      `Fixed-gamma* difference` = full - disabled,
      `MCSE (ablation)` = mcse(disabled, N_sim_sens),
      `MCSE (full)` = mcse(full, N_sim_sens),
      `MCSE (difference)` = sqrt(
        mcse(disabled, N_sim_sens)^2 + mcse(full, N_sim_sens)^2))
  }))
}

composite_null_static <- run_null_sweep(
  delta = 0, seed_disabled = 900000L, seed_full = 1000000L)
contemporaneous_null_trend <- run_null_sweep(
  delta = delta_trend, seed_disabled = 600000L, seed_full = 700000L)

format_null_sweep <- function(x) {
  x |>
    mutate(
      across(
        c(`SSR + RAR disabled`, `All mechanisms`,
          `MCSE (ablation)`, `MCSE (full)`, `MCSE (difference)`),
        ~ sprintf("%.4f", .x)),
      `Fixed-gamma* difference` =
        sprintf("%+.4f", `Fixed-gamma* difference`))
}

format_null_sweep(composite_null_static) |>
  kable(caption = paste(
    "Pointwise Type I error along the constant-rate equality boundary",
    "(p_e = p_c = p, delta = 0). Both series hold the full-design threshold",
    "gamma* fixed; 'SSR + RAR disabled' is an ablation, not a separately",
    "calibrated design. Monte Carlo standard errors are conditional on gamma*."),
    align = rep("r", 7))

format_null_sweep(contemporaneous_null_trend) |>
  kable(caption = paste(
    "Rejection probability under contemporaneous no treatment effect",
    "(p_e(i) = p_c(i), delta = 0.05) with a time-constant analysis.",
    "Both series hold the full-design threshold gamma* fixed. Read the",
    "difference against its own MCSE; individual rows near zero are not",
    "distinguishable from no difference at N = 5,000 per cell."),
    align = rep("r", 7))
```

```{r composite-null-plot, fig.width=13, fig.height=5.2, fig.cap="Pointwise rejection probability across the common-rate grid under Rule B. The left panel is the constant-rate equality boundary of the superiority null and reports Type I error; the right panel is a shared-trend contemporaneous no-effect stress scenario analyzed with a time-constant model. Grey holds the full-design threshold gamma* with SSR and RAR disabled; green is the full design. Both are fixed-threshold comparisons, not separately calibrated designs. Dashed red line = nominal alpha = 0.025."}
null_plot_data <- bind_rows(
  composite_null_static |>
    mutate(DGM = "Constant rate (δ = 0): Type I error"),
  contemporaneous_null_trend |>
    mutate(DGM = "Shared trend (δ = 0.05): rejection probability"))

null_plot_data |>
  select(p, DGM, `SSR + RAR disabled`, `All mechanisms`) |>
  pivot_longer(c(`SSR + RAR disabled`, `All mechanisms`),
               names_to = "Design", values_to = "Rejection") |>
  ggplot(aes(p, Rejection, color = Design, shape = Design)) +
  geom_hline(yintercept = alpha, linetype = "dashed",
             color = ZET$red, linewidth = 0.8) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.8) +
  facet_wrap(~DGM, nrow = 1) +
  scale_color_manual(values = c(
    `SSR + RAR disabled` = ZET$neutralL,
    `All mechanisms` = ZET$green)) +
  scale_y_continuous(breaks = seq(0, 0.30, by = 0.05),
                     labels = percent_format(accuracy = 1)) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = bquote("Common baseline rate " * p),
       y = "Rejection probability", color = NULL, shape = NULL) +
  theme_zetyra() +
  # Type sizes are set above the theme defaults: this figure is two panels wide
  # in the manuscript, so everything in it renders at roughly half the linear
  # scale of a single-panel figure and the theme defaults come out small.
  theme(axis.title  = element_text(size = 17, face = "bold", color = ZET$ink,
                                   margin = margin(t = 4, r = 4)),
        axis.text   = element_text(size = 15, color = ZET$slate),
        strip.text  = element_text(size = 15, face = "bold", color = ZET$ink,
                                   margin = margin(t = 4, b = 4)),
        legend.text = element_text(size = 14, color = ZET$ink),
        legend.position = "bottom")
```

---

## 5.7 Sensitivity Analysis


All three sweeps pass the relevant parameter directly into `simulate_one_trial()`.
No global variables are mutated. These are **frozen-threshold perturbations**:
each setting retains the anchor-calibrated $\gamma^*$, rather than being
recalibrated as a new design. The 95% simulation intervals show Monte Carlo
uncertainty conditional on that threshold.

```{r sens-prior-weight}
#| cache: true
# Swept quantity is the weight on the FLAT robustify() component: larger
# w_vague = less borrowing from the historical prior.
w_grid <- seq(0, 0.90, by = 0.10)
set.seed(300L)

t1e_w <- sapply(w_grid, function(w) {
  t1e_at(replicate(N_sim_sens,
                 simulate_one_trial(mix    = set_vague_weight(prior_mix, w),
                                    p_true_e = p0_conflict,
                                    p_c_data = p0_conflict,
                                    delta    = delta_trend,
                                    use_ssr  = TRUE,
                                    use_rar  = TRUE)), gamma_star)
})

sens_w <- tibble(
  w = w_grid,
  t1e = t1e_w,
  se = mcse(t1e_w, N_sim_sens),
  lo = pmax(0, t1e - 1.96 * se),
  hi = pmin(1, t1e + 1.96 * se))

p_w <- sens_w |>
  ggplot(aes(w, t1e)) +
  geom_ribbon(aes(ymin = lo, ymax = hi),
              fill = ZET$green, alpha = 0.15, color = NA) +
  geom_line(color = ZET$green, linewidth = 1.2) +
  geom_point(color = ZET$green, size = 2.5) +
  geom_hline(yintercept = alpha, linetype = "dashed", color = ZET$red) +
  scale_x_continuous(labels = percent_format()) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(title = "Vague-Component Weight", x = bquote(w[vague]),
       y = "Rejection probability") +
  theme_zetyra()
```

```{r sens-time-trend}
#| cache: true
delta_grid <- seq(0, 0.15, by = 0.025)
set.seed(400L)

t1e_d <- sapply(delta_grid, function(d) {
  t1e_at(replicate(N_sim_sens,
                 simulate_one_trial(mix    = prior_mix,
                                    p_true_e = p0_conflict,
                                    p_c_data = p0_conflict,
                                    delta    = d,          # varied here
                                    use_ssr  = TRUE,
                                    use_rar  = TRUE)), gamma_star)
})

sens_d <- tibble(
  d = delta_grid,
  t1e = t1e_d,
  se = mcse(t1e_d, N_sim_sens),
  lo = pmax(0, t1e - 1.96 * se),
  hi = pmin(1, t1e + 1.96 * se))

p_d <- sens_d |>
  ggplot(aes(d, t1e)) +
  geom_ribbon(aes(ymin = lo, ymax = hi),
              fill = ZET$green, alpha = 0.15, color = NA) +
  geom_line(color = ZET$green, linewidth = 1.2) +
  geom_point(color = ZET$green, size = 2.5) +
  geom_hline(yintercept = alpha, linetype = "dashed", color = ZET$red) +
  scale_x_continuous(labels = percent_format()) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(title = "Time Trend Magnitude", x = bquote(δ),
       y = "Rejection probability") +
  theme_zetyra()
```

```{r sens-ssr-timing}
#| cache: true
ssr_grid <- c(20L, 30L, 40L, 45L, 55L, 60L)
set.seed(500L)

t1e_ssr <- sapply(ssr_grid, function(nt) {
  t1e_at(replicate(N_sim_sens,
                 simulate_one_trial(mix      = prior_mix,
                                    p_true_e   = p0_conflict,  # H0: p_e = p_c = p0_conflict
                                    p_c_data   = p0_conflict,
                                    delta      = delta_trend,
                                    n_ssr_trig = nt,       # varied here
                                    use_ssr    = TRUE,
                                    use_rar    = TRUE)), gamma_star)
})

p_ssr <- tibble(
  nt    = ssr_grid,
  t1e   = t1e_ssr,
  se    = mcse(t1e_ssr, N_sim_sens),
  lo    = pmax(0, t1e - 1.96 * se),
  hi    = pmin(1, t1e + 1.96 * se),
  label = sprintf("n=%d\n(%d%%)", ssr_grid,
                  as.integer(round(ssr_grid / N_base * 100)))
) |>
  ggplot(aes(label, t1e)) +
  geom_col(fill = ZET$green, width = 0.55, alpha = 0.90) +
  geom_errorbar(aes(ymin = lo, ymax = hi), width = 0.14,
                linewidth = 0.6, color = ZET$ink) +
  geom_hline(yintercept = alpha, linetype = "dashed", color = ZET$red) +
  geom_text(aes(y = hi, label = sprintf("%.4f", t1e)),
            vjust = -0.25, size = 3.5, color = ZET$ink) +
  scale_y_continuous(limits = c(0, max(pmin(1, t1e_ssr +
                                             1.96 * mcse(t1e_ssr, N_sim_sens))) * 1.18),
                     labels = percent_format(accuracy = 0.1)) +
  labs(title = "SSR Trigger Timing", x = bquote(n[SSR]),
       y = "Rejection probability") +
  theme_zetyra()
```

```{r sens-panel, fig.width=13, fig.height=4.5, fig.cap="Frozen-threshold sensitivity of rejection probability to prior weight (left), time trend (centre), and Rule B timing (right). Bands and bars are pointwise 95% Monte Carlo simulation intervals conditional on gamma*. Dashed line = nominal alpha = 0.025."}
p_w | p_d | p_ssr
```

```{r sens-table}
tribble(
  ~Parameter, ~Range, ~`Observed rejection range`, ~`Largest MCSE`,
  "Vague-component weight (w<sub>vague</sub>)", "0% – 90%",
    sprintf("%.4f – %.4f", min(t1e_w), max(t1e_w)),
    sprintf("%.4f", max(mcse(t1e_w, N_sim_sens))),
  "Time trend (δ)",  "0% – 15%",
    sprintf("%.4f – %.4f", min(t1e_d), max(t1e_d)),
    sprintf("%.4f", max(mcse(t1e_d, N_sim_sens))),
  "SSR timing (n<sub>SSR</sub>)",      "n = 20, 30, 40, 45, 55, 60",
    sprintf("%.4f – %.4f", min(t1e_ssr), max(t1e_ssr)),
    sprintf("%.4f", max(mcse(t1e_ssr, N_sim_sens)))
) |>
  kable(caption = paste(
    "Observed frozen-threshold sensitivity ranges across the evaluated grids.",
    "The ranges are descriptive extrema of noisy point estimates, not bounds"),
        align   = c("l","l","r","r"),
        escape  = FALSE)
```

---

## 5.8 Reproducibility

**Append-only grid extension, and what it demonstrated.** The common-rate grid
was extended from seven points to nine (adding $p = 0.40$ and $0.45$) after the
seven-point results had been reviewed. Both sweeps seed per grid *position*
(`seed_base + 1000*i`), so appending the new rates left $i = 1,\dots,7$
untouched, and the seven original rows were expected to reproduce exactly.

The development comparison found all 42 retained point estimates across both
sweeps — ablation, full design, and their difference at seven rates in two
sweeps — bit-identical before and after the extension; the MCSEs are
deterministic functions of the corresponding proportions. The pre-extension
render was overwritten and is not part of the release, so this comparison is
recorded rather than independently auditable from the released files. The
per-cell seeding nevertheless ensures that appended points cannot perturb the
random streams used by earlier positions.

Had a point been *inserted* rather than appended, every position after it would
have shifted, silently reseeding every subsequent cell and changing numbers
that had already been reviewed. Any future extension of this grid should
likewise append.

```{r auto-summary}
#| code-fold: false
cat(sprintf("gamma*                = %.3f\n", gamma_star))
cat(sprintf("Type I error (calibration)  = %.4f  (target: %.3f)\n",
            t1e_at_star, alpha))
cat(sprintf("Type I error (fresh anchor) = %.4f  (95%% SI [%.4f, %.4f])\n",
            anchor_validation_t1e,
            anchor_validation_ci["lower"], anchor_validation_ci["upper"]))
cat(sprintf("Rejection probability (+2 pp from anchor, p=0.27) = %.4f  (+%.1f%%)\n",
            t1e_mild, inflation_pct_mild))
cat(sprintf("Rejection probability (+5 pp from anchor, p=0.30) = %.4f  (+%.1f%%)\n",
            t1e_primary, inflation_pct))
cat(sprintf("Rejection probability (earlier SSR, n_ssr=30) = %.4f  (+%.1f%%)\n",
            t1e_neg, (t1e_neg - alpha)/alpha*100))
cat(sprintf("Max constant-rate null Type I error on grid = %.4f\n",
            max(composite_null_static$`All mechanisms`)))
cat(sprintf("Max trend-active rejection probability on grid = %.4f\n",
            max(contemporaneous_null_trend$`All mechanisms`)))
cat(sprintf("Earlier-vs-reference change = %+.4f (%+.1f%% of stress excess)\n",
            timing_change, timing_change_pct_excess))
cat(sprintf("\nFixed-threshold factorial ablation:\n"))
cat(sprintf("  Delta_SSR         = %+.4f\n", delta_f_ssr))
cat(sprintf("  Delta_RAR         = %+.4f\n", delta_f_rar))
cat(sprintf("  Delta_interaction (pooled) = %+.4f\n", pooled))
cat(sprintf("\nBaseline-rate departure x trend:\n"))
cat(sprintf("  Departure only    = +%.4f\n", t1e_conf_only  - t1e_neither))
cat(sprintf("  Trend only        = +%.4f\n", t1e_trend_only - t1e_neither))
cat(sprintf("  Both (trend stress) = +%.4f\n", t1e_both_mech - t1e_neither))
cat(sprintf("  Interaction term  = %+.4f\n", interaction_term))
```

```{r mcse-table}
mcse_tbl <- tribble(
  ~Estimate,                     ~Probability,   ~N_sim,
  "Calibration (in-sample at gamma*)", t1e_at_star, N_sim_calib,
  "Independent anchor validation", anchor_validation_t1e, N_sim_sens,
  "+2 pp from anchor + trend (p=0.27)", t1e_mild, N_sim_main,
  "+5 pp from anchor + trend (p=0.30)", t1e_primary, N_sim_main,
  "Trend stress: earlier SSR (n = 30)", t1e_neg, N_sim_main,
  "Factorial: baseline",         t1e_f_base,     N_sim_sens,
  "Factorial: SSR only",         t1e_f_ssr,      N_sim_sens,
  "Factorial: RAR only",         t1e_f_rar,      N_sim_sens,
  "Factorial: SSR + RAR",        t1e_f_both,     N_sim_sens,
  "Departure x trend: neither",  t1e_neither,    N_sim_sens,
  "Departure x trend: departure only", t1e_conf_only, N_sim_sens,
  "Departure x trend: trend only", t1e_trend_only, N_sim_sens,
  "Departure x trend: both",     t1e_both_mech,  N_sim_sens
) |>
  mutate(
    MCSE   = mcse(Probability, N_sim),
    CI_lo  = Probability - 1.96 * MCSE,
    CI_hi  = Probability + 1.96 * MCSE,
    Probability = sprintf("%.4f", Probability),
    MCSE   = sprintf("%.4f", MCSE),
    `95% SI` = sprintf("[%.4f, %.4f]", CI_lo, CI_hi),
    N_sim  = formatC(N_sim, format = "d", big.mark = ",")
  ) |>
  # The calibration row is in-sample: gamma* was chosen by bisection ON this
  # same vector to be the smallest threshold achieving <= alpha, so the value is
  # near-deterministic by construction (its across-seed SD is O(1/N_sim), not
  # the binomial sqrt(p(1-p)/N_sim)). Printing a binomial MCSE and a simulation
  # interval for it would describe an estimator that was not used. The honest
  # out-of-sample companion is the row directly below it.
  mutate(MCSE     = if_else(grepl("^Calibration", Estimate), "---", MCSE),
         `95% SI` = if_else(grepl("^Calibration", Estimate), "---", `95% SI`)) |>
  select(Estimate, Probability, MCSE, `95% SI`, N_sim)

kable(mcse_tbl,
      caption = paste("Monte Carlo standard errors and 95% simulation intervals",
                      "for key rejection-probability estimates, conditional on",
                      "the selected gamma*. No-trend rows estimate Type I error;",
                      "trend-active rows are contemporaneous no-effect stress",
                      "scenarios under a time-constant analysis. The calibration row is",
                      "in-sample and near-deterministic by construction, so no",
                      "MCSE is reported for it; use the independent anchor",
                      "validation row for the out-of-sample anchor value."),
      align = c("l", "r", "r", "c", "r"))

# ── Propagated MCSEs for derived interaction terms ──────────────────────────
# Interaction = T1E_both - T1E_A - T1E_B + T1E_neither
# All four simulations are independent, so Var(interaction) = sum of variances

# Departure x trend interaction
se_mech <- sqrt(
  mcse(t1e_both_mech,  N_sim_sens)^2 +
  mcse(t1e_conf_only,  N_sim_sens)^2 +
  mcse(t1e_trend_only, N_sim_sens)^2 +
  mcse(t1e_neither,    N_sim_sens)^2
)

int_tbl <- tribble(
  ~`Interaction term`,                    ~Estimate,        ~MCSE,    ~`95% SI`,
  "Departure × trend", interaction_term, se_mech,
    sprintf("[%+.4f, %+.4f]", interaction_term - 1.96*se_mech, interaction_term + 1.96*se_mech),
  "Factorial (SSR × RAR; pooled)", pooled, se_p,
    sprintf("[%+.4f, %+.4f]", pooled - 1.96*se_p, pooled + 1.96*se_p)
) |>
  mutate(Estimate = sprintf("%+.4f", Estimate),
         MCSE    = sprintf("%.4f", MCSE))

kable(int_tbl,
      caption = paste("Propagated MCSEs for derived interaction terms.",
                      "The factorial row is the inverse-variance pooled estimate",
                      "from the N = 5,000 and N = 20,000 per-cell runs reported above.",
                      "Intervals are Monte Carlo simulation intervals, not",
                      "sampling-confidence intervals for clinical data."),
      align = c("l", "r", "r", "c"))
```

**Seeds:**

| Simulation | Seed |
|---|---|
| MAP prior MCMC | 42 (via `set.seed()` before `gMAP()`) |
| RAR validation | 42 |
| Threshold calibration | 1 |
| Calibration-threshold bootstrap | 2 |
| Independent anchor validation | 101 |
| Trend stress: +5 pp from anchor (p=0.30) | 2026 |
| Trend stress: +2 pp from anchor (p=0.27) | 2028 |
| Earlier-SSR timing sensitivity (n<sub>SSR</sub> = 30) | 2027 |
| Fixed-threshold factorial ablation | per cell: 300000 + 1000·k, k = 1..4 |
| Departure × trend (2x2) | per cell: 400000 + 1000·k, k = 1..4 |
| Fresh-seed interaction check | per cell: 800000 + 1000·k, k = 1..4 |
| Shared-trend grid — ablation series | per grid point: 600000 + 1000·i, i = 1..9 |
| Shared-trend grid — full design | per grid point: 700000 + 1000·i, i = 1..9 |
| Constant-rate null grid — ablation series | per grid point: 900000 + 1000·i, i = 1..9 |
| Constant-rate null grid — full design | per grid point: 1000000 + 1000·i, i = 1..9 |
| Sensitivity — prior weight | 300 |
| Sensitivity — time trend | 400 |
| Sensitivity — SSR timing | 500 |

```{r session-info}
#| code-fold: false
# Also written to disk. Printing into the rendered HTML records the producing
# environment nowhere a reader can reach -- the HTML is gitignored, and the MAP
# prior is fitted through rstan, so the RBesT and Stan versions are part of the
# result. SESSION_INFO.txt sits beside this source and is committed.
session_record <- sub("[[:space:]]+$", "", capture.output(sessionInfo()))
writeLines(session_record, "SESSION_INFO.txt")
sessionInfo()
```

```{r map-prior-release-restate}
#| cache: false
# Restates the preflight outcome beside sessionInfo() so the release record
# carries it, and re-reads the committed file to confirm the object that
# actually fed the simulations is still that file. That last identity IS a hard
# stop wherever it runs: if it fails, the results in this document were not
# produced by the committed prior, which is a correctness failure rather than
# an environment difference.
map_release_check <- unname(dget(MAP_FROZEN_PATH))
dimnames(map_release_check) <- list(c("w", "a", "b"), NULL)
stopifnot(identical(prior_mix, map_release_check))
cat(sprintf("Simulation input: %s (confirmed unchanged)\n", MAP_FROZEN_PATH))
cat("Layer 1, published 4 dp prior table: PASS (hard gate)\n")
cat(sprintf("Layer 2, refit reproduces the frozen mixture: %s  (max dev %.3e, bound %.0e)\n",
            if (map_frozen_ok) "yes" else "NO -- see the preflight warning",
            map_frozen_dev, MAP_FROZEN_TOL))
cat(sprintf("Release-certifiable render: %s\n",
            if (map_frozen_ok) "yes" else
              "NO -- reproduction is valid, certification is not"))
cat("Status written to map_prior_status.txt; asserted by ",
    "tools/release/check_release.py --only map_prior\n", sep = "")
```

---

## References

::: {#refs}
:::
