---
title: "Section 5: Simulation Study"
subtitle: "Composed Pipeline Operating Characteristics in Bayesian Clinical Trial Design"
author: "Lu Qian · Zetyra | Evidence in the Wild"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-float: true
    code-fold: true
    code-tools: true
    theme:
      - cosmo
      - custom.scss
    fig-width: 9
    fig-height: 5.5
    df-print: kable
    embed-resources: true
  pdf:
    toc: true
    number-sections: true
    fig-width: 7
    fig-height: 4.5
    geometry: margin=1in
    fontsize: 11pt
execute:
  echo: true
  warning: false
  message: false
  cache: true
  cache-path: "_cache/"
bibliography: references.bib
---

```{r setup}
#| cache: false
library(RBesT)
library(ggplot2)
library(dplyr)
library(tibble)
library(purrr)
library(tidyr)
library(knitr)
library(patchwork)
library(scales)

# ── Color palette ─────────────────────────────────────────────────────────────
# Ink/green retained as primary accent; backgrounds switched to clean white/slate
ZET <- list(
  ink      = "#1A1A2E",   # deep navy-black — titles, axis text
  green    = "#1B3A2E",   # forest green  — primary series, headers
  green2   = "#2D6A4F",   # mid green     — secondary series
  teal     = "#40916C",   # light green   — tertiary / Version B
  slate    = "#4B5563",   # dark slate    — axis text, subtitles, legends
  gridline = "#E5E7EB",   # cool light gray — major gridlines
  red      = "#B91C1C",   # accessible red — reference lines, FAIL
  amber    = "#D97706",   # amber         — warning / border
  bg       = "#FFFFFF",   # pure white    — plot background
  panel    = "#FFFFFF",   # white         — panel background
  muted    = "#545C6B",   # dark grey     — captions, reference annotations (6.7:1)
  neutralL = "#46505F",   # darker grey   — grey DATA series (8.0:1 on white)
  neutralF = "#8B95A3"    # mid grey fill — baseline/neutral BARS (3.0:1, WCAG 1.4.11)
)

# ── Cache invalidation on palette change ──────────────────────────────────────
# knitr keys a chunk's cache on that chunk's own code text. The palette lives
# here, in an uncached setup chunk, so editing a colour does NOT invalidate the
# cached plotting chunks — a palette-only edit would silently leave most figures
# at the OLD colours while the few uncached ones update, producing a deck with
# mismatched greys. Making the palette part of every chunk's cache key fixes
# that: change a colour and every figure regenerates; leave it alone and the
# expensive simulations stay cached.
knitr::opts_chunk$set(cache.extra = list(zet_palette = ZET))

# ── Modern publication theme ──────────────────────────────────────────────────
theme_zetyra <- function(base_size = 12) {
  theme_minimal(base_size = base_size, base_family = "sans") +
    theme(
      # Backgrounds — clean white, no colored fills
      plot.background   = element_rect(fill = ZET$bg,    color = NA),
      panel.background  = element_rect(fill = ZET$panel, color = NA),

      # Grid — thin horizontal only; vertical removed for clarity
      panel.grid.major.x = element_blank(),
      panel.grid.major.y = element_line(color = ZET$gridline, linewidth = 0.45),
      panel.grid.minor   = element_blank(),

      # Axes — subtle bottom/left rule; no full box
      axis.line.x  = element_line(color = "#D1D5DB", linewidth = 0.5),
      axis.ticks.x = element_line(color = "#D1D5DB", linewidth = 0.4),
      axis.ticks.length = unit(3, "pt"),
      axis.text    = element_text(color = ZET$slate,  size = base_size - 1),
      axis.title   = element_text(color = ZET$ink,    size = base_size,
                                   face = "bold", margin = margin(t = 4, r = 4)),

      # Titles — left-aligned, strong hierarchy
      plot.title    = element_text(color = ZET$ink,   size = base_size + 3,
                                   face = "bold",     hjust = 0,
                                   margin = margin(b = 4)),
      plot.subtitle = element_text(color = ZET$slate, size = base_size - 1,
                                   hjust = 0,         margin = margin(b = 10)),
      plot.caption  = element_text(color = ZET$muted, size = base_size - 2,
                                   hjust = 0,         margin = margin(t = 8)),
      plot.margin   = margin(12, 16, 10, 12),

      # Legend — top-right, no box
      legend.position    = "top",
      legend.justification = "left",
      legend.background  = element_rect(fill = NA, color = NA),
      legend.key         = element_rect(fill = NA, color = NA),
      legend.text        = element_text(color = ZET$slate, size = base_size - 1),
      legend.title       = element_text(color = ZET$ink,   size = base_size - 1,
                                         face = "bold"),
      legend.spacing.x   = unit(6, "pt"),

      # Facet strips — light gray pill-style
      strip.background = element_rect(fill = ZET$gridline, color = NA,
                                       linewidth = 0),
      strip.text       = element_text(color = ZET$ink, face = "bold",
                                       size = base_size - 1, margin = margin(4,6,4,6))
    )
}
```

---

## Overview

This document implements the simulation study for a two-arm Phase II oncology
design with historical borrowing on the control arm, Bayesian posterior
monitoring, blinded sample size re-estimation (SSR) and response-adaptive
randomization (RAR).

**Primary claim — robustness.** An efficacy threshold calibrated at a single
assumed control rate does not deliver Type I error control across the composite
null. Ablations that switch off the adaptive components while *holding that
threshold fixed* do not restore control either.

**Secondary claim — composition.** SSR and RAR modify that excess substantially
and nonlinearly, and the sign of their combined effect depends on where in the
null the design is evaluated (protective at low rates, harmful at high ones).

**Scope.** The grid *demonstrates* a control failure and establishes that the
supremum is at least 0.2618; it does not locate the supremum. Scenarios are
departures of the baseline rate from the calibration anchor $p_0 = 0.25$, not
quantified prior–data conflict. On the plausibility of departures of this size,
@tang_2010 treat a five-percentage-point control-rate drift as clinically modest
in an oncology simulation, and @viele_2014 review the conditions under which
historical and concurrent control rates diverge. Allocation targets follow the
optimal-allocation literature [@rosenberger_optimal; @rosenberger_lachin].

The platform and its benchmark framework are documented in the Zetyra white
paper [@zetyra_whitepaper]; the Bayesian calculators used here are documented
separately [@zetyra_bayesian_toolkit].

Simulation code:
`github.com/evidenceinthewild/Composed-Adaptive-Trial-Validation`

---

## 5.1 Design Parameters

```{r design-params}
# ── Trial parameters ──────────────────────────────────────────────────────────
p0       <- 0.25   # null / historical control response rate
p_alt    <- 0.45   # alternative response rate
alpha    <- 0.025  # one-sided Type I error target

# ── Sample size & looks ───────────────────────────────────────────────────────
N_base   <- 90L
N_max    <- 150L
n_looks  <- c(30L, 45L, 60L, 90L)
n_ssr    <- 45L    # SSR trigger at 50% of N_base

# ── MAP prior ────────────────────────────────────────────────────────────────
# The prior is FITTED from historical data in 5.2.1 and stored in `prior_mix`,
# a matrix with rows (w, a, b), one column per mixture component.
#
# NOTE (2026-07 correction): earlier versions declared scalar defaults
#   a_inf = 6.5; b_inf = 19.5; w_inf = 0.70; a_vag = 0.5; b_vag = 0.5
# both here AND as default arguments of simulate_one_trial(). Because R
# evaluates default arguments as literals in the function frame, the fitted
# values assigned here never reached the simulation: every result was produced
# by the hand-specified 0.70*Beta(6.5,19.5) + 0.30*Beta(0.5,0.5) mixture
# (mean 0.325), not by the RBesT MAP fit. The prior is now passed explicitly
# as a single object so the two cannot diverge.
prior_mix <- NULL   # assigned in 5.2.1; simulations require it

# ── Departures from the calibration anchor & time trend ──────────────────────
# NAMING NOTE: the variables below are still called `p0_conflict` for historical
# reasons. They are NOT quantified prior-data conflict, which would have to be
# defined against the prior predictive distribution. They are departures of the
# common baseline rate from the calibration anchor p0 = 0.25. The prose and all
# reported labels use the accurate term; only the identifiers are legacy.
#
# Departure scenarios are defined relative to the DESIGN NULL p0 = 0.25 -- the
# historical control rate the efficacy threshold is calibrated at -- NOT
# relative to the mean of the robustified MAP mixture.
#
# NOTE (2026-07 correction): these were labelled "2pp / 5pp above the prior
# mean". That is false for every candidate prior: the fitted robustified
# mixture has mean 0.3132, so p = 0.27 and p = 0.30 both lie BELOW it. The
# reference point is now named explicitly wherever these are described.
# For orientation, the relevant means are:
#   design null / historical rate       0.2500
#   dominant (informative) component    0.2568
#   full robustified mixture            0.3132
p0_conflict      <- 0.30   # +5 pp above the design null p0 = 0.25 (stress test)
p0_conflict_mild <- 0.27   # +2 pp above the design null p0 = 0.25 (headline)
delta_trend <- 0.05   # linear drift over full enrollment

# ── Monte Carlo & simulation counts ──────────────────────────────────────────
# (N_mc removed: posterior probabilities are now computed by deterministic
#  quadrature rather than Monte Carlo -- see map_posterior_prob().)
N_sim_calib  <- 20000L
N_sim_main   <- 10000L
N_sim_sens   <-  5000L

# ── Monte Carlo SE helper ───────────────────────────────────────────────────
# Binomial MCSE for a proportion estimated from N_sim replications
mcse <- function(p_hat, n_sim) sqrt(p_hat * (1 - p_hat) / n_sim)

# 95% simulation interval
mcse_ci <- function(p_hat, n_sim) {
  se <- mcse(p_hat, n_sim)
  c(lower = p_hat - 1.96 * se, upper = p_hat + 1.96 * se)
}
```

---

## 5.2 Component-Level Validation

### 5.2.1 MAP Prior — Benchmarked Against RBesT [@weber_rbest]

```{r map-prior-fit}
hist_data <- data.frame(
  study = c("H1", "H2", "H3"),
  n     = c( 50,   40,   60),
  r     = c( 13,   10,   15)
)

set.seed(42)
map_fit <- gMAP(
  cbind(r, n - r) ~ 1 | study,
  data       = hist_data,
  family     = binomial,
  tau.dist   = "HalfNormal",
  tau.prior  = 1,
  beta.prior = 2,
  chains     = 4,
  iter       = 4000,
  warmup     = 1000
)

map_mix <- automixfit(map_fit, Nc = 2)
rob_mix <- robustify(map_mix, weight = 0.20, mean = 0.5)
print(rob_mix)
```

```{r map-prior-validation, fig.cap="MAP prior density: components and mixture from the fitted rob_mix object. Analytic density reproduces dmix() to floating-point precision."}

p_seq <- seq(0.01, 0.99, length.out = 500)

# Parameters from fitted rob_mix (matrix with rows "w","a","b"; cols = components)
w_fit   <- rob_mix["w", ]   # mixture weights
a_fit   <- rob_mix["a", ]   # Beta alpha parameters
b_fit   <- rob_mix["b", ]   # Beta beta  parameters

cat("Fitted mixture components:\n")
print(rbind(w = round(w_fit, 4), a = round(a_fit, 3), b = round(b_fit, 3)))

# Analytic density using the EXTRACTED parameters
rbest_dens    <- dmix(rob_mix, p_seq)          # RBesT reference
analytic_dens <- sapply(p_seq, function(p)
  sum(w_fit * dbeta(p, a_fit, b_fit)))         # our formula with fitted params

max_dev_map <- max(abs(rbest_dens - analytic_dens))

# ── Carry the FULL fitted mixture into the simulation ────────────────────────
# automixfit(Nc = 2) + robustify() returns THREE components. Earlier versions
# assumed two: they took the most informative component, treated the *first*
# leftover as "vague", and dropped the Beta(1,1) robustify() component
# entirely. The full mixture is now passed through intact.
prior_mix <- rbind(w = w_fit, a = a_fit, b = b_fit)

cat(sprintf("\nPrior carried into simulation (%d components):\n", ncol(prior_mix)))
print(round(prior_mix, 4))
cat(sprintf("  mixture mean = %.4f\n", sum(w_fit * a_fit / (a_fit + b_fit))))
stopifnot(abs(sum(w_fit) - 1) < 1e-8)

# Component densities for plot
df_comps <- tibble(
  p     = rep(p_seq, length(w_fit) + 1),
  dens  = c(sapply(seq_along(w_fit), function(j)
               w_fit[j] * dbeta(p_seq, a_fit[j], b_fit[j])),
             rbest_dens),
  label = rep(c(sprintf("Component %d (w=%.2f)", seq_along(w_fit), w_fit),
                "Mixture (RBesT)"), each = 500)
)

comp_colors <- c("#2166AC", "#4393C3", "#D6604D", "#B2182B")
comp_labels <- levels(factor(df_comps$label))
names(comp_colors) <- comp_labels[seq_along(comp_colors)]

ggplot(df_comps, aes(p, dens, color = label, linetype = label)) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = comp_colors) +
  labs(title    = "MAP Prior: Fitted Mixture Components and Density",
       subtitle = sprintf(
         "Max |analytic - dmix| = %.2e  (threshold: 1e-10, floating-point only)",
         max_dev_map),
       x = bquote("Control response rate" ~ p[c]), y = "Density",
       color = NULL, linetype = NULL) +
  theme_zetyra() + theme(legend.position = "bottom")

cat(sprintf("\nMAP PRIOR:  max dev = %.2e  [floating-point precision]  %s\n",
            max_dev_map, ifelse(max_dev_map < 1e-8, "PASS", "FAIL")))
```

### 5.2.2 Posterior Probability Helper

The monitoring rule is a posterior-probability threshold applied at each
interim look; see @zhou_ji_2024 for the design family and the question of
whether such rules require adjustment for interim analyses.

```{r posterior-prob-fn}
# ── Conjugate update of an arbitrary Beta-mixture prior ──────────────────────
# mix: matrix with rows (w, a, b), one column per component.
mixture_posterior <- function(mix, x_c, n_c) {
  w <- mix["w", ]; a <- mix["a", ]; b <- mix["b", ]
  a_post <- a + x_c
  b_post <- b + (n_c - x_c)
  lml <- lbeta(a_post, b_post) - lbeta(a, b)      # log marginal likelihood
  lw  <- log(w) + lml
  lw  <- lw - max(lw)                              # numerically stable
  w_post <- exp(lw); w_post <- w_post / sum(w_post)
  rbind(w = w_post, a = a_post, b = b_post)
}

# ── Gauss-Legendre nodes on [0,1] for deterministic integration ──────────────
# Golub-Welsch: nodes/weights are the eigenvalues/first-components of the
# symmetric tridiagonal Jacobi matrix. Base R only -- no extra dependency.
gauss_legendre_01 <- function(n) {
  k <- 1:(n - 1)
  b <- k / sqrt(4 * k^2 - 1)
  J <- diag(0, n)
  J[cbind(k, k + 1)] <- b
  J[cbind(k + 1, k)] <- b
  e <- eigen(J, symmetric = TRUE)
  ord <- order(e$values)
  x <- e$values[ord]                  # nodes on [-1, 1]
  w <- 2 * (e$vectors[1, ord])^2      # weights on [-1, 1]
  list(x = 0.5 * (x + 1), w = 0.5 * w)   # mapped to [0, 1]
}
.gl <- gauss_legendre_01(400)
.QX <- .gl$x
.QW <- .gl$w

# Sanity checks on the rule actually used below.
#
# Normalisation alone is too weak to certify a quadrature rule: a set of
# positive weights summing to 1 passes it whether or not the nodes are right.
# A standalone diagnostic written for this project built the Jacobi matrix with
# the off-diagonal terms on the MAIN diagonal, still integrated the constant 1
# to 1.0, still produced nodes inside (0, 1) -- and integrated x to 0.7887
# instead of 0.5. Moment exactness to degree 2n-1 is the property that actually
# distinguishes a Gauss-Legendre rule, so it is checked here, at render time,
# on the rule the simulation uses.
.g8 <- gauss_legendre_01(8L)
stopifnot(
  abs(sum(.QW) - 1)                     < 1e-12,   # int 1    = 1
  abs(sum(.QW * .QX) - 0.5)             < 1e-12,   # int x    = 1/2
  abs(sum(.QW * .QX^2) - 1 / 3)         < 1e-12,   # int x^2  = 1/3
  abs(sum(.g8$w * .g8$x^15) - 1 / 16)   < 1e-10,   # exact to degree 2n-1
  abs(sum(.QW * dbeta(.QX, 3, 7)) - 1)  < 1e-10)   # a smooth Beta density
rm(.g8)

# ── P(p_e > p_c | data), by deterministic quadrature ─────────────────────────
# = int f_e(x) F_c(x) dx, where p_e ~ Beta(.5+x_e, .5+n_e-x_e) (Jeffreys, no
# borrowing on the experimental arm) and p_c ~ posterior mixture.
#
# NOTE (2026-07 correction): this was previously a 2,000-draw Monte Carlo
# estimate. Because the monitoring rule takes a maximum over interim looks,
# Monte Carlo noise in the estimate inflated that maximum, and identical trial
# data could yield different decisions. Deterministic 400-node Gauss-Legendre
# quadrature removes that randomness.
#
# It is DETERMINISTIC, not exact. The sanity checks above verify the
# Gauss-Legendre weights and one smooth Beta integral; they do not bound the
# approximation error for every state the adaptive design can visit. Extreme
# all-success or all-failure states can concentrate posterior mass near a
# boundary and are the most demanding cases for a fixed-node rule.
#
# All reported simulations use the specified 400-node rule. This supplement
# does not claim machine-precision agreement with a Monte Carlo reference:
# a 4e5-draw reference has Monte Carlo error of order 1e-3. Nor does it claim
# invariance of gamma* or the T1E cells to the node count; establishing that
# would require a paired rerun of calibration and each reported experiment
# under both the 400-node rule and a higher-order reference.
map_posterior_prob <- function(x_e, n_e, x_c, n_c, mix) {
  post <- mixture_posterior(mix, x_c, n_c)
  f_e <- dbeta(.QX, 0.5 + x_e, 0.5 + (n_e - x_e))
  F_c <- rowSums(vapply(seq_len(ncol(post)),
                        function(j) post["w", j] *
                          pbeta(.QX, post["a", j], post["b", j]),
                        numeric(length(.QX))))
  sum(.QW * f_e * F_c)
}
```

### 5.2.3 Single-Arm SSR — Conditional-Power Self-Consistency

```{r ssr-helpers}
# Conditional power: Wald approximation for binary endpoint (blinded interim).
conditional_power <- function(x_int, n_int, n_total, p0, alpha = 0.025) {
  p_hat <- x_int / n_int
  if (n_total <= n_int || p_hat <= p0) return(0)
  z_a <- qnorm(1 - alpha)
  mu  <- (p_hat - p0) / sqrt(p0 * (1 - p0) / n_total)
  pmax(0, pnorm(mu - z_a))
}

# SSR: find smallest n >= n_int achieving target CP, capped at n_max.
ssr_n_target <- function(x_int, n_int, p0,
                          target_cp = 0.80, alpha = 0.025, n_max = N_max) {
  p_hat <- x_int / n_int
  if (p_hat <= p0) return(n_int)
  lo <- n_int;  hi <- n_max
  for (k in seq_len(50)) {
    mid <- (lo + hi) %/% 2L
    if (conditional_power(x_int, n_int, mid, p0, alpha) >= target_cp)
      hi <- mid else lo <- mid
    if (hi - lo <= 1L) break
  }
  min(hi, n_max)
}
```

```{r ssr-validation}
# ── Self-consistency validation for ssr_n_target() ────────────────────────────
#
# Approach: for each test scenario (x_int successes at n_int patients),
# (1) call ssr_n_target() to get the adjusted sample size n*,
# (2) verify that conditional_power(n*) is within 2% of the target CP.
#
# This confirms the SSR function correctly solves the inversion problem:
# "find the smallest n such that CP(n) ≥ target_cp."
#
# intentionally returns n_int unchanged (no increase warranted). In these cases
# conditional_power(n_int, n_int, ...) = 0 by construction, giving deviation = 0.80.
# These are correct behavior, not failures. Exclude them from max_dev calculation.
# The pass column flags them separately as "futile — correct".

ssr_test <- tribble(
  ~x_int, ~n_int, ~target_cp,
  # Futile zone: p_hat ≤ p0 → ssr_n_target returns n_int unchanged (no SSR)
  11, 45, 0.80,    # p_hat = 0.244 < p0 = 0.25 → futile (correct: no increase)
  # Active zone: p_hat > p0 → SSR should find n* achieving target CP
  14, 45, 0.80,    # p_hat = 0.311 → SSR should increase n
  16, 45, 0.80,    # p_hat = 0.356
  18, 45, 0.80,    # p_hat = 0.400
  20, 45, 0.80,    # p_hat = 0.444
  14, 30, 0.80,    # earlier interim, same effect size
  18, 60, 0.80     # later interim
)

ssr_res <- ssr_test |>
  mutate(
    p_hat  = x_int / n_int,
    futile = p_hat <= p0,
    n_star = pmap_int(list(x_int, n_int, target_cp), function(x, n, tcp) {
      ssr_n_target(x, n, p0, target_cp = tcp, n_max = N_max)
    }),
    cp_at_nstar = pmap_dbl(list(x_int, n_int, n_star), function(x, n, ns) {
      conditional_power(x, n, ns, p0, alpha)
    }),
    # CP at n*-1: confirms n* is the MINIMUM solution.
    # Because n is an integer, CP(n*) can legitimately overshoot the target
    # by a significant margin — measuring |CP(n*) - target| incorrectly
    # penalizes correct discrete optimization. The right test is minimality:
    #   (1) CP(n*)   >= target  — target is achieved
    #   (2) CP(n*-1) <  target  — n* cannot be reduced by 1
    cp_at_nstar_m1 = pmap_dbl(
      list(x_int, n_int, n_star, futile),
      function(x, n, ns, fut) {
        if (fut || ns <= n) return(NA_real_)
        conditional_power(x, n, ns - 1L, p0, alpha)
      }
    ),
    achieves_target = cp_at_nstar >= target_cp | n_star == N_max,
    is_minimal      = is.na(cp_at_nstar_m1) | cp_at_nstar_m1 < target_cp,
    pass = case_when(
      futile      ~ n_star == n_int,               # futile: no increase is correct
      n_star == N_max ~ TRUE,                      # capped: acceptable
      TRUE        ~ achieves_target & is_minimal   # active: minimality check
    )
  )

kable(
  ssr_res |>
    select(x_int, n_int, target_cp, p_hat, n_star,
           cp_at_nstar, cp_at_nstar_m1, pass) |>
    mutate(across(where(is.numeric), ~ round(.x, 3))),
  caption = paste(
    "SSR minimality check: n* achieves target CP and cannot be reduced by 1.",
    "Futile (p̂ ≤ p₀): correct if n* = n<sub>int</sub>.",
    "Active: CP(n*) >= target AND CP(n*-1) < target.",
    "Note: CP(n*) may legitimately overshoot target — discrete n is expected."),
  col.names = c("x<sub>int</sub>","n<sub>int</sub>","Target CP","p̂","n*",
                "CP(n*)","CP(n*−1)","Pass"),
  escape = FALSE)

active_noncap <- !ssr_res$futile & ssr_res$n_star < N_max

cat(sprintf("SSR:  CP(n*) >= target for all active cases:   %s\n",
            ifelse(all(ssr_res$achieves_target[active_noncap]), "PASS", "FAIL")))
cat(sprintf("      CP(n*-1) < target (n* is minimal):       %s\n",
            ifelse(all(ssr_res$is_minimal[active_noncap], na.rm=TRUE), "PASS", "FAIL")))
cat(sprintf("      Futile: n_star == n_int for p_hat <= p0:  %s\n",
            ifelse(all(ssr_res$n_star[ssr_res$futile] ==
                       ssr_res$n_int[ssr_res$futile]), "PASS", "FAIL")))
```

### 5.2.4 RAR — Benchmarked Against Analytic Allocation [@rosenberger_optimal]

```{r rar-validation}
#| cache: true
thompson_alloc <- function(n, p_e, p_c, a0 = 0.5, b0 = 0.5, n_burn = 10L) {
  ae <- a0; be <- b0; ac <- a0; bc <- b0; n_e <- 0L
  for (i in seq_len(n)) {
    arm <- if (i <= n_burn) (i - 1L) %% 2L else
           as.integer(rbeta(1, ae, be) >= rbeta(1, ac, bc))
    if (arm == 1L) {
      y <- rbinom(1, 1, p_e); ae <- ae + y; be <- be + (1 - y); n_e <- n_e + 1L
    } else {
      y <- rbinom(1, 1, p_c); ac <- ac + y; bc <- bc + (1 - y)
    }
  }
  n_e / n
}


# P(X > Y) where X ~ Beta(a1,b1), Y ~ Beta(a2,b2) via numerical integration
prob_beta_gt <- function(a1, b1, a2, b2) {
  integrate(function(x) dbeta(x, a1, b1) * pbeta(x, a2, b2),
            lower = 0, upper = 1, rel.tol = 1e-8)$value
}

# Validate: after observing fixed outcomes, does one-step TS allocation
# proportion match P(Beta(ae,be) > Beta(ac,bc))?
# We do n_reps single-step allocations from a known posterior state.

single_step_ts <- function(ae, be, ac, bc, n_reps = 100000L) {
  mean(rbeta(n_reps, ae, be) >= rbeta(n_reps, ac, bc))
}

set.seed(42)
# Test scenarios: known posterior states after observing specific data
# ae = 0.5 + x_e, be = 0.5 + (n_e - x_e); similarly for control
rar_scen <- tribble(
  ~x_e, ~n_e, ~x_c, ~n_c,
    5,   10,    2,   10,   # clearly favors experimental
    3,   10,    5,   10,   # clearly favors control
    5,   10,    5,   10,   # symmetric (should give ~0.5)
    8,   20,    4,   20,   # moderate asymmetry
   12,   30,    8,   30    # larger sample
)

rar_res <- rar_scen |>
  mutate(
    ae = 0.5 + x_e, be = 0.5 + (n_e - x_e),
    ac = 0.5 + x_c, bc = 0.5 + (n_c - x_c),
    # Analytical reference: P(Beta(ae,be) > Beta(ac,bc))
    ref_alloc = pmap_dbl(list(ae, be, ac, bc),
                         ~ prob_beta_gt(..1, ..2, ..3, ..4)),
    # Empirical: 100k single-step TS draws
    emp_alloc = pmap_dbl(list(ae, be, ac, bc),
                         ~ single_step_ts(..1, ..2, ..3, ..4)),
    deviation = abs(emp_alloc - ref_alloc),
    pass      = deviation < 0.005   # Monte Carlo error at n=100k is ~0.001
  )

kable(
  rar_res |> select(x_e, n_e, x_c, n_c, ref_alloc, emp_alloc, deviation, pass),
  digits = 4,
  caption = paste(
    "RAR validation: Thompson Sampling one-step allocation probability vs.",
    "analytical P(Beta(ae,be) > Beta(ac,bc)) via numerical integration.",
    "n=100k MC draws; expected Monte Carlo SE ≈ 0.001."),
  col.names = c("x<sub>e</sub>","n<sub>e</sub>","x<sub>c</sub>","n<sub>c</sub>","Analytical P(θ<sub>e</sub>>θ<sub>c</sub>)",
                "TS proportion","|dev|","Pass"),
  escape = FALSE)

cat(sprintf("RAR:  max dev = %.5f  [threshold 0.005]  %s\n",
            max(rar_res$deviation), ifelse(all(rar_res$pass), "PASS", "FAIL")))
```

---

## 5.3 Composed Pipeline Simulation

### Core simulation function

```{r sim-fn}
# simulate_one_trial() -----------------------------------------------------------
#
# Returns the maximum posterior probability across interim looks; see the note
# below. Convert to a rejection decision with t1e_at(pp, gamma).
#
# Two-arm Phase II design with historical borrowing on the control arm via a
# MAP prior, three co-occurring adaptive mechanisms:
#   MAP Bayesian monitoring, blinded SSR (if use_ssr), Thompson Sampling RAR (if use_rar)
#
# Version B (SSR-timing sensitivity, NOT causal isolation -- moving SSR also
#   changes the realised final-N distribution):
#   Same monitoring schedule as primary (looks = c(30,45,60,90)).
#   Same use_ssr=TRUE, use_rar=TRUE.
#   ONLY difference: n_ssr_trig=30 instead of 45.
#   => SSR co-occurs with the n=30 look (not n=45), breaking co-occurrence
#      at the richer n=45 interim where conflict and trend have accumulated.
#
# Key property of Version B: monitoring boundaries evaluated at identical
# sample sizes in primary and control; only SSR timing differs.
#
# Factorial decomposition: use_ssr x use_rar gives four conditions.
# Factor crossing: vary p_c_data (baseline-rate departure) and delta (trend).

# Returns the MAXIMUM posterior probability across interim looks.
#
# Data generation does not depend on gamma -- the enrollment loop completes
# before monitoring begins -- so rejection at threshold gamma is exactly
# {max_pp > gamma}. Returning max_pp lets a single simulated sample yield the
# Type I error at EVERY threshold, which makes exact calibration cheap.
#
# NOTE (2026-07 correction): `mix` has no default. It previously did
# (w_inf = 0.70, a_inf = 6.5, b_inf = 19.5), and because every caller omitted
# the prior arguments, R used those literals and the fitted MAP prior was
# never actually used. Making the argument mandatory makes that failure
# impossible.
simulate_one_trial <- function(
    mix,                  # REQUIRED: fitted prior, rows (w, a, b)
    p_true_e   = 0.25,
    p_c_data   = 0.25,    # control data-generating rate (conflict: != p0)
    delta      = 0,       # time trend magnitude
    n_ssr_trig = 45L,     # when SSR fires (primary=45, Version B control=30)
    use_ssr    = TRUE,    # FALSE = no sample size adaptation
    use_rar    = TRUE,    # FALSE = fixed 1:1 alternating allocation
    n_base     = 90L,
    n_cap      = 150L,
    looks      = c(30L, 45L, 60L, 90L),
    n_burn_rar = 10L) {

  stopifnot(!missing(mix), is.matrix(mix), all(c("w","a","b") %in% rownames(mix)))

  n_total  <- n_base
  ssr_done <- FALSE

  alloc_v <- integer(n_cap)
  resp_v  <- numeric(n_cap)

  ae <- 0.5; be <- 0.5   # TS posteriors (experimental)
  ac <- 0.5; bc <- 0.5   # TS posteriors (control)

  i <- 1L
  while (i <= n_total) {

    # Calendar drift on a FIXED per-patient scale.
    #
    # NOTE (2026-07 correction): this was (i - 1L) / n_total, and n_total is
    # mutated when SSR expands the trial. Normalized time therefore jumped
    # BACKWARD at the moment SSR fired (e.g. 44/90 = 0.489 -> 45/150 = 0.300),
    # so the response probability fell. The artifact was induced by SSR, which
    # contaminated the estimated SSR effect and the conflict x trend
    # interaction. The denominator is now fixed at the planned horizon: drift
    # accrues at a constant rate per patient and is invariant to adaptation.
    # Note t_frac may exceed 1 when SSR extends enrollment past n_base: drift
    # keeps accruing at the same per-patient rate, which is the intended
    # behaviour (a longer trial spans more calendar time). Clamp to [0,1] so
    # rbinom() can never receive an out-of-range probability under large delta.
    t_frac  <- (i - 1L) / (n_base - 1L)
    p_e_now <- min(max(p_true_e + delta * t_frac, 0), 1)
    p_c_now <- min(max(p_c_data + delta * t_frac, 0), 1)

    # Allocation: Thompson Sampling or fixed 1:1
    arm <- if (!use_rar || i <= n_burn_rar)
      (i - 1L) %% 2L
    else
      as.integer(rbeta(1, ae, be) >= rbeta(1, ac, bc))

    # Outcome
    if (arm == 1L) {
      y <- rbinom(1, 1, p_e_now)
      ae <- ae + y; be <- be + (1L - y)
    } else {
      y <- rbinom(1, 1, p_c_now)
      ac <- ac + y; bc <- bc + (1L - y)
    }
    alloc_v[i] <- arm
    resp_v[i]  <- y

    # SSR (fires at n_ssr_trig, uses all data up to that point)
    if (use_ssr && i == n_ssr_trig && !ssr_done) {
      ssr_done <- TRUE
      x_pool <- sum(resp_v[1:i])
      n_new  <- ssr_n_target(x_pool, i, p0 = 0.25,
                              target_cp = 0.80, n_max = n_cap)
      if (n_new > n_total) n_total <- n_new
    }

    i <- i + 1L
  }

  enrolled <- i - 1L
  alloc_v  <- alloc_v[1:enrolled]
  resp_v   <- resp_v[1:enrolled]

  # Bayesian monitoring: same schedule regardless of use_ssr / use_rar
  actual_looks <- sort(unique(c(looks[looks <= n_base], n_total)))
  actual_looks <- actual_looks[actual_looks <= enrolled]

  cum_ne <- cumsum(alloc_v == 1L)
  cum_xe <- cumsum(resp_v  * (alloc_v == 1L))
  cum_xc <- cumsum(resp_v  * (alloc_v == 0L))

  max_pp <- 0

  for (lk in actual_looks) {
    ne <- cum_ne[lk];  xe <- cum_xe[lk]
    nc <- lk - ne;     xc <- cum_xc[lk]
    if (ne == 0L || nc == 0L) next

    # Posterior always uses all data up to this look (no stagger)
    pp <- map_posterior_prob(xe, ne, xc, nc, mix)
    if (pp > max_pp) max_pp <- pp
  }

  max_pp
}

# ── Helpers built on the max-pp representation ───────────────────────────────
# T1E at a given threshold, from a vector of per-trial max posterior probs.
t1e_at <- function(max_pp, gamma) mean(max_pp > gamma)

# Exact threshold: smallest gamma whose T1E does not exceed alpha.
#
# NOTE (2026-07 correction): gamma* was previously chosen from a 0.005-step
# grid, which left calibration T1E at 0.0290 against a 0.025 target. Reporting
# inflation relative to 0.025 then charged that calibration error to the
# scenario. Binary search removes it.
calibrate_gamma <- function(max_pp, alpha = 0.025,
                            lo = 0.80, hi = 0.9999, tol = 1e-6) {
  for (k in seq_len(200)) {
    mid <- 0.5 * (lo + hi)
    if (mean(max_pp > mid) > alpha) lo <- mid else hi <- mid
    if (hi - lo < tol) break
  }
  hi
}

# Replicate a scenario, returning the vector of per-trial max-pp values.
#
# NOTE: do NOT write replicate(n_sim, simulate_one_trial(...)). replicate()
# is defined as sapply(integer(n), eval.parent(substitute(function(...) expr))),
# so `...` inside the expression binds to that anonymous function's dots --
# which receive the sapply loop index -- rather than to this function's dots.
# The loop index would be passed positionally as `mix`. Capture the arguments
# first and use do.call().
max_pp_sample <- function(n_sim, ...) {
  args <- list(...)
  vapply(seq_len(n_sim),
         function(i) do.call(simulate_one_trial, args),
         numeric(1))
}

# ── Re-weight the vague component of a fitted mixture ────────────────────────
# Used by the prior-weight sensitivity sweep. The vague component is the flat
# one added by robustify(); its weight is set to `w_vague` and the informative
# components share the remainder in their original proportions.
#
# The vague component is identified as the one with the smallest effective
# sample size (a + b) rather than by matching exact parameters: robustify()
# parameterises the added component by ESS, so it is Beta(1,1) or Beta(.5,.5)
# depending on the `n` argument and RBesT version.
#
# NOTE (2026-07 correction): the sweep previously passed a scalar `w_inf`
# straight into simulate_one_trial(), which -- given the default-argument
# defect -- meant it swept the weight of a Beta(6.5,19.5) that had nothing to
# do with the fitted prior.
set_vague_weight <- function(mix, w_vague) {
  ess <- mix["a", ] + mix["b", ]
  if (ncol(mix) < 2L)
    stop("mixture has a single component; nothing to re-weight")
  idx_vague <- which.min(ess)
  is_vague <- seq_len(ncol(mix)) == idx_vague
  w <- mix["w", ]
  w_new <- numeric(length(w))
  w_new[is_vague]  <- w_vague
  inf_w <- w[!is_vague]
  w_new[!is_vague] <- (1 - w_vague) * inf_w / sum(inf_w)
  stopifnot(abs(sum(w_new) - 1) < 1e-8)
  rbind(w = w_new, a = mix["a", ], b = mix["b", ])
}
```

### Threshold calibration

$\gamma^*$ is calibrated under $H_0$ at the anchor $p_0 = 0.25$ with no trend.

```{r calibration}
#| cache: true
# One sample of max-pp values yields T1E at every threshold, so the whole
# calibration curve and the exact gamma* come from a single simulation run.
set.seed(1L)
calib_pp <- max_pp_sample(N_sim_calib,
                          mix      = prior_mix,
                          p_true_e = p0,
                          p_c_data = p0,      # no conflict
                          delta    = 0,       # no trend
                          use_ssr  = TRUE,    # calibrate with mechanisms on
                          use_rar  = TRUE)

gamma_star  <- calibrate_gamma(calib_pp, alpha)
t1e_at_star <- t1e_at(calib_pp, gamma_star)

gamma_grid <- seq(0.90, 0.99, by = 0.005)
t1e_calib  <- sapply(gamma_grid, function(g) t1e_at(calib_pp, g))

cat(sprintf("gamma* = %.5f  (achieved alpha = %.4f, target %.4f)\n",
            gamma_star, t1e_at_star, alpha))
cat(sprintf("MCSE at gamma* = %.4f  (N = %d)\n",
            mcse(t1e_at_star, N_sim_calib), N_sim_calib))

tibble(gamma = gamma_grid, t1e = t1e_calib) |>
  ggplot(aes(gamma, t1e)) +
  geom_line(color = ZET$ink, linewidth = 1) +
  geom_point(size = 2, color = ZET$ink) +
  geom_hline(yintercept = alpha,      linetype = "dashed", color = ZET$red) +
  geom_vline(xintercept = gamma_star, linetype = "dashed", color = ZET$green) +
  annotate("text", x = gamma_star + 0.002, y = 0.075,
           label = sprintf("γ* = %.3f", gamma_star),
           color = ZET$green2, hjust = 0, size = 3.5, fontface = "bold") +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(title    = "Monitoring Threshold Calibration",
       subtitle = sprintf("Target α = %.3f under H₀ (at anchor, no trend)", alpha),
       x = "Efficacy threshold γ", y = "Simulated Type I error") +
  theme_zetyra()
```

### Primary result: shared $\mathcal{F}_t$

```{r primary-sim}
#| cache: true
# PRIMARY: SSR fires at n=45 (same look as monitoring)
# Factors: baseline-rate departure (p = 0.30 vs. calibration anchor 0.25) + trend
#
# H0 SPECIFICATION: Both arms generated from p0_conflict (= 0.30).
# gamma* was calibrated at 0.25 -- this is the departure from that anchor.
# H0 (no treatment effect) is preserved: p_true_e = p_c_data = 0.30.
set.seed(2026L)
pp_primary <- replicate(
  N_sim_main,
  simulate_one_trial(mix      = prior_mix,
                     p_true_e   = p0_conflict,  # H0: same rate as control
                     p_c_data   = p0_conflict,
                     delta      = delta_trend,
                     n_ssr_trig = 45L,    # SSR fires at n=45 monitoring look
                     use_ssr    = TRUE,
                     use_rar    = TRUE))

t1e_primary   <- t1e_at(pp_primary, gamma_star)
inflation_abs <- t1e_primary - alpha
inflation_pct <- inflation_abs / alpha * 100

cat(sprintf("PRIMARY  T1E:  %.4f  (nominal %.3f)\n", t1e_primary, alpha))
cat(sprintf("Inflation:     +%.4f  (+%.1f%%)\n", inflation_abs, inflation_pct))
```

```{r mild-conflict-sim}
#| cache: true

# MILD DEPARTURE: p_e = p_c = 0.27, i.e. +2 pp above the DESIGN NULL p0 = 0.25
# (the rate at which gamma* was calibrated). Not defined relative to the
# robustified mixture mean, which is 0.3132 -- above both departure scenarios.
# All other parameters identical to primary simulation
set.seed(2028L)
pp_mild <- replicate(
  N_sim_main,
  simulate_one_trial(mix      = prior_mix,
                     p_true_e   = p0_conflict_mild,
                     p_c_data   = p0_conflict_mild,
                     delta      = delta_trend,
                     n_ssr_trig = 45L,
                     use_ssr    = TRUE,
                     use_rar    = TRUE))

t1e_mild      <- t1e_at(pp_mild, gamma_star)
inflation_mild <- t1e_mild - alpha
inflation_pct_mild <- inflation_mild / alpha * 100

cat(sprintf("MILD DEPARTURE T1E:  %.4f  (nominal %.3f)\n", t1e_mild, alpha))
cat(sprintf("Inflation:          +%.4f  (+%.1f%%)\n", inflation_mild, inflation_pct_mild))
cat(sprintf("vs. strong departure: %.4f  (strong = %.4f)\n", t1e_mild, t1e_primary))

# Comparison table
tribble(
  ~Scenario,                              ~`p (both arms)`, ~`T1E`,    ~`vs. nominal`,
  "At anchor (calibration)",              "0.25",        t1e_at_star,   "—",
  "Mild departure (p=0.27)",    "0.27",        t1e_mild,
    sprintf("%+.1f%%", inflation_pct_mild),
  "Strong departure (p=0.30)","0.30",        t1e_primary,
    sprintf("%+.1f%%", inflation_pct)
) |>
  mutate(T1E = sprintf("%.4f", T1E)) |>
  kable(caption = paste(
    "Excess grows with the size of the departure.",
    "Mild departure (+2pp from anchor 0.25): within plausible inter-study variation.",
    "Strong departure (+5pp from anchor): stress test.",
    "All other parameters identical across rows."),
    align = c("l","c","r","r"))
```

### Version B: SSR-timing sensitivity (same monitoring schedule)

Version B varies *when* SSR fires while holding the monitoring schedule fixed.

**Version B design:**

| Look | Primary (shared) | Version B control |
|------|-----------------|-------------------|
| n=30 | monitor | **monitor + SSR** |
| n=45 | **monitor + SSR** | monitor |
| n=60 | monitor | monitor |
| final | monitoring | monitoring |

The scheduled monitoring looks are the same in both designs; only the interim
at which SSR fires differs.

**This is a sensitivity analysis, not a clean negative control.** Moving SSR
from $n=45$ to $n=30$ also changes the data on which re-estimation is computed,
and therefore the distribution of the realised final sample size — and the final
look is taken at `n_total`, which that distribution determines. The comparison
confounds SSR timing with its consequences for final $N$. Read it as evidence on
SSR timing; isolating a co-occurrence effect would require reporting and
controlling the final-$N$ distributions.

```{r negative-control}
#| cache: true

# Version B control: n_ssr_trig=30 — SSR fires at n=30 look, not n=45
# All other parameters identical to primary (including H0 specification)
set.seed(2027L)
pp_neg <- replicate(
  N_sim_main,
  simulate_one_trial(mix      = prior_mix,
                     p_true_e   = p0_conflict,  # H0: same rate as control
                     p_c_data   = p0_conflict,
                     delta      = delta_trend,
                     n_ssr_trig = 30L,    # SSR fires at n=30 look (not n=45)
                     use_ssr    = TRUE,
                     use_rar    = TRUE))

t1e_neg       <- t1e_at(pp_neg, gamma_star)
pct_recovered <- (t1e_primary - t1e_neg) / inflation_abs * 100

cat(sprintf("Version B T1E: %.4f  (nominal %.3f)\n", t1e_neg, alpha))
cat(sprintf("Primary T1E:   %.4f\n", t1e_primary))
cat(sprintf("Attenuation:   %.1f%% of inflation removed by moving SSR to n=30\n",
            pct_recovered))
```

```{r results-plot, fig.cap="Type I error: calibration target, primary (SSR at n=45), and Version B (SSR at n=30). Moving SSR earlier changes the result little within the evaluated range, recovering about 6.5% of the excess over nominal. This is an SSR-timing sensitivity analysis, not a clean negative control: moving SSR also changes the realised final-N distribution. It does not apportion the excess among components, and no such attribution is made here."}
tibble(
  Scenario = factor(
    c("Calibration\n(at anchor,\nno trend)",
      "Primary\n(SSR at n = 45)",
      "Version B\n(SSR at n = 30)"),
    levels = c("Calibration\n(at anchor,\nno trend)",
               "Primary\n(SSR at n = 45)",
               "Version B\n(SSR at n = 30)")),
  T1E  = c(t1e_at_star, t1e_primary, t1e_neg),
  fill = c("a", "b", "c")
) |>
  ggplot(aes(Scenario, T1E, fill = fill)) +
  geom_col(width = 0.50) +
  geom_hline(yintercept = alpha, linetype = "dashed",
             color = ZET$red, linewidth = 0.8) +
  geom_text(aes(label = sprintf("%.4f", T1E)),
            vjust = -0.5, size = 3.8, fontface = "bold", color = ZET$ink) +
  scale_fill_manual(values = c(a = ZET$neutralF, b = ZET$green, c = ZET$teal)) +
  scale_y_continuous(limits = c(0, max(t1e_primary, t1e_neg, 0.04) * 1.30),
                     labels = percent_format(accuracy = 0.1)) +
  labs(title    = "Pipeline-Level Type I Error: SSR Timing Sensitivity",
       subtitle = bquote(
         "Two-arm design, MAP prior, departure" ~ p[c] == .(sprintf("%.2f", p0_conflict)) ~
         "(anchor" ~ .(sprintf("%.2f,", p0)) ~ "trend" ~ δ == .(sprintf("%.2f)", delta_trend))),
       x = NULL, y = "Type I error") +
  theme_zetyra() + theme(legend.position = "none")
```

```{r results-table}
tribble(
  ~Scenario, ~T1E, ~`vs. nominal`, ~Note,
  "Calibration (at anchor, no trend)", t1e_at_star, "—", "threshold target",
  "Primary: SSR at n = 45", t1e_primary,
    sprintf("%+.1f%%", inflation_pct), "SSR at the n=45 look",
  "Version B: SSR at n = 30", t1e_neg,
    sprintf("%+.1f%%", (t1e_neg - alpha)/alpha*100), "SSR at the n=30 look"
) |>
  mutate(T1E = sprintf("%.4f", T1E)) |>
  kable(caption = paste("Type I error summary. Version B tests SSR timing as a causal variable;",
                        "changing SSR timing alters the result little within the evaluated range."),
        align = c("l","r","r","l"))
```

---

## 5.4 Fixed-Threshold Factorial Ablation

Running four designs (SSR × RAR) shows what happens when each adaptive component
is switched off.

**These are ablations at a fixed threshold, not a decomposition.** $\gamma^*$ was
calibrated with SSR and RAR *active*. The ablation cells retain that threshold,
so they are not calibrated to $\alpha$: with both components disabled the
monitoring rule under $\gamma^*$ sits near 0.05 at the anchor with no trend,
twice nominal, before any departure or trend. Do not read the baseline cell as
an attributable share.

All four conditions use the same monitoring threshold $\gamma^*$ and the same
baseline-rate departure and time trend as the primary simulation.

```{r factorial}
#| cache: true

cat("Factorial decomposition: use_ssr x use_rar\n")
set.seed(3000L)

# Four conditions: 2x2 SSR x RAR
factorial_grid <- expand.grid(use_ssr = c(FALSE, TRUE),
                               use_rar = c(FALSE, TRUE))

t1e_factorial <- apply(factorial_grid, 1, function(row) {
  t1e_at(replicate(N_sim_sens,
    simulate_one_trial(mix    = prior_mix,
                       p_true_e = p0_conflict,  # H0: p_e = p_c = p0_conflict
                       p_c_data = p0_conflict,
                       delta    = delta_trend,
                       use_ssr  = as.logical(row["use_ssr"]),
                       use_rar  = as.logical(row["use_rar"]))), gamma_star)
})

factorial_res <- factorial_grid |>
  mutate(T1E = t1e_factorial,
         label = case_when(
           !use_ssr & !use_rar ~ "Baseline\n(no SSR, no RAR)",
           use_ssr  & !use_rar ~ "SSR only\n(fixed allocation)",
           !use_ssr & use_rar  ~ "RAR only\n(no SSR)",
           use_ssr  & use_rar  ~ "SSR + RAR\n(primary)"
         ))

# Factorial effects — prefixed _f_ to avoid collision with sensitivity t1e_ssr
t1e_f_base <- factorial_res$T1E[!factorial_res$use_ssr & !factorial_res$use_rar]
t1e_f_ssr  <- factorial_res$T1E[ factorial_res$use_ssr & !factorial_res$use_rar]
t1e_f_rar  <- factorial_res$T1E[!factorial_res$use_ssr &  factorial_res$use_rar]
t1e_f_both <- factorial_res$T1E[ factorial_res$use_ssr &  factorial_res$use_rar]

# Compute deltas directly from factorial_res — no intermediate variables
# that could be shadowed by later chunks
delta_f_ssr  <- t1e_f_ssr  - t1e_f_base
delta_f_rar  <- t1e_f_rar  - t1e_f_base
delta_f_int  <- t1e_f_both - t1e_f_ssr - t1e_f_rar + t1e_f_base

# Report the CELL VALUES first, then the contrasts.
#
# NOTE (2026-07 correction): earlier versions printed only the three contrasts.
# That concealed the single most important number in the study -- the baseline
# cell, in which Bayesian monitoring runs with NO adaptive component. If the
# baseline is already far above alpha, then switching the adaptive components
# off does not restore control. NOTE this cell keeps gamma*, which was
# calibrated with SSR+RAR active, so it is NOT at nominal even at the anchor
# (~0.05). It is an ablation, not an attributable share.
cat("Cell values (Type I error):\n")
cat(sprintf("  SSR + RAR disabled     = %.4f  (MCSE %.4f)\n",
            t1e_f_base, mcse(t1e_f_base, N_sim_sens)))
cat(sprintf("  + SSR only             = %.4f  (MCSE %.4f)\n",
            t1e_f_ssr,  mcse(t1e_f_ssr,  N_sim_sens)))
cat(sprintf("  + RAR only             = %.4f  (MCSE %.4f)\n",
            t1e_f_rar,  mcse(t1e_f_rar,  N_sim_sens)))
cat(sprintf("  + SSR and RAR          = %.4f  (MCSE %.4f)\n",
            t1e_f_both, mcse(t1e_f_both, N_sim_sens)))
cat(sprintf("\n  NOTE: ablation cells retain the full-design threshold and are\n"))
cat(sprintf("        not separately calibrated; do not read as attributable shares.\n"))
cat("\nContrasts:\n")
cat(sprintf("  Δ_SSR         = %+.4f\n", delta_f_ssr))
cat(sprintf("  Δ_RAR         = %+.4f\n", delta_f_rar))
cat(sprintf("  Δ_interaction = %+.4f  (super-additive if > 0)\n", delta_f_int))

# Bar chart
factorial_res |>
  mutate(label = factor(label, levels = label)) |>
  ggplot(aes(label, T1E, fill = use_ssr & use_rar)) +
  geom_col(width = 0.52) +
  geom_hline(yintercept = alpha, linetype = "dashed",
             color = ZET$red, linewidth = 0.8) +
  geom_text(aes(label = sprintf("%.4f", T1E)),
            vjust = -0.5, size = 3.8, fontface = "bold", color = ZET$ink) +
  scale_fill_manual(values = c(`FALSE` = ZET$neutralF, `TRUE` = ZET$green)) +
  scale_y_continuous(limits = c(0, max(t1e_factorial) * 1.30),
                     labels = percent_format(accuracy = 0.1)) +
  labs(title    = "Factorial Decomposition: SSR × RAR",
       subtitle = bquote(
         Delta[SSR] == .(sprintf("%+.4f", delta_f_ssr)) ~~ "|" ~~
         Delta[RAR] == .(sprintf("%+.4f", delta_f_rar)) ~~ "|" ~~
         Delta[interaction] == .(sprintf("%+.4f", delta_f_int))),
       x = NULL, y = "Type I error") +
  theme_zetyra() + theme(legend.position = "none")
```

```{r factorial-table}
tribble(
  ~Condition, ~`T1E`, ~`vs. baseline`,
  "Baseline (no SSR, no RAR)", t1e_f_base, "—",
  "SSR only (fixed allocation)", t1e_f_ssr, sprintf("%+.4f", delta_f_ssr),
  "RAR only (no SSR)",          t1e_f_rar, sprintf("%+.4f", delta_f_rar),
  "SSR + RAR (primary)",        t1e_f_both,sprintf("%+.4f", t1e_f_both - t1e_f_base)
) |>
  mutate(T1E = sprintf("%.4f", T1E)) |>
  kable(caption = paste("Factorial decomposition. Interaction term =",
                         sprintf("%.4f", delta_f_int),
                         ". If positive, co-occurrence produces super-additive inflation."),
        align = c("l","r","r"))
```

---

### Fresh-seed check on the SSR × RAR interaction

The interaction is the one composition result whose magnitude matters, so it is
re-estimated here on an independent seed with an independent replication count.
This is a repeat within the same R implementation, not an independent
reimplementation: it guards against a seed artefact, not against a shared
modelling error.

```{r factorial-fresh-seed}
#| cache: true
set.seed(73028L)
N_fresh <- 3000L

fresh <- vapply(seq_len(4), function(k) {
  us <- k %in% c(2, 4)          # SSR on for cells 2 and 4
  ur <- k %in% c(3, 4)          # RAR on for cells 3 and 4
  t1e_at(replicate(N_fresh,
    simulate_one_trial(mix = prior_mix,
                       p_true_e = p0_conflict, p_c_data = p0_conflict,
                       delta = delta_trend,
                       use_ssr = us, use_rar = ur)), gamma_star)
}, numeric(1))

names(fresh) <- c("base", "ssr", "rar", "both")
int_fresh <- fresh["both"] - fresh["ssr"] - fresh["rar"] + fresh["base"]
se_fresh  <- sqrt(sum(mcse(fresh, N_fresh)^2))

cat(sprintf("Fresh seed 73028, N = %d per cell\n", N_fresh))
for (nm in names(fresh))
  cat(sprintf("  %-5s = %.4f  (MCSE %.4f)\n", nm, fresh[nm], mcse(fresh[nm], N_fresh)))
cat(sprintf("\n  interaction = %+.4f  (MCSE %.4f)  95%% SI [%+.4f, %+.4f]\n",
            int_fresh, se_fresh, int_fresh - 1.96*se_fresh, int_fresh + 1.96*se_fresh))
cat(sprintf("  primary-run interaction for comparison: %+.4f\n", delta_f_int))
```

---

## 5.5 Baseline-Rate Departure × Time Trend

The primary simulation activates both factors simultaneously. Here we *cross* them
in a 2×2 design (baseline-rate departure × time trend) to see whether each raises
Type I error on its own, and whether their joint effect exceeds the additive
prediction.

This is a factor-crossing experiment, **not** mechanism isolation. Both factors feed
the monitoring rule, the SSR trigger and the RAR allocation rule simultaneously, and
the design does not separate those channels. We also describe the departures as
departures from the calibration anchor $p_0 = 0.25$, not as quantified prior–data
conflict, which would have to be defined against the prior predictive distribution.

```{r mechanism-isolation}
#| cache: true

cat("Factor crossing: 2x2 baseline-rate departure x trend\n")
set.seed(4000L)

mech_grid <- expand.grid(
  conflict = c(FALSE, TRUE),
  trend    = c(FALSE, TRUE)
)

t1e_mech <- apply(mech_grid, 1, function(row) {
  p_c <- if (row["conflict"]) p0_conflict else p0   # 0.30 vs 0.25
  d   <- if (row["trend"])    delta_trend  else 0    # 0.05 vs 0
  # H0: p_true_e = p_c_data (both arms at same rate, MAP prior is what conflicts)
  t1e_at(replicate(N_sim_sens,
    simulate_one_trial(mix    = prior_mix,
                       p_true_e = p_c,   # H0 preserved: same rate as control
                       p_c_data = p_c,
                       delta    = d,
                       use_ssr  = TRUE,
                       use_rar  = TRUE)), gamma_star)
})

mech_res <- mech_grid |>
  mutate(T1E   = t1e_mech,
         label = case_when(
           !conflict & !trend ~ "Neither\n(calibration)",
           conflict  & !trend ~ "Departure only\n(no trend)",
           !conflict & trend  ~ "Trend only\n(at anchor)",
           conflict  & trend  ~ "Both\n(primary)"
         ))

# Additive prediction vs. actual
t1e_neither  <- mech_res$T1E[!mech_res$conflict & !mech_res$trend]
t1e_conf_only<- mech_res$T1E[ mech_res$conflict & !mech_res$trend]
t1e_trend_only<-mech_res$T1E[!mech_res$conflict &  mech_res$trend]
t1e_both_mech<- mech_res$T1E[ mech_res$conflict &  mech_res$trend]

additive_pred <- t1e_conf_only + t1e_trend_only - t1e_neither
interaction_term <- t1e_both_mech - additive_pred

cat(sprintf("Departure only:   %.4f  (+%.4f vs. calibration)\n",
            t1e_conf_only,  t1e_conf_only  - t1e_neither))
cat(sprintf("Trend only:       %.4f  (+%.4f vs. calibration)\n",
            t1e_trend_only, t1e_trend_only - t1e_neither))
cat(sprintf("Additive prediction: %.4f\n", additive_pred))
cat(sprintf("Observed (both):     %.4f\n", t1e_both_mech))
cat(sprintf("Interaction term:    %+.4f  (amplification if > 0)\n", interaction_term))

mech_res |>
  mutate(label = factor(label, levels = label)) |>
  ggplot(aes(label, T1E,
             fill = factor(conflict + trend * 2,
                           labels = c("Neither","Departure","Trend","Both")))) +
  geom_col(width = 0.52) +
  geom_hline(yintercept = alpha, linetype = "dashed",
             color = ZET$red, linewidth = 0.8) +
  geom_hline(yintercept = additive_pred, linetype = "dotted",
             color = ZET$slate, linewidth = 1.0) +
  # ── in-panel annotation ─────────────────────────────────────────────────────
  # This figure carries no title or subtitle: the manuscript prints a caption
  # beneath every figure, so a title only repeats it. What the subtitle used to
  # carry is annotated inside the panel instead.
  #
  # The three positions below were checked against the bar tops, their value
  # labels and the legend box at the rendered 9 x 5.5 in size; each has under
  # one line height of slack in at least one direction. Re-check before nudging.
  #
  # Additive-prediction label: left-hand end, vertically centred between the
  # dotted additive line and the next gridline above it, so it crowds neither.
  # floor()+1 rather than ceiling() keeps this correct if additive_pred ever
  # lands exactly on a gridline.
  annotate("text", x = 0.55,
           y = (additive_pred + (floor(additive_pred / 0.05) + 1) * 0.05) / 2,
           label = sprintf("Additive prediction: %.4f", additive_pred),
           size = 4.1, fontface = "bold", color = ZET$ink, hjust = 0) +
  # The arc runs from the additive-prediction label to the observed value on the
  # "Both" bar. Its two endpoints are the two numbers whose difference IS the
  # interaction, so the span carries the meaning and the label above it only has
  # to name the quantity. Both endpoints sit above every bar, so the path stays
  # clear however the bar heights move; offsets are taken off the data rather
  # than hardcoded for the same reason.
  annotate("curve", x = 1.95, y = additive_pred + 0.012,
           xend = 3.82, yend = t1e_both_mech + 0.006,
           curvature = -0.20, angle = 90, ncp = 12,
           arrow = arrow(length = unit(0.18, "cm"), type = "closed"),
           color = ZET$green, linewidth = 0.6) +
  annotate("text", x = 2.85, y = t1e_both_mech + 0.026,
           label = sprintf("Interaction = %+.4f", interaction_term),
           size = 4.1, fontface = "bold", color = ZET$green, hjust = 0.5) +
  # Short form for the nominal-alpha line: the caption spells it out in full,
  # and only the gap between the first two bars is wide enough for a label.
  annotate("text", x = 1.5, y = alpha, label = "α = 0.025",
           hjust = 0.5, vjust = -0.45, size = 3.5,
           fontface = "bold", color = ZET$red) +
  geom_text(aes(label = sprintf("%.4f", T1E)),
            vjust = -0.5, size = 3.8, fontface = "bold", color = ZET$ink) +
  scale_fill_manual(values = c(Neither  = ZET$neutralF,
                                Departure = "#2563EB",
                                Trend    = ZET$teal,
                                Both     = ZET$green)) +
  scale_y_continuous(limits = c(0, max(t1e_mech) * 1.35),
                     labels = percent_format(accuracy = 0.1)) +
  labs(x = NULL, y = "Type I error", fill = NULL) +
  theme_zetyra() +
  theme(axis.title = element_text(size = 15, face = "bold", color = ZET$ink,
                                  margin = margin(t = 4, r = 4)),
        axis.text  = element_text(size = 13, color = ZET$slate),
        legend.position             = "inside",
        legend.position.inside      = c(0.005, 0.995),
        legend.justification.inside = c(0, 1),
        legend.direction            = "vertical",
        legend.text  = element_text(size = 11, color = ZET$slate),
        legend.background = element_rect(fill = "white", color = NA),
        legend.margin     = margin(2, 6, 2, 2))
```

```{r mechanism-table}
tribble(
  ~Condition, ~T1E, ~`vs. calibration`, ~Note,
  "Neither (calibration)",    t1e_neither,   "—",              "threshold target",
  "Departure only (no trend)", t1e_conf_only,
    sprintf("%+.4f", t1e_conf_only - t1e_neither),  "prior away from realised rate",
  "Trend only (at anchor)", t1e_trend_only,
    sprintf("%+.4f", t1e_trend_only - t1e_neither), "rate drifts from anchor",
  "Both (primary)",           t1e_both_mech,
    sprintf("%+.4f", t1e_both_mech - t1e_neither),  sprintf("interaction = %+.4f", interaction_term)
) |>
  mutate(T1E = sprintf("%.4f", T1E)) |>
  kable(caption = paste("2x2 crossing of baseline-rate departure and time trend.",
                        "Each factor alone raises T1E; jointly they exceed the",
                        "additive prediction. Both act through the monitoring rule,",
                        "so this crosses factors rather than isolating mechanisms."),
        align = c("l","r","r","l"))
```

---

## 5.6 Type I Error Across the Composite Null

The scenarios above each evaluate Type I error at a single control rate. But
$H_0$ here is composite: it holds for *every* common rate $p_e = p_c = p$, not
only at $p = 0.25$. Type I error control is a statement about the supremum
over that set, so a single-point evaluation cannot establish it.

Sweeping $p$ separates two questions that the single-point analysis conflates:
how the calibrated design behaves away from its calibration point, and what
enabling the adaptive components changes at that same threshold.  Note that the
comparison series holds $\gamma^*$ at the value calibrated for the *full*
design and merely disables SSR and RAR; it is a fixed-threshold ablation, not a
separately calibrated monitoring-only design.

Three caveats govern how the sweep should be read.

**The crossing point is not sharply identified.** The fixed-$\gamma^*$ difference is
clearly negative at low evaluated rates and clearly positive at high ones, but
the individual effects at $p = 0.25$ ($-0.0070$) and $p = 0.27$ ($+0.0078$)
are each within their Monte Carlo error. Locating the crossing would require
more replications or paired common-random-number contrasts; the defensible
claim is the direction change across the range, not its position.

**The SSR rule shares a threshold with the sweep.** SSR is a pooled
one-sample conditional-power rule tested against $p_0 = 0.25$, while the final
estimand is the two-arm comparison $p_e > p_c$. Because SSR activates near
that same 0.25, the apparent crossing may partly reflect the construction of
the SSR rule rather than a general property of composition.

**Extended trials accrue more drift.** With drift anchored to the planned
horizon $n_{\text{base}} = 90$, a trial expanded to $n = 150$ experiences
total drift of $0.05 \times 149/89 \approx 0.084$, not $0.05$. This is
coherent -- a longer trial spans more calendar time -- but it means the
estimated SSR effect and the interaction include the additional secular-trend
exposure created by extending enrollment. A fixed-terminal-drift alternative
would separate the two and is worth reporting alongside.

```{r composite-null-sweep}
#| cache: true
p_null_grid <- c(0.20, 0.22, 0.25, 0.27, 0.30, 0.33, 0.36)
set.seed(600L)

sweep_rows <- lapply(p_null_grid, function(pn) {
  mon <- t1e_at(replicate(N_sim_sens,
           simulate_one_trial(mix = prior_mix, p_true_e = pn, p_c_data = pn,
                              delta = delta_trend,
                              use_ssr = FALSE, use_rar = FALSE)), gamma_star)
  all <- t1e_at(replicate(N_sim_sens,
           simulate_one_trial(mix = prior_mix, p_true_e = pn, p_c_data = pn,
                              delta = delta_trend,
                              use_ssr = TRUE,  use_rar = TRUE)), gamma_star)
  tibble(p = pn, `SSR + RAR disabled` = mon, `All mechanisms` = all,
         `Fixed-gamma* difference` = all - mon)
})
composite_null <- bind_rows(sweep_rows)

composite_null |>
  mutate(across(c(`SSR + RAR disabled`, `All mechanisms`),
                ~ sprintf("%.4f", .x)),
         `Fixed-gamma* difference` = sprintf("%+.4f", `Fixed-gamma* difference`)) |>
  kable(caption = paste("Type I error across the composite null",
                        "(p_e = p_c = p, trend active). Both columns hold the",
                        "full-design threshold gamma* fixed; 'SSR + RAR",
                        "disabled' is an ablation, not a separately calibrated",
                        "design. Neither attains nominal across the range, and",
                        "the sign of the fixed-threshold difference depends on where in",
                        "the null the design is evaluated."),
        align = c("r","r","r","r"))
```

```{r composite-null-plot, fig.cap="Type I error across the composite null. Grey holds the full-design threshold gamma* with SSR and RAR disabled; green is the full design. At that fixed threshold, enabling SSR and RAR lowers T1E at small p and raises it at large p. Grey is a fixed-threshold ablation, not a separately calibrated monitoring-only design, so the gap between the series is not an attributable share."}
composite_null |>
  select(p, `SSR + RAR disabled`, `All mechanisms`) |>
  pivot_longer(-p, names_to = "Design", values_to = "T1E") |>
  ggplot(aes(p, T1E, color = Design, shape = Design)) +
  geom_hline(yintercept = alpha, linetype = "dashed",
             color = ZET$red, linewidth = 0.8) +
  # No title or subtitle: the manuscript captions every figure, so a title only
  # repeats it. The nominal-alpha reference therefore moves into the panel,
  # right-aligned at the high-p end where both series run far above the line --
  # the same placement as the presentation version of this figure.
  annotate("text", x = max(composite_null$p), y = alpha,
           label = "nominal α = 0.025", hjust = 1, vjust = -0.45,
           size = 4.3, fontface = "bold", color = ZET$red) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.8) +
  scale_color_manual(values = c(`SSR + RAR disabled` = ZET$neutralL,
                                `All mechanisms`  = ZET$green)) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = bquote("Common baseline rate " * p), y = "Type I error",
       color = NULL, shape = NULL) +
  theme_zetyra() +
  theme(axis.title = element_text(size = 15, face = "bold", color = ZET$ink,
                                  margin = margin(t = 4, r = 4)),
        axis.text  = element_text(size = 13, color = ZET$slate),
        legend.position             = "inside",
        legend.position.inside      = c(0.01, 0.99),
        legend.justification.inside = c(0, 1),
        legend.direction            = "vertical",
        legend.text  = element_text(size = 12, color = ZET$slate),
        legend.background = element_rect(fill = "white", color = "#D8DCE2"),
        legend.margin     = margin(4, 8, 4, 6))
```

---

## 5.7 Sensitivity Analysis


All three sweeps pass the relevant parameter directly into `simulate_one_trial()`.
No global variables are mutated.

```{r sens-prior-weight}
#| cache: true
# Swept quantity is the weight on the FLAT robustify() component: larger
# w_vague = less borrowing from the historical prior.
w_grid <- seq(0, 0.90, by = 0.10)
set.seed(300L)

t1e_w <- sapply(w_grid, function(w) {
  t1e_at(replicate(N_sim_sens,
                 simulate_one_trial(mix    = set_vague_weight(prior_mix, w),
                                    p_true_e = p0_conflict,
                                    p_c_data = p0_conflict,
                                    delta    = delta_trend,
                                    use_ssr  = TRUE,
                                    use_rar  = TRUE)), gamma_star)
})

p_w <- tibble(w = w_grid, t1e = t1e_w) |>
  ggplot(aes(w, t1e)) +
  geom_line(color = ZET$green, linewidth = 1.2) +
  geom_point(color = ZET$green, size = 2.5) +
  geom_hline(yintercept = alpha, linetype = "dashed", color = ZET$red) +
  scale_x_continuous(labels = percent_format()) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(title = "Vague-Component Weight", x = bquote(w[vague]), y = "T1E") +
  theme_zetyra()
```

```{r sens-time-trend}
#| cache: true
delta_grid <- seq(0, 0.15, by = 0.025)
set.seed(400L)

t1e_d <- sapply(delta_grid, function(d) {
  t1e_at(replicate(N_sim_sens,
                 simulate_one_trial(mix    = prior_mix,
                                    p_true_e = p0_conflict,
                                    p_c_data = p0_conflict,
                                    delta    = d,          # varied here
                                    use_ssr  = TRUE,
                                    use_rar  = TRUE)), gamma_star)
})

p_d <- tibble(d = delta_grid, t1e = t1e_d) |>
  ggplot(aes(d, t1e)) +
  geom_line(color = ZET$green, linewidth = 1.2) +
  geom_point(color = ZET$green, size = 2.5) +
  geom_hline(yintercept = alpha, linetype = "dashed", color = ZET$red) +
  scale_x_continuous(labels = percent_format()) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(title = "Time Trend Magnitude", x = bquote(δ), y = "T1E") +
  theme_zetyra()
```

```{r sens-ssr-timing}
#| cache: true
ssr_grid <- c(20L, 30L, 40L, 45L, 55L, 60L)
set.seed(500L)

t1e_ssr <- sapply(ssr_grid, function(nt) {
  t1e_at(replicate(N_sim_sens,
                 simulate_one_trial(mix      = prior_mix,
                                    p_true_e   = p0_conflict,  # H0: p_e = p_c = p0_conflict
                                    p_c_data   = p0_conflict,
                                    delta      = delta_trend,
                                    n_ssr_trig = nt,       # varied here
                                    use_ssr    = TRUE,
                                    use_rar    = TRUE)), gamma_star)
})

p_ssr <- tibble(
  nt    = ssr_grid,
  t1e   = t1e_ssr,
  label = sprintf("n=%d\n(%d%%)", ssr_grid,
                  as.integer(round(ssr_grid / N_base * 100)))
) |>
  ggplot(aes(label, t1e)) +
  geom_col(fill = ZET$green, width = 0.55, alpha = 0.90) +
  geom_hline(yintercept = alpha, linetype = "dashed", color = ZET$red) +
  geom_text(aes(label = sprintf("%.4f", t1e)),
            vjust = -0.5, size = 3.5, color = ZET$ink) +
  scale_y_continuous(limits = c(0, max(t1e_ssr) * 1.25),
                     labels = percent_format(accuracy = 0.1)) +
  labs(title = "SSR Trigger Timing", x = bquote(n[SSR]), y = "T1E") +
  theme_zetyra()
```

```{r sens-panel, fig.width=13, fig.height=4.5, fig.cap="Sensitivity: Type I error inflation vs. prior weight (left), time trend (centre), SSR timing (right). Dashed line = nominal alpha = 0.025. Direction does not reverse across any examined range."}
p_w | p_d | p_ssr
```

```{r sens-table}
tribble(
  ~Parameter,        ~Range,              ~`Min T1E`,             ~`Max T1E`,             ~`Direction reversal`,
  "Prior weight (w<sub>inf</sub>)",    "0% – 90%",
    sprintf("%.4f", min(t1e_w)),   sprintf("%.4f", max(t1e_w)),   "No",
  "Time trend (δ)",  "0% – 15%",
    sprintf("%.4f", min(t1e_d)),   sprintf("%.4f", max(t1e_d)),   "No",
  "SSR timing (n<sub>SSR</sub>)",      "n = 20, 30, 40, 45, 55, 60",
    sprintf("%.4f", min(t1e_ssr)), sprintf("%.4f", max(t1e_ssr)), "No"
) |>
  kable(caption = "Sensitivity bounds across all parameter ranges",
        align   = c("l","l","r","r","c"),
        escape  = FALSE)
```

---

## 5.8 Reproducibility

```{r auto-summary}
#| code-fold: false
cat(sprintf("gamma*                = %.3f\n", gamma_star))
cat(sprintf("T1E (calibration)     = %.4f  (target: %.3f)\n", t1e_at_star, alpha))
cat(sprintf("T1E (mild departure, p=0.27) = %.4f  (+%.1f%%)\n",
            t1e_mild, inflation_pct_mild))
cat(sprintf("T1E (strong departure, p=0.30) = %.4f  (+%.1f%%)\n",
            t1e_primary, inflation_pct))
cat(sprintf("T1E (Version B, n_ssr=30)     = %.4f  (+%.1f%%)\n",
            t1e_neg, (t1e_neg - alpha)/alpha*100))
cat(sprintf("Version B attenuation  = %.1f%% of primary inflation\n", pct_recovered))
cat(sprintf("\nFactorial decomposition:\n"))
cat(sprintf("  Delta_SSR         = %+.4f\n", delta_f_ssr))
cat(sprintf("  Delta_RAR         = %+.4f\n", delta_f_rar))
cat(sprintf("  Delta_interaction = %+.4f\n", delta_f_int))
cat(sprintf("\nBaseline-rate departure x trend:\n"))
cat(sprintf("  Departure only    = +%.4f\n", t1e_conf_only  - t1e_neither))
cat(sprintf("  Trend only        = +%.4f\n", t1e_trend_only - t1e_neither))
cat(sprintf("  Both (primary)    = +%.4f\n", t1e_both_mech  - t1e_neither))
cat(sprintf("  Interaction term  = %+.4f\n", interaction_term))
```

```{r mcse-table}
mcse_tbl <- tribble(
  ~Estimate,                     ~T1E,           ~N_sim,
  "Calibration (at anchor)",     t1e_at_star,    N_sim_calib,
  "Mild departure (p=0.27)",     t1e_mild,       N_sim_main,
  "Strong departure (p=0.30)",   t1e_primary,    N_sim_main,
  "Version B (SSR at n = 30)",     t1e_neg,        N_sim_main,
  "Factorial: baseline",         t1e_f_base,     N_sim_sens,
  "Factorial: SSR only",         t1e_f_ssr,      N_sim_sens,
  "Factorial: RAR only",         t1e_f_rar,      N_sim_sens,
  "Factorial: SSR + RAR",        t1e_f_both,     N_sim_sens,
  "Mech. isol.: neither",        t1e_neither,    N_sim_sens,
  "Departure x trend: departure only", t1e_conf_only, N_sim_sens,
  "Mech. isol.: trend only",     t1e_trend_only, N_sim_sens,
  "Mech. isol.: both",           t1e_both_mech,  N_sim_sens
) |>
  mutate(
    MCSE   = mcse(T1E, N_sim),
    CI_lo  = T1E - 1.96 * MCSE,
    CI_hi  = T1E + 1.96 * MCSE,
    T1E    = sprintf("%.4f", T1E),
    MCSE   = sprintf("%.4f", MCSE),
    `95% SI` = sprintf("[%.4f, %.4f]", CI_lo, CI_hi),
    N_sim  = formatC(N_sim, format = "d", big.mark = ",")
  ) |>
  select(Estimate, T1E, MCSE, `95% SI`, N_sim)

kable(mcse_tbl,
      caption = "Monte Carlo standard errors and 95% simulation intervals for all key T1E estimates.",
      align = c("l", "r", "r", "c", "r"))

# ── Propagated MCSEs for derived interaction terms ──────────────────────────
# Interaction = T1E_both - T1E_A - T1E_B + T1E_neither
# All four simulations are independent, so Var(interaction) = sum of variances

# Departure x trend interaction
se_mech <- sqrt(
  mcse(t1e_both_mech,  N_sim_sens)^2 +
  mcse(t1e_conf_only,  N_sim_sens)^2 +
  mcse(t1e_trend_only, N_sim_sens)^2 +
  mcse(t1e_neither,    N_sim_sens)^2
)

# Factorial interaction
se_fact <- sqrt(
  mcse(t1e_f_both, N_sim_sens)^2 +
  mcse(t1e_f_ssr,  N_sim_sens)^2 +
  mcse(t1e_f_rar,  N_sim_sens)^2 +
  mcse(t1e_f_base, N_sim_sens)^2
)

int_tbl <- tribble(
  ~`Interaction term`,                    ~Estimate,        ~MCSE,    ~`95% SI`,
  "Departure × trend", interaction_term, se_mech,
    sprintf("[%+.4f, %+.4f]", interaction_term - 1.96*se_mech, interaction_term + 1.96*se_mech),
  "Factorial (SSR × RAR)",                  delta_f_int,     se_fact,
    sprintf("[%+.4f, %+.4f]", delta_f_int - 1.96*se_fact, delta_f_int + 1.96*se_fact)
) |>
  mutate(Estimate = sprintf("%+.4f", Estimate),
         MCSE    = sprintf("%.4f", MCSE))

kable(int_tbl,
      caption = paste("Propagated MCSEs for derived interaction terms.",
                      "Both simulation intervals exclude zero. Note the",
                      "factorial interaction is symmetric: RAR has no",
                      "marginal effect without SSR but raises T1E when SSR",
                      "is active (equivalently, SSR's effect is larger under",
                      "RAR)."),
      align = c("l", "r", "r", "c"))
```

**Seeds:**

| Simulation | Seed |
|---|---|
| MAP prior MCMC | 42 (via `set.seed()` before `gMAP()`) |
| RAR validation | 42 |
| Threshold calibration | 1 |
| Primary: strong departure (p=0.30) | 2026 |
| Primary: mild departure (p=0.27) | 2028 |
| Version B control (n<sub>SSR</sub> = 30) | 2027 |
| Factorial decomposition | 3000 |
| Departure × trend (2x2) | 4000 |
| Sensitivity — prior weight | 300 |
| Sensitivity — time trend | 400 |
| Sensitivity — SSR timing | 500 |

```{r session-info}
#| code-fold: false
sessionInfo()
```

---

## References

::: {#refs}
:::
