---
title: "Section 5: Simulation Study — Rule B"
subtitle: "Composed Pipeline Operating Characteristics with Blinded Nuisance-Parameter SSR"
author: "Lu Qian · Zetyra | Evidence in the Wild"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-float: true
    code-fold: true
    code-tools: true
    theme:
      - cosmo
      - custom.scss
    fig-width: 9
    fig-height: 5.5
    df-print: kable
    embed-resources: true
  pdf:
    toc: true
    number-sections: true
    fig-width: 7
    fig-height: 4.5
    geometry: margin=1in
    fontsize: 11pt
execute:
  echo: true
  warning: false
  message: false
  cache: true
  cache-path: "_cache/"
bibliography: references.bib
---

```{r setup}
#| cache: false
library(RBesT)
library(ggplot2)
library(dplyr)
library(tibble)
library(purrr)
library(tidyr)
library(knitr)
library(patchwork)
library(scales)

# ── Color palette ─────────────────────────────────────────────────────────────
# Publication palette with white backgrounds and accessible contrast.
ZET <- list(
  ink      = "#1A1A2E",   # deep navy-black — titles, axis text
  green    = "#1B3A2E",   # forest green  — primary series, headers
  green2   = "#2D6A4F",   # mid green     — secondary series
  teal     = "#40916C",   # light green   — tertiary / timing sensitivity
  slate    = "#4B5563",   # dark slate    — axis text, subtitles, legends
  gridline = "#E5E7EB",   # cool light gray — major gridlines
  red      = "#B91C1C",   # accessible red — reference lines, FAIL
  amber    = "#D97706",   # amber         — warning / border
  bg       = "#FFFFFF",   # pure white    — plot background
  panel    = "#FFFFFF",   # white         — panel background
  muted    = "#545C6B",   # dark grey     — captions, reference annotations (6.7:1)
  neutralL = "#46505F",   # darker grey   — grey DATA series (8.0:1 on white)
  neutralF = "#8B95A3"    # mid grey fill — baseline/neutral BARS (3.0:1, WCAG 1.4.11)
)

# ── Cache invalidation on palette change ──────────────────────────────────────
# Include the palette in every cache key so all figures regenerate together
# whenever a colour changes.
knitr::opts_chunk$set(cache.extra = list(zet_palette = ZET))

# ── Modern publication theme ──────────────────────────────────────────────────
theme_zetyra <- function(base_size = 12) {
  theme_minimal(base_size = base_size, base_family = "sans") +
    theme(
      # Backgrounds — clean white, no colored fills
      plot.background   = element_rect(fill = ZET$bg,    color = NA),
      panel.background  = element_rect(fill = ZET$panel, color = NA),

      # Grid — thin horizontal only; vertical removed for clarity
      panel.grid.major.x = element_blank(),
      panel.grid.major.y = element_line(color = ZET$gridline, linewidth = 0.45),
      panel.grid.minor   = element_blank(),

      # Axes — subtle bottom/left rule; no full box
      axis.line.x  = element_line(color = "#D1D5DB", linewidth = 0.5),
      axis.ticks.x = element_line(color = "#D1D5DB", linewidth = 0.4),
      axis.ticks.length = unit(3, "pt"),
      axis.text    = element_text(color = ZET$slate,  size = base_size - 1),
      axis.title   = element_text(color = ZET$ink,    size = base_size,
                                   face = "bold", margin = margin(t = 4, r = 4)),

      # Titles — left-aligned, strong hierarchy
      plot.title    = element_text(color = ZET$ink,   size = base_size + 3,
                                   face = "bold",     hjust = 0,
                                   margin = margin(b = 4)),
      plot.subtitle = element_text(color = ZET$slate, size = base_size - 1,
                                   hjust = 0,         margin = margin(b = 10)),
      plot.caption  = element_text(color = ZET$muted, size = base_size - 2,
                                   hjust = 0,         margin = margin(t = 8)),
      plot.margin   = margin(12, 16, 10, 12),

      # Legend — top-right, no box
      legend.position    = "top",
      legend.justification = "left",
      legend.background  = element_rect(fill = NA, color = NA),
      legend.key         = element_rect(fill = NA, color = NA),
      legend.text        = element_text(color = ZET$slate, size = base_size - 1),
      legend.title       = element_text(color = ZET$ink,   size = base_size - 1,
                                         face = "bold"),
      legend.spacing.x   = unit(6, "pt"),

      # Facet strips — light gray pill-style
      strip.background = element_rect(fill = ZET$gridline, color = NA,
                                       linewidth = 0),
      strip.text       = element_text(color = ZET$ink, face = "bold",
                                       size = base_size - 1, margin = margin(4,6,4,6))
    )
}
```

---

## Overview

This document implements the simulation study for a two-arm Phase II oncology
design with historical borrowing on the control arm, Bayesian posterior
monitoring, a blinded nuisance-parameter sample size re-estimation (SSR) rule,
and response-adaptive randomization (RAR).

**Rule B.** At the SSR look, the pooled response rate is used only as a blinded
nuisance-parameter estimate. The planned treatment effect remains fixed at
$0.45-0.25=0.20$, and the two-arm sample size is recomputed under a prespecified
working 1:1 allocation. The rule permits increases only, from $N=90$ up to
the prespecified cap $N=150$; with these planning inputs its actual mapping at
the $n=45$ look ranges only from $N=90$ to $N=100$. Because the operational
design uses Thompson-sampling RAR and a Bayesian final decision, this is a
conventional-style working nuisance rule, not an exact power calculation for
the full adaptive design.

**Primary claim — robustness.** An efficacy threshold calibrated at a single
assumed control rate does not deliver Type I error control across the composite
null. Ablations that switch off the adaptive components while *holding that
threshold fixed* do not restore control either.

**Secondary finding — composition.** At the frozen full-design threshold, the
combined SSR/RAR contrast is negative at the lower evaluated null rates and
positive at the upper evaluated rates. The factorial interaction estimates are
imprecise and are not interpreted as evidence of a stable super-additive effect.

**Scope.** The grid demonstrates a control failure: the largest observed
pointwise Type I error is 0.2272, so the supremum is at least 0.2272. It does
not locate the true supremum. Scenarios are
departures of the baseline rate from the calibration anchor $p_0 = 0.25$, not
quantified prior–data conflict. On the plausibility of departures of this size,
@tang_2010 treat a five-percentage-point control-rate drift as clinically modest
in an oncology simulation, and @viele_2014 review the conditions under which
historical and concurrent control rates diverge. Allocation targets follow the
optimal-allocation literature [@rosenberger_optimal; @rosenberger_lachin].

The platform and its benchmark framework are documented in the Zetyra white
paper [@zetyra_whitepaper]; the Bayesian calculators used here are documented
separately [@zetyra_bayesian_toolkit].

Simulation code:
`github.com/evidenceinthewild/Composed-Adaptive-Trial-Validation`

---

## 5.1 Design Parameters

```{r design-params}
# ── Trial parameters ──────────────────────────────────────────────────────────
p0       <- 0.25   # null / historical control response rate
p_alt    <- 0.45   # alternative response rate
alpha    <- 0.025  # one-sided Type I error target

# ── Sample size & looks ───────────────────────────────────────────────────────
N_base   <- 90L
N_max    <- 150L
n_ssr    <- 45L    # SSR trigger at 50% of N_base

# ── MAP prior ────────────────────────────────────────────────────────────────
# The prior is FITTED from historical data in 5.2.1 and stored in `prior_mix`,
# a matrix with rows (w, a, b), one column per mixture component.
# The simulation function requires this fitted object explicitly.
prior_mix <- NULL   # assigned in 5.2.1; simulations require it

# ── Departures from the calibration anchor & time trend ──────────────────────
# These are departures from the design anchor p0 = 0.25, not quantified
# prior-data conflict. The relevant reference rates are:
#   design null / historical rate       0.2500
#   dominant (informative) component    0.2568
#   full robustified mixture            0.3132
p_departure_strong <- 0.30   # +5 pp above the design anchor
p_departure_mild   <- 0.27   # +2 pp above the design anchor
delta_trend <- 0.05   # linear drift over full enrollment
p_null_grid <- c(0.20, 0.22, 0.25, 0.27, 0.30, 0.33, 0.36)

# ── Monte Carlo & simulation counts ──────────────────────────────────────────
# Posterior probabilities use deterministic quadrature; N_sim controls only
# the number of simulated trials.
N_sim_calib  <- 20000L
N_sim_main   <- 10000L
N_sim_sens   <-  5000L

# ── Monte Carlo SE helper ───────────────────────────────────────────────────
# Binomial MCSE for a proportion estimated from N_sim replications
mcse <- function(p_hat, n_sim) sqrt(p_hat * (1 - p_hat) / n_sim)

# 95% simulation interval
mcse_ci <- function(p_hat, n_sim) {
  se <- mcse(p_hat, n_sim)
  c(lower = p_hat - 1.96 * se, upper = p_hat + 1.96 * se)
}
```

---

## 5.2 Component-Level Validation

### 5.2.1 MAP Prior — Benchmarked Against RBesT [@weber_rbest]

```{r map-prior-fit}
hist_data <- data.frame(
  study = c("H1", "H2", "H3"),
  n     = c( 50,   40,   60),
  r     = c( 13,   10,   15)
)

set.seed(42)
map_fit <- gMAP(
  cbind(r, n - r) ~ 1 | study,
  data       = hist_data,
  family     = binomial,
  tau.dist   = "HalfNormal",
  tau.prior  = 1,
  beta.prior = 2,
  chains     = 4,
  iter       = 4000,
  warmup     = 1000
)

map_mix <- automixfit(map_fit, Nc = 2)
rob_mix <- robustify(map_mix, weight = 0.20, mean = 0.5)
print(rob_mix)
```

```{r map-prior-validation, fig.cap="MAP prior density: components and mixture from the fitted rob_mix object. Analytic density reproduces dmix() to floating-point precision."}

p_seq <- seq(0.01, 0.99, length.out = 500)

# Parameters from fitted rob_mix (matrix with rows "w","a","b"; cols = components)
w_fit   <- rob_mix["w", ]   # mixture weights
a_fit   <- rob_mix["a", ]   # Beta alpha parameters
b_fit   <- rob_mix["b", ]   # Beta beta  parameters

cat("Fitted mixture components:\n")
print(rbind(w = round(w_fit, 4), a = round(a_fit, 3), b = round(b_fit, 3)))

# Analytic density using the EXTRACTED parameters
rbest_dens    <- dmix(rob_mix, p_seq)          # RBesT reference
analytic_dens <- sapply(p_seq, function(p)
  sum(w_fit * dbeta(p, a_fit, b_fit)))         # our formula with fitted params

max_dev_map <- max(abs(rbest_dens - analytic_dens))

# Carry every component returned by automixfit() and robustify() into the
# simulation.
prior_mix <- rbind(w = w_fit, a = a_fit, b = b_fit)

cat(sprintf("\nPrior carried into simulation (%d components):\n", ncol(prior_mix)))
print(round(prior_mix, 4))
cat(sprintf("  mixture mean = %.4f\n", sum(w_fit * a_fit / (a_fit + b_fit))))
stopifnot(abs(sum(w_fit) - 1) < 1e-8)

# Component densities for plot
df_comps <- tibble(
  p     = rep(p_seq, length(w_fit) + 1),
  dens  = c(sapply(seq_along(w_fit), function(j)
               w_fit[j] * dbeta(p_seq, a_fit[j], b_fit[j])),
             rbest_dens),
  label = rep(c(sprintf("Component %d (w=%.2f)", seq_along(w_fit), w_fit),
                "Mixture (RBesT)"), each = 500)
)

comp_colors <- c("#2166AC", "#4393C3", "#D6604D", "#B2182B")
comp_labels <- levels(factor(df_comps$label))
names(comp_colors) <- comp_labels[seq_along(comp_colors)]

ggplot(df_comps, aes(p, dens, color = label, linetype = label)) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = comp_colors) +
  labs(title    = "MAP Prior: Fitted Mixture Components and Density",
       subtitle = sprintf(
         "Max |analytic - dmix| = %.2e  (threshold: 1e-10, floating-point only)",
         max_dev_map),
       x = bquote("Control response rate" ~ p[c]), y = "Density",
       color = NULL, linetype = NULL) +
  theme_zetyra() + theme(legend.position = "bottom")

cat(sprintf("\nMAP PRIOR:  max dev = %.2e  [floating-point precision]  %s\n",
            max_dev_map, ifelse(max_dev_map < 1e-8, "PASS", "FAIL")))
```

### 5.2.2 Posterior Probability Helper

The monitoring rule is a posterior-probability threshold applied at each
interim look; see @zhou_ji_2024 for the design family and the question of
whether such rules require adjustment for interim analyses.

```{r posterior-prob-fn}
# ── Conjugate update of an arbitrary Beta-mixture prior ──────────────────────
# mix: matrix with rows (w, a, b), one column per component.
mixture_posterior <- function(mix, x_c, n_c) {
  w <- mix["w", ]; a <- mix["a", ]; b <- mix["b", ]
  a_post <- a + x_c
  b_post <- b + (n_c - x_c)
  lml <- lbeta(a_post, b_post) - lbeta(a, b)      # log marginal likelihood
  lw  <- log(w) + lml
  lw  <- lw - max(lw)                              # numerically stable
  w_post <- exp(lw); w_post <- w_post / sum(w_post)
  rbind(w = w_post, a = a_post, b = b_post)
}

# ── Gauss-Legendre nodes on [0,1] for deterministic integration ──────────────
# Golub-Welsch: nodes/weights are the eigenvalues/first-components of the
# symmetric tridiagonal Jacobi matrix. Base R only -- no extra dependency.
gauss_legendre_01 <- function(n) {
  k <- 1:(n - 1)
  b <- k / sqrt(4 * k^2 - 1)
  J <- diag(0, n)
  J[cbind(k, k + 1)] <- b
  J[cbind(k + 1, k)] <- b
  e <- eigen(J, symmetric = TRUE)
  ord <- order(e$values)
  x <- e$values[ord]                  # nodes on [-1, 1]
  w <- 2 * (e$vectors[1, ord])^2      # weights on [-1, 1]
  list(x = 0.5 * (x + 1), w = 0.5 * w)   # mapped to [0, 1]
}
.gl <- gauss_legendre_01(400)
.QX <- .gl$x
.QW <- .gl$w

# Verify normalization and polynomial moment exactness on the quadrature rule
# used by the simulation.
.g8 <- gauss_legendre_01(8L)
stopifnot(
  abs(sum(.QW) - 1)                     < 1e-12,   # int 1    = 1
  abs(sum(.QW * .QX) - 0.5)             < 1e-12,   # int x    = 1/2
  abs(sum(.QW * .QX^2) - 1 / 3)         < 1e-12,   # int x^2  = 1/3
  abs(sum(.g8$w * .g8$x^15) - 1 / 16)   < 1e-10,   # exact to degree 2n-1
  abs(sum(.QW * dbeta(.QX, 3, 7)) - 1)  < 1e-10)   # a smooth Beta density
rm(.g8)

# ── P(p_e > p_c | data), by deterministic quadrature ─────────────────────────
# = int f_e(x) F_c(x) dx, where p_e ~ Beta(.5+x_e, .5+n_e-x_e) (Jeffreys, no
# borrowing on the experimental arm) and p_c ~ posterior mixture.
#
# This calculation is deterministic, not exact. The sanity checks above verify the
# Gauss-Legendre weights and one smooth Beta integral; they do not bound the
# approximation error for every state the adaptive design can visit. Extreme
# all-success or all-failure states can concentrate posterior mass near a
# boundary and are the most demanding cases for a fixed-node rule.
#
# All reported simulations use the specified 400-node rule. A dedicated
# convergence study would be needed to establish invariance to the node count.
map_posterior_prob <- function(x_e, n_e, x_c, n_c, mix) {
  post <- mixture_posterior(mix, x_c, n_c)
  f_e <- dbeta(.QX, 0.5 + x_e, 0.5 + (n_e - x_e))
  F_c <- rowSums(vapply(seq_len(ncol(post)),
                        function(j) post["w", j] *
                          pbeta(.QX, post["a", j], post["b", j]),
                        numeric(length(.QX))))
  sum(.QW * f_e * F_c)
}
```

### 5.2.3 Rule B — Blinded Nuisance-Parameter Re-estimation

Rule B holds the planned treatment effect fixed and re-estimates only the
nuisance event-rate level from the pooled interim responses. The reference
calculation uses a prespecified working 1:1 allocation. Its target is the
fixed-sample Wald power implied by the original $N=90$ design at
$(p_c,p_e)=(0.25,0.45)$; it is not presented as the actual power of the
Thompson-sampling RAR design or of the Bayesian monitoring rule. The pooled
construction follows the blinded binary internal-pilot principle described by
@friede_kieser_2004.

The SSR decision consumes only $(x_{\mathrm{pool}},n_{\mathrm{int}})$, permits
increases only, and is capped at $N=150$. Arm labels, arm-specific outcomes,
posterior probabilities, and the realised allocation ratio are not inputs.

```{r ssr-helpers}
# Fixed planning quantities. The power target is derived from, rather than
# imposed on, the original N = 90 design.
delta_plan <- p_alt - p0
z_alpha_ssr <- qnorm(1 - alpha)
var_plan <- p0 * (1 - p0) + p_alt * (1 - p_alt)
n_arm_plan <- N_base / 2
z_power_plan <- delta_plan * sqrt(n_arm_plan / var_plan) - z_alpha_ssr
power_plan <- pnorm(z_power_plan)

# Core Rule B calculation from a blinded pooled rate. Under working 1:1
# allocation, the fixed planned effect is imposed symmetrically around p_pool.
ssr_nuisance_rate <- function(p_pool, n_int, n_base = N_base, n_max = N_max,
                              delta = delta_plan,
                              target_power = power_plan) {
  stopifnot(length(p_pool) == 1L, is.finite(p_pool),
            p_pool >= 0, p_pool <= 1,
            n_int >= 1L, n_base >= n_int, n_max >= n_base,
            delta > 0, delta < 1,
            target_power > 0, target_power < 1)

  # Truncation keeps both working rates inside (0,1) at extreme pooled counts.
  p_c_work <- min(max(p_pool - delta / 2, 1e-6), 1 - delta - 1e-6)
  p_e_work <- p_c_work + delta
  var_work <- p_c_work * (1 - p_c_work) +
              p_e_work * (1 - p_e_work)

  n_arm <- (z_alpha_ssr + qnorm(target_power))^2 * var_work / delta^2
  n_new <- 2L * as.integer(ceiling(n_arm))
  min(max(n_new, as.integer(n_base), as.integer(n_int)), as.integer(n_max))
}

# Operational interface: the only outcome input is the pooled response count.
ssr_nuisance <- function(x_int, n_int, n_base = N_base, n_max = N_max,
                         delta = delta_plan,
                         target_power = power_plan) {
  stopifnot(length(x_int) == 1L, is.finite(x_int),
            x_int >= 0, x_int <= n_int)
  ssr_nuisance_rate(x_int / n_int, n_int, n_base, n_max,
                    delta, target_power)
}

cat(sprintf("Planned effect held fixed: %.2f\n", delta_plan))
cat(sprintf("Working allocation: 1:1 (%d per arm at N = %d)\n",
            as.integer(n_arm_plan), N_base))
cat(sprintf("Reference Wald power preserved by Rule B: %.4f\n", power_plan))
```

```{r ssr-validation}
# Exhaust the complete support of the blinded count at the planned SSR look.
ssr_lookup <- tibble(
  x_pool = 0:n_ssr,
  n_int = n_ssr,
  p_pool = x_pool / n_int,
  final_N = vapply(x_pool, function(x)
    ssr_nuisance(x, n_ssr, n_base = N_base, n_max = N_max), integer(1))
)

# Independent algebra checks: exact planning rates reproduce N = 90; every
# operational decision is even, increase-only, and inside the prespecified cap.
n_at_planning_rates <- ssr_nuisance_rate(
  (p0 + p_alt) / 2, n_ssr, n_base = N_base, n_max = N_max)
stopifnot(
  n_at_planning_rates == N_base,
  all(ssr_lookup$final_N >= N_base),
  all(ssr_lookup$final_N <= N_max),
  all(ssr_lookup$final_N %% 2L == 0L)
)

# Selected blinded states make the adaptation's magnitude transparent.
ssr_demo <- ssr_lookup |>
  filter(x_pool %in% c(0L, 9L, 11L, 14L, 16L, 18L, 22L, 27L, 36L, 45L)) |>
  mutate(
    p_c_work = pmin(pmax(p_pool - delta_plan / 2, 1e-6),
                     1 - delta_plan - 1e-6),
    p_e_work = p_c_work + delta_plan
  ) |>
  select(x_pool, n_int, p_pool, p_c_work, p_e_work, final_N)

kable(ssr_demo, digits = 3, row.names = FALSE,
      caption = paste(
        "Rule B decisions at selected pooled interim counts.",
        "The planned effect remains 0.20; only the blinded nuisance level changes."),
      col.names = c("Pooled responses", "Interim N", "Pooled rate",
                    "Working p_c", "Working p_e", "Final N"))

# Exact Poisson-binomial distribution of the pooled count under each composite-
# null grid point with the active time trend. Under p_e(t) = p_c(t), this pooled
# distribution is independent of the realised treatment-allocation path.
poisson_binomial_pmf <- function(prob) {
  out <- c(1, rep(0, length(prob)))
  for (j in seq_along(prob)) {
    old <- out
    out[1] <- old[1] * (1 - prob[j])
    for (k in 2:(j + 1L))
      out[k] <- old[k] * (1 - prob[j]) + old[k - 1L] * prob[j]
  }
  out
}

ssr_size_summary <- bind_rows(lapply(p_null_grid, function(p) {
  prob <- pmin(pmax(p + delta_trend * (0:(n_ssr - 1L)) /
                      (N_base - 1L), 0), 1)
  pmf <- poisson_binomial_pmf(prob)
  final_n <- ssr_lookup$final_N
  tibble(
    p = p,
    `Pr(N > 90)` = sum(pmf[final_n > N_base]),
    `Expected N` = sum(pmf * final_n),
    `Maximum possible N` = max(final_n[pmf > 0])
  )
}))

kable(ssr_size_summary, digits = 4, row.names = FALSE,
      caption = paste(
        "Exact sample-size behavior at the n = 45 SSR look under the evaluated",
        "composite-null scenarios with the active time trend."))

cat(sprintf("Rule B validation: PASS\n"))
cat(sprintf("  N at exact planning nuisance rate: %d\n", n_at_planning_rates))
cat(sprintf("  Operational final-N range over x = 0,...,%d: %d to %d\n",
            n_ssr, min(ssr_lookup$final_N), max(ssr_lookup$final_N)))
```

### 5.2.4 RAR — Benchmarked Against Analytic Allocation [@rosenberger_optimal]

```{r rar-validation}
#| cache: true
# P(X > Y) where X ~ Beta(a1,b1), Y ~ Beta(a2,b2) via numerical integration
prob_beta_gt <- function(a1, b1, a2, b2) {
  integrate(function(x) dbeta(x, a1, b1) * pbeta(x, a2, b2),
            lower = 0, upper = 1, rel.tol = 1e-8)$value
}

# Validate: after observing fixed outcomes, does one-step TS allocation
# proportion match P(Beta(ae,be) > Beta(ac,bc))?
# We do n_reps single-step allocations from a known posterior state.

single_step_ts <- function(ae, be, ac, bc, n_reps = 100000L) {
  mean(rbeta(n_reps, ae, be) >= rbeta(n_reps, ac, bc))
}

set.seed(42)
# Test scenarios: known posterior states after observing specific data
# ae = 0.5 + x_e, be = 0.5 + (n_e - x_e); similarly for control
rar_scen <- tribble(
  ~x_e, ~n_e, ~x_c, ~n_c,
    5,   10,    2,   10,   # clearly favors experimental
    3,   10,    5,   10,   # clearly favors control
    5,   10,    5,   10,   # symmetric (should give ~0.5)
    8,   20,    4,   20,   # moderate asymmetry
   12,   30,    8,   30    # larger sample
)

rar_res <- rar_scen |>
  mutate(
    ae = 0.5 + x_e, be = 0.5 + (n_e - x_e),
    ac = 0.5 + x_c, bc = 0.5 + (n_c - x_c),
    # Analytical reference: P(Beta(ae,be) > Beta(ac,bc))
    ref_alloc = pmap_dbl(list(ae, be, ac, bc),
                         ~ prob_beta_gt(..1, ..2, ..3, ..4)),
    # Empirical: 100k single-step TS draws
    emp_alloc = pmap_dbl(list(ae, be, ac, bc),
                         ~ single_step_ts(..1, ..2, ..3, ..4)),
    deviation = abs(emp_alloc - ref_alloc),
    pass      = deviation < 0.005   # Monte Carlo error at n=100k is ~0.001
  )

kable(
  rar_res |> select(x_e, n_e, x_c, n_c, ref_alloc, emp_alloc, deviation, pass),
  digits = 4,
  caption = paste(
    "RAR validation: Thompson Sampling one-step allocation probability vs.",
    "analytical P(Beta(ae,be) > Beta(ac,bc)) via numerical integration.",
    "n=100k MC draws; expected Monte Carlo SE ≈ 0.001."),
  col.names = c("x<sub>e</sub>","n<sub>e</sub>","x<sub>c</sub>","n<sub>c</sub>","Analytical P(θ<sub>e</sub>>θ<sub>c</sub>)",
                "TS proportion","|dev|","Pass"),
  escape = FALSE)

cat(sprintf("RAR:  max dev = %.5f  [threshold 0.005]  %s\n",
            max(rar_res$deviation), ifelse(all(rar_res$pass), "PASS", "FAIL")))
```

---

## 5.3 Composed Pipeline Simulation

### Core simulation function

```{r sim-fn}
# simulate_one_trial() -----------------------------------------------------------
#
# Returns the maximum posterior probability across interim looks; see the note
# below. Convert to a rejection decision with t1e_at(pp, gamma).
#
# Two-arm Phase II design with historical borrowing on the control arm via a
# MAP prior, three co-occurring adaptive mechanisms:
#   MAP Bayesian monitoring, Rule B blinded nuisance SSR (if use_ssr),
#   Thompson Sampling RAR (if use_rar)
#
# The earlier-SSR sensitivity changes n_ssr_trig from 45 to 30 while retaining
# the same monitoring schedule and active mechanisms. Because timing also
# changes the final-N distribution, it is not a causal isolation exercise.
#
# Factorial decomposition: use_ssr x use_rar gives four conditions.
# Factor crossing: vary p_c_data (baseline-rate departure) and delta (trend).

# Returns the MAXIMUM posterior probability across interim looks.
#
# Data generation does not depend on gamma -- the enrollment loop completes
# before monitoring begins -- so rejection at threshold gamma is exactly
# {max_pp > gamma}. Returning max_pp lets a single simulated sample yield the
# Type I error at EVERY threshold, which makes exact calibration cheap.
#
simulate_one_trial <- function(
    mix,                  # REQUIRED: fitted prior, rows (w, a, b)
    p_true_e   = 0.25,
    p_c_data   = 0.25,    # control data-generating rate
    delta      = 0,       # time trend magnitude
    n_ssr_trig = 45L,     # when SSR fires (primary=45, timing sensitivity=30)
    use_ssr    = TRUE,    # FALSE = no sample size adaptation
    use_rar    = TRUE,    # FALSE = fixed 1:1 alternating allocation
    n_base     = 90L,
    n_cap      = 150L,
    looks      = c(30L, 45L, 60L, 90L),
    n_burn_rar = 10L) {

  stopifnot(!missing(mix), is.matrix(mix), all(c("w","a","b") %in% rownames(mix)))

  n_total  <- n_base
  ssr_done <- FALSE

  alloc_v <- integer(n_cap)
  resp_v  <- numeric(n_cap)

  ae <- 0.5; be <- 0.5   # TS posteriors (experimental)
  ac <- 0.5; bc <- 0.5   # TS posteriors (control)

  i <- 1L
  while (i <= n_total) {

    # Calendar drift uses the fixed planned horizon. Thus an extension beyond
    # n_base accrues additional drift at the same per-patient rate. Clamp the
    # resulting probabilities to the valid Bernoulli range.
    t_frac  <- (i - 1L) / (n_base - 1L)
    p_e_now <- min(max(p_true_e + delta * t_frac, 0), 1)
    p_c_now <- min(max(p_c_data + delta * t_frac, 0), 1)

    # Allocation: Thompson Sampling or fixed 1:1
    arm <- if (!use_rar || i <= n_burn_rar)
      (i - 1L) %% 2L
    else
      as.integer(rbeta(1, ae, be) >= rbeta(1, ac, bc))

    # Outcome
    if (arm == 1L) {
      y <- rbinom(1, 1, p_e_now)
      ae <- ae + y; be <- be + (1L - y)
    } else {
      y <- rbinom(1, 1, p_c_now)
      ac <- ac + y; bc <- bc + (1L - y)
    }
    alloc_v[i] <- arm
    resp_v[i]  <- y

    # SSR (fires at n_ssr_trig, uses all data up to that point)
    if (use_ssr && i == n_ssr_trig && !ssr_done) {
      ssr_done <- TRUE
      x_pool <- sum(resp_v[1:i])
      n_new  <- ssr_nuisance(x_pool, i, n_base = n_base, n_max = n_cap)
      if (n_new > n_total) n_total <- n_new
    }

    i <- i + 1L
  }

  enrolled <- i - 1L
  alloc_v  <- alloc_v[1:enrolled]
  resp_v   <- resp_v[1:enrolled]

  # Bayesian monitoring: same schedule regardless of use_ssr / use_rar
  actual_looks <- sort(unique(c(looks[looks <= n_base], n_total)))
  actual_looks <- actual_looks[actual_looks <= enrolled]

  cum_ne <- cumsum(alloc_v == 1L)
  cum_xe <- cumsum(resp_v  * (alloc_v == 1L))
  cum_xc <- cumsum(resp_v  * (alloc_v == 0L))

  max_pp <- 0

  for (lk in actual_looks) {
    ne <- cum_ne[lk];  xe <- cum_xe[lk]
    nc <- lk - ne;     xc <- cum_xc[lk]
    if (ne == 0L || nc == 0L) next

    # Posterior always uses all data up to this look (no stagger)
    pp <- map_posterior_prob(xe, ne, xc, nc, mix)
    if (pp > max_pp) max_pp <- pp
  }

  max_pp
}

# ── Helpers built on the max-pp representation ───────────────────────────────
# T1E at a given threshold, from a vector of per-trial max posterior probs.
t1e_at <- function(max_pp, gamma) mean(max_pp > gamma)

# Empirical threshold: smallest gamma whose simulated T1E does not exceed alpha.
calibrate_gamma <- function(max_pp, alpha = 0.025,
                            lo = 0.80, hi = 0.9999, tol = 1e-6) {
  for (k in seq_len(200)) {
    mid <- 0.5 * (lo + hi)
    if (mean(max_pp > mid) > alpha) lo <- mid else hi <- mid
    if (hi - lo < tol) break
  }
  hi
}

# Replicate a scenario, returning the vector of per-trial max-pp values.
#
# Do not write replicate(n_sim, simulate_one_trial(...)). replicate()
# is defined as sapply(integer(n), eval.parent(substitute(function(...) expr))),
# so `...` inside the expression binds to that anonymous function's dots --
# which receive the sapply loop index -- rather than to this function's dots.
# The loop index would be passed positionally as `mix`. Capture the arguments
# first and use do.call().
max_pp_sample <- function(n_sim, ...) {
  args <- list(...)
  vapply(seq_len(n_sim),
         function(i) do.call(simulate_one_trial, args),
         numeric(1))
}

# ── Re-weight the vague component of a fitted mixture ────────────────────────
# Used by the prior-weight sensitivity sweep. The vague component is the flat
# one added by robustify(); its weight is set to `w_vague` and the informative
# components share the remainder in their original proportions.
#
# The vague component is identified as the one with the smallest effective
# sample size (a + b) rather than by matching exact parameters: robustify()
# parameterises the added component by ESS, so it is Beta(1,1) or Beta(.5,.5)
# depending on the `n` argument and RBesT version.
set_vague_weight <- function(mix, w_vague) {
  ess <- mix["a", ] + mix["b", ]
  if (ncol(mix) < 2L)
    stop("mixture has a single component; nothing to re-weight")
  idx_vague <- which.min(ess)
  is_vague <- seq_len(ncol(mix)) == idx_vague
  w <- mix["w", ]
  w_new <- numeric(length(w))
  w_new[is_vague]  <- w_vague
  inf_w <- w[!is_vague]
  w_new[!is_vague] <- (1 - w_vague) * inf_w / sum(inf_w)
  stopifnot(abs(sum(w_new) - 1) < 1e-8)
  rbind(w = w_new, a = mix["a", ], b = mix["b", ])
}
```

### Threshold calibration

$\gamma^*$ is recalibrated with Rule B active under $H_0$ at the anchor
$p_0 = 0.25$ with no trend.

```{r calibration}
#| cache: true
# One sample of max-pp values yields T1E at every threshold, so the whole
# calibration curve and the exact gamma* come from a single simulation run.
set.seed(1L)
calib_pp <- max_pp_sample(N_sim_calib,
                          mix      = prior_mix,
                          p_true_e = p0,
                          p_c_data = p0,      # no anchor departure
                          delta    = 0,       # no trend
                          use_ssr  = TRUE,    # calibrate with mechanisms on
                          use_rar  = TRUE)

gamma_star  <- calibrate_gamma(calib_pp, alpha)
t1e_at_star <- t1e_at(calib_pp, gamma_star)

gamma_grid <- seq(0.90, 0.99, by = 0.005)
t1e_calib  <- sapply(gamma_grid, function(g) t1e_at(calib_pp, g))

cat(sprintf("gamma* = %.5f  (achieved alpha = %.4f, target %.4f)\n",
            gamma_star, t1e_at_star, alpha))
cat(sprintf("MCSE at gamma* = %.4f  (N = %d)\n",
            mcse(t1e_at_star, N_sim_calib), N_sim_calib))

tibble(gamma = gamma_grid, t1e = t1e_calib) |>
  ggplot(aes(gamma, t1e)) +
  geom_line(color = ZET$ink, linewidth = 1) +
  geom_point(size = 2, color = ZET$ink) +
  geom_hline(yintercept = alpha,      linetype = "dashed", color = ZET$red) +
  geom_vline(xintercept = gamma_star, linetype = "dashed", color = ZET$green) +
  annotate("text", x = gamma_star + 0.002, y = 0.075,
           label = sprintf("γ* = %.3f", gamma_star),
           color = ZET$green2, hjust = 0, size = 3.5, fontface = "bold") +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(title    = "Monitoring Threshold Calibration",
       subtitle = sprintf("Target α = %.3f under H₀ (at anchor, no trend)", alpha),
       x = "Efficacy threshold γ", y = "Simulated Type I error") +
  theme_zetyra()
```

```{r anchor-validation}
#| cache: true
# Fresh-seed validation is separate from the sample used to choose gamma*.
set.seed(101L)
anchor_validation_pp <- max_pp_sample(
  N_sim_sens,
  mix      = prior_mix,
  p_true_e = p0,
  p_c_data = p0,
  delta    = 0,
  use_ssr  = TRUE,
  use_rar  = TRUE)
anchor_validation_t1e <- t1e_at(anchor_validation_pp, gamma_star)
anchor_validation_ci <- mcse_ci(anchor_validation_t1e, N_sim_sens)

cat(sprintf(
  "Independent anchor validation: T1E = %.4f, MCSE = %.4f, 95%% SI [%.4f, %.4f] (N = %d)\n",
  anchor_validation_t1e,
  mcse(anchor_validation_t1e, N_sim_sens),
  anchor_validation_ci["lower"], anchor_validation_ci["upper"],
  N_sim_sens))
```

### Primary result: shared $\mathcal{F}_t$

```{r primary-sim}
#| cache: true
# PRIMARY: SSR fires at n=45 (same look as monitoring)
# Factors: baseline-rate departure (p = 0.30 vs. calibration anchor 0.25) + trend
#
# H0 SPECIFICATION: Both arms generated from p_departure_strong (= 0.30).
# gamma* was calibrated at 0.25 -- this is the departure from that anchor.
# H0 (no treatment effect) is preserved: p_true_e = p_c_data = 0.30.
set.seed(2026L)
pp_primary <- replicate(
  N_sim_main,
  simulate_one_trial(mix      = prior_mix,
                     p_true_e   = p_departure_strong,  # H0: same rate as control
                     p_c_data   = p_departure_strong,
                     delta      = delta_trend,
                     n_ssr_trig = 45L,    # SSR fires at n=45 monitoring look
                     use_ssr    = TRUE,
                     use_rar    = TRUE))

t1e_primary   <- t1e_at(pp_primary, gamma_star)
inflation_abs <- t1e_primary - alpha
inflation_pct <- inflation_abs / alpha * 100

cat(sprintf("PRIMARY  T1E:  %.4f  (nominal %.3f)\n", t1e_primary, alpha))
cat(sprintf("Inflation:     +%.4f  (+%.1f%%)\n", inflation_abs, inflation_pct))
```

```{r mild-departure-sim}
#| cache: true

# MILD DEPARTURE: p_e = p_c = 0.27, i.e. +2 pp above the DESIGN NULL p0 = 0.25
# (the rate at which gamma* was calibrated). Not defined relative to the
# robustified mixture mean, which is 0.3132 -- above both departure scenarios.
# All other parameters identical to primary simulation
set.seed(2028L)
pp_mild <- replicate(
  N_sim_main,
  simulate_one_trial(mix      = prior_mix,
                     p_true_e   = p_departure_mild,
                     p_c_data   = p_departure_mild,
                     delta      = delta_trend,
                     n_ssr_trig = 45L,
                     use_ssr    = TRUE,
                     use_rar    = TRUE))

t1e_mild      <- t1e_at(pp_mild, gamma_star)
inflation_mild <- t1e_mild - alpha
inflation_pct_mild <- inflation_mild / alpha * 100

cat(sprintf("MILD DEPARTURE T1E:  %.4f  (nominal %.3f)\n", t1e_mild, alpha))
cat(sprintf("Inflation:          +%.4f  (+%.1f%%)\n", inflation_mild, inflation_pct_mild))
cat(sprintf("vs. strong departure: %.4f  (strong = %.4f)\n", t1e_mild, t1e_primary))

# Comparison table
tribble(
  ~Scenario,                              ~`p (both arms)`, ~`T1E`,    ~`vs. nominal`,
  "At anchor (calibration)",              "0.25",        t1e_at_star,   "—",
  "Mild departure (p=0.27)",    "0.27",        t1e_mild,
    sprintf("%+.1f%%", inflation_pct_mild),
  "Strong departure (p=0.30)","0.30",        t1e_primary,
    sprintf("%+.1f%%", inflation_pct)
) |>
  mutate(T1E = sprintf("%.4f", T1E)) |>
  kable(caption = paste(
    "Excess grows with the size of the departure.",
    "Mild departure (+2pp from anchor 0.25): within plausible inter-study variation.",
    "Strong departure (+5pp from anchor): stress test.",
    "All other parameters identical across rows."),
    align = c("l","c","r","r"))
```

### Earlier-SSR timing sensitivity (same monitoring schedule)

This sensitivity varies *when* Rule B fires while holding the monitoring
schedule fixed.

**Timing-sensitivity design:**

| Look | Primary (shared) | Earlier-SSR sensitivity |
|------|-----------------|-------------------|
| n=30 | monitor | **monitor + SSR** |
| n=45 | **monitor + SSR** | monitor |
| n=60 | monitor | monitor |
| final | monitoring | monitoring |

The scheduled monitoring looks are the same in both designs; only the interim
at which SSR fires differs.

**This is a sensitivity analysis, not a clean negative control.** Moving SSR
from $n=45$ to $n=30$ also changes the data on which re-estimation is computed,
and therefore the distribution of the realised final sample size — and the final
look is taken at `n_total`, which that distribution determines. The comparison
confounds SSR timing with its consequences for final $N$. Read it as evidence on
SSR timing; isolating a co-occurrence effect would require reporting and
controlling the final-$N$ distributions.

```{r ssr-timing-comparison}
#| cache: true

# Earlier-SSR sensitivity: n_ssr_trig=30, not n=45
# All other parameters identical to primary (including H0 specification)
set.seed(2027L)
pp_neg <- replicate(
  N_sim_main,
  simulate_one_trial(mix      = prior_mix,
                     p_true_e   = p_departure_strong,  # H0: same rate as control
                     p_c_data   = p_departure_strong,
                     delta      = delta_trend,
                     n_ssr_trig = 30L,    # SSR fires at n=30 look (not n=45)
                     use_ssr    = TRUE,
                     use_rar    = TRUE))

t1e_neg       <- t1e_at(pp_neg, gamma_star)
timing_change <- t1e_neg - t1e_primary
timing_change_pct_excess <- timing_change / inflation_abs * 100

cat(sprintf("Earlier-SSR T1E: %.4f  (nominal %.3f)\n", t1e_neg, alpha))
cat(sprintf("Primary T1E:   %.4f\n", t1e_primary))
cat(sprintf("Timing change: %+.4f (%+.1f%% of the primary excess over nominal)\n",
            timing_change, timing_change_pct_excess))
```

```{r results-plot, fig.cap="Type I error under Rule B: calibration target, primary SSR timing (n=45), and earlier SSR timing (n=30). This is a timing sensitivity, not a clean negative control: moving SSR also changes the realised final-N distribution. It does not apportion the excess among components."}
tibble(
  Scenario = factor(
    c("Calibration\n(at anchor,\nno trend)",
      "Primary\n(SSR at n = 45)",
      "Earlier SSR\n(n = 30)"),
    levels = c("Calibration\n(at anchor,\nno trend)",
               "Primary\n(SSR at n = 45)",
               "Earlier SSR\n(n = 30)")),
  T1E  = c(t1e_at_star, t1e_primary, t1e_neg),
  fill = c("a", "b", "c")
) |>
  ggplot(aes(Scenario, T1E, fill = fill)) +
  geom_col(width = 0.50) +
  geom_hline(yintercept = alpha, linetype = "dashed",
             color = ZET$red, linewidth = 0.8) +
  geom_text(aes(label = sprintf("%.4f", T1E)),
            vjust = -0.5, size = 3.8, fontface = "bold", color = ZET$ink) +
  scale_fill_manual(values = c(a = ZET$neutralF, b = ZET$green, c = ZET$teal)) +
  scale_y_continuous(limits = c(0, max(t1e_primary, t1e_neg, 0.04) * 1.30),
                     labels = percent_format(accuracy = 0.1)) +
  labs(title    = "Pipeline-Level Type I Error: SSR Timing Sensitivity",
       subtitle = bquote(
         "Two-arm design, MAP prior, departure" ~ p[c] == .(sprintf("%.2f", p_departure_strong)) ~
         "(anchor" ~ .(sprintf("%.2f,", p0)) ~ "trend" ~ δ == .(sprintf("%.2f)", delta_trend))),
       x = NULL, y = "Type I error") +
  theme_zetyra() + theme(legend.position = "none")
```

```{r results-table}
tribble(
  ~Scenario, ~T1E, ~`vs. nominal`, ~Note,
  "Calibration (at anchor, no trend)", t1e_at_star, "—", "threshold target",
  "Primary: SSR at n = 45", t1e_primary,
    sprintf("%+.1f%%", inflation_pct), "SSR at the n=45 look",
  "Earlier SSR: n = 30", t1e_neg,
    sprintf("%+.1f%%", (t1e_neg - alpha)/alpha*100), "SSR at the n=30 look"
) |>
  mutate(T1E = sprintf("%.4f", T1E)) |>
  kable(caption = paste("Type I error summary under Rule B.",
                        "The n = 30 row is an SSR-timing sensitivity, not a causal isolation."),
        align = c("l","r","r","l"))
```

---

## 5.4 Fixed-Threshold Factorial Ablation

Running four designs (SSR × RAR) shows what happens when each adaptive component
is switched off.

**These are ablations at a fixed threshold, not a decomposition.** $\gamma^*$ was
calibrated with SSR and RAR *active*. The ablation cells retain that threshold,
so they are not calibrated to $\alpha$. Do not read the baseline cell as an
attributable share.

All four conditions use the same monitoring threshold $\gamma^*$ and the same
baseline-rate departure and time trend as the primary simulation.

```{r factorial}
#| cache: true
#| fig-cap: "Fixed-threshold factorial ablation under Rule B. The displayed interaction is a point estimate; its propagated 95% Monte Carlo simulation interval includes zero, and the fresh-seed check below is also compatible with zero."

cat("Fixed-threshold factorial ablation: use_ssr x use_rar\n")
set.seed(3000L)

# Four conditions: 2x2 SSR x RAR
factorial_grid <- expand.grid(use_ssr = c(FALSE, TRUE),
                               use_rar = c(FALSE, TRUE))

t1e_factorial <- apply(factorial_grid, 1, function(row) {
  t1e_at(replicate(N_sim_sens,
    simulate_one_trial(mix    = prior_mix,
                       p_true_e = p_departure_strong,  # H0: p_e = p_c = p_departure_strong
                       p_c_data = p_departure_strong,
                       delta    = delta_trend,
                       use_ssr  = as.logical(row["use_ssr"]),
                       use_rar  = as.logical(row["use_rar"]))), gamma_star)
})

factorial_res <- factorial_grid |>
  mutate(T1E = t1e_factorial,
         label = case_when(
           !use_ssr & !use_rar ~ "Baseline\n(no SSR, no RAR)",
           use_ssr  & !use_rar ~ "SSR only\n(fixed allocation)",
           !use_ssr & use_rar  ~ "RAR only\n(no SSR)",
           use_ssr  & use_rar  ~ "SSR + RAR\n(primary)"
         ))

# Factorial effects — prefixed _f_ to avoid collision with sensitivity t1e_ssr
t1e_f_base <- factorial_res$T1E[!factorial_res$use_ssr & !factorial_res$use_rar]
t1e_f_ssr  <- factorial_res$T1E[ factorial_res$use_ssr & !factorial_res$use_rar]
t1e_f_rar  <- factorial_res$T1E[!factorial_res$use_ssr &  factorial_res$use_rar]
t1e_f_both <- factorial_res$T1E[ factorial_res$use_ssr &  factorial_res$use_rar]

# Compute deltas directly from factorial_res — no intermediate variables
# that could be shadowed by later chunks
delta_f_ssr  <- t1e_f_ssr  - t1e_f_base
delta_f_rar  <- t1e_f_rar  - t1e_f_base
delta_f_int  <- t1e_f_both - t1e_f_ssr - t1e_f_rar + t1e_f_base

# Report the cell values before the contrasts. Every cell retains gamma*, so
# these are fixed-threshold ablations rather than attributable shares.
cat("Cell values (Type I error):\n")
cat(sprintf("  SSR + RAR disabled     = %.4f  (MCSE %.4f)\n",
            t1e_f_base, mcse(t1e_f_base, N_sim_sens)))
cat(sprintf("  + SSR only             = %.4f  (MCSE %.4f)\n",
            t1e_f_ssr,  mcse(t1e_f_ssr,  N_sim_sens)))
cat(sprintf("  + RAR only             = %.4f  (MCSE %.4f)\n",
            t1e_f_rar,  mcse(t1e_f_rar,  N_sim_sens)))
cat(sprintf("  + SSR and RAR          = %.4f  (MCSE %.4f)\n",
            t1e_f_both, mcse(t1e_f_both, N_sim_sens)))
cat(sprintf("\n  Ablation caveat: cells retain the full-design threshold and are\n"))
cat(sprintf("        not separately calibrated; do not read as attributable shares.\n"))
cat("\nContrasts:\n")
cat(sprintf("  Δ_SSR         = %+.4f\n", delta_f_ssr))
cat(sprintf("  Δ_RAR         = %+.4f\n", delta_f_rar))
cat(sprintf("  Δ_interaction = %+.4f  (point estimate; MC uncertainty below)\n",
            delta_f_int))

# Bar chart
factorial_res |>
  mutate(label = factor(label, levels = label)) |>
  ggplot(aes(label, T1E, fill = use_ssr & use_rar)) +
  geom_col(width = 0.52) +
  geom_hline(yintercept = alpha, linetype = "dashed",
             color = ZET$red, linewidth = 0.8) +
  geom_text(aes(label = sprintf("%.4f", T1E)),
            vjust = -0.5, size = 3.8, fontface = "bold", color = ZET$ink) +
  scale_fill_manual(values = c(`FALSE` = ZET$neutralF, `TRUE` = ZET$green)) +
  scale_y_continuous(limits = c(0, max(t1e_factorial) * 1.30),
                     labels = percent_format(accuracy = 0.1)) +
  labs(title    = "Fixed-Threshold Factorial Ablation: SSR × RAR",
       subtitle = bquote(
         Delta[SSR] == .(sprintf("%+.4f", delta_f_ssr)) ~~ "|" ~~
         Delta[RAR] == .(sprintf("%+.4f", delta_f_rar)) ~~ "|" ~~
         Delta[interaction] == .(sprintf("%+.4f", delta_f_int)) ~ "(point estimate)"),
       x = NULL, y = "Type I error") +
  theme_zetyra() + theme(legend.position = "none")
```

```{r factorial-table}
tribble(
  ~Condition, ~`T1E`, ~`vs. baseline`,
  "Baseline (no SSR, no RAR)", t1e_f_base, "—",
  "SSR only (fixed allocation)", t1e_f_ssr, sprintf("%+.4f", delta_f_ssr),
  "RAR only (no SSR)",          t1e_f_rar, sprintf("%+.4f", delta_f_rar),
  "SSR + RAR (primary)",        t1e_f_both,sprintf("%+.4f", t1e_f_both - t1e_f_base)
) |>
  mutate(T1E = sprintf("%.4f", T1E)) |>
  kable(caption = paste("Fixed-threshold factorial ablation. Interaction point estimate =",
                         sprintf("%.4f", delta_f_int),
                         "; uncertainty and a fresh-seed check are reported below."),
        align = c("l","r","r"))
```

---

### Fresh-seed check on the SSR × RAR interaction

The interaction is the one composition result whose magnitude matters, so it is
re-estimated here on an independent seed with an independent replication count.
This is a repeat within the same R implementation, not an independent
reimplementation: it guards against a seed artefact, not against a shared
modelling error.

```{r factorial-fresh-seed}
#| cache: true
set.seed(73028L)
N_fresh <- 3000L

fresh <- vapply(seq_len(4), function(k) {
  us <- k %in% c(2, 4)          # SSR on for cells 2 and 4
  ur <- k %in% c(3, 4)          # RAR on for cells 3 and 4
  t1e_at(replicate(N_fresh,
    simulate_one_trial(mix = prior_mix,
                       p_true_e = p_departure_strong, p_c_data = p_departure_strong,
                       delta = delta_trend,
                       use_ssr = us, use_rar = ur)), gamma_star)
}, numeric(1))

names(fresh) <- c("base", "ssr", "rar", "both")
int_fresh <- fresh["both"] - fresh["ssr"] - fresh["rar"] + fresh["base"]
se_fresh  <- sqrt(sum(mcse(fresh, N_fresh)^2))

cat(sprintf("Fresh seed 73028, N = %d per cell\n", N_fresh))
for (nm in names(fresh))
  cat(sprintf("  %-5s = %.4f  (MCSE %.4f)\n", nm, fresh[nm], mcse(fresh[nm], N_fresh)))
cat(sprintf("\n  interaction = %+.4f  (MCSE %.4f)  95%% SI [%+.4f, %+.4f]\n",
            int_fresh, se_fresh, int_fresh - 1.96*se_fresh, int_fresh + 1.96*se_fresh))
cat(sprintf("  primary-run interaction for comparison: %+.4f\n", delta_f_int))
```

The fresh-seed estimate and the primary estimate are both compatible with zero
at their Monte Carlo precision. Rule B therefore does not support a stable
factorial SSR-by-RAR interaction claim in this run.

---

## 5.5 Baseline-Rate Departure × Time Trend

The primary simulation activates both factors simultaneously. Here we *cross* them
in a 2×2 design (baseline-rate departure × time trend) to see whether each raises
Type I error on its own, and whether their joint effect exceeds the additive
prediction.

This is a factor-crossing experiment, **not** mechanism isolation. Both factors feed
the monitoring rule, the SSR trigger and the RAR allocation rule simultaneously, and
the design does not separate those channels. We also describe the departures as
departures from the calibration anchor $p_0 = 0.25$, not as quantified prior–data
conflict, which would have to be defined against the prior predictive distribution.

```{r departure-trend-crossing}
#| cache: true
#| fig-cap: "Type I error in the baseline-rate-departure by time-trend crossing under Rule B. The annotated interaction is a point estimate; its propagated 95% Monte Carlo simulation interval includes zero."

cat("Factor crossing: 2x2 baseline-rate departure x trend\n")
set.seed(4000L)

departure_trend_grid <- expand.grid(
  departure = c(FALSE, TRUE),
  trend    = c(FALSE, TRUE)
)

t1e_departure_trend <- apply(departure_trend_grid, 1, function(row) {
  p_c <- if (row["departure"]) p_departure_strong else p0   # 0.30 vs 0.25
  d   <- if (row["trend"])    delta_trend  else 0    # 0.05 vs 0
  # H0 is preserved because both arms use the same data-generating rate.
  t1e_at(replicate(N_sim_sens,
    simulate_one_trial(mix    = prior_mix,
                       p_true_e = p_c,   # H0 preserved: same rate as control
                       p_c_data = p_c,
                       delta    = d,
                       use_ssr  = TRUE,
                       use_rar  = TRUE)), gamma_star)
})

departure_trend_res <- departure_trend_grid |>
  mutate(T1E   = t1e_departure_trend,
         label = case_when(
           !departure & !trend ~ "Neither\n(calibration)",
           departure  & !trend ~ "Departure only\n(no trend)",
           !departure & trend  ~ "Trend only\n(at anchor)",
           departure  & trend  ~ "Both\n(primary)"
         ))

# Additive prediction vs. actual
t1e_neither <- departure_trend_res$T1E[
  !departure_trend_res$departure & !departure_trend_res$trend]
t1e_departure_only <- departure_trend_res$T1E[
  departure_trend_res$departure & !departure_trend_res$trend]
t1e_trend_only <- departure_trend_res$T1E[
  !departure_trend_res$departure & departure_trend_res$trend]
t1e_both_factors <- departure_trend_res$T1E[
  departure_trend_res$departure & departure_trend_res$trend]

additive_pred <- t1e_departure_only + t1e_trend_only - t1e_neither
interaction_term <- t1e_both_factors - additive_pred

cat(sprintf("Departure only:   %.4f  (+%.4f vs. calibration)\n",
            t1e_departure_only,  t1e_departure_only  - t1e_neither))
cat(sprintf("Trend only:       %.4f  (+%.4f vs. calibration)\n",
            t1e_trend_only, t1e_trend_only - t1e_neither))
cat(sprintf("Additive prediction: %.4f\n", additive_pred))
cat(sprintf("Observed (both):     %.4f\n", t1e_both_factors))
cat(sprintf("Interaction term:    %+.4f  (point estimate; MC uncertainty below)\n",
            interaction_term))

departure_trend_res |>
  mutate(label = factor(label, levels = label)) |>
  ggplot(aes(label, T1E,
             fill = factor(departure + trend * 2,
                           labels = c("Neither","Departure","Trend","Both")))) +
  geom_col(width = 0.52) +
  geom_hline(yintercept = alpha, linetype = "dashed",
             color = ZET$red, linewidth = 0.8) +
  geom_hline(yintercept = additive_pred, linetype = "dotted",
             color = ZET$slate, linewidth = 1.0) +
  # Keep the additive-prediction annotation between adjacent gridlines.
  annotate("text", x = 0.55,
           y = (additive_pred + (floor(additive_pred / 0.05) + 1) * 0.05) / 2,
           label = sprintf("Additive prediction: %.4f", additive_pred),
           size = 4.1, fontface = "bold", color = ZET$ink, hjust = 0) +
  # The arc links the additive prediction to the observed value.
  annotate("curve", x = 1.95, y = additive_pred + 0.012,
           xend = 3.82, yend = t1e_both_factors + 0.006,
           curvature = -0.20, angle = 90, ncp = 12,
           arrow = arrow(length = unit(0.18, "cm"), type = "closed"),
           color = ZET$green, linewidth = 0.6) +
  annotate("text", x = 2.85, y = t1e_both_factors + 0.020,
           label = sprintf("Interaction point estimate = %+.4f", interaction_term),
           size = 4.1, fontface = "bold", color = ZET$green, hjust = 0.5) +
  annotate("text", x = 1.5, y = alpha, label = "α = 0.025",
           hjust = 0.5, vjust = -0.45, size = 3.5,
           fontface = "bold", color = ZET$red) +
  geom_text(aes(label = sprintf("%.4f", T1E)),
            vjust = -0.5, size = 3.8, fontface = "bold", color = ZET$ink) +
  # The x-axis already names all four fill categories, so no legend is needed.
  scale_fill_manual(values = c(Neither  = ZET$neutralF,
                                Departure = "#2563EB",
                                Trend    = ZET$teal,
                                Both     = ZET$green),
                    guide = "none") +
  scale_y_continuous(limits = c(0, max(t1e_departure_trend) * 1.20),
                     labels = percent_format(accuracy = 0.1)) +
  labs(x = NULL, y = "Type I error", fill = NULL) +
  theme_zetyra() +
  theme(axis.title = element_text(size = 15, face = "bold", color = ZET$ink,
                                  margin = margin(t = 4, r = 4)),
        axis.text  = element_text(size = 13, color = ZET$slate))
```

```{r departure-trend-table}
tribble(
  ~Condition, ~T1E, ~`vs. calibration`, ~Note,
  "Neither (calibration)",    t1e_neither,   "—",              "threshold target",
  "Departure only (no trend)", t1e_departure_only,
    sprintf("%+.4f", t1e_departure_only - t1e_neither),  "prior away from realised rate",
  "Trend only (at anchor)", t1e_trend_only,
    sprintf("%+.4f", t1e_trend_only - t1e_neither), "rate drifts from anchor",
  "Both (primary)",           t1e_both_factors,
    sprintf("%+.4f", t1e_both_factors - t1e_neither),  sprintf("interaction = %+.4f", interaction_term)
) |>
  mutate(T1E = sprintf("%.4f", T1E)) |>
  kable(caption = paste("2x2 crossing of baseline-rate departure and time trend.",
                        "Both factors act through the monitoring rule, so this",
                        "crosses factors rather than isolating mechanisms."),
        align = c("l","r","r","l"))
```

---

## 5.6 Type I Error Across the Composite Null

The scenarios above each evaluate Type I error at a single control rate. But
$H_0$ here is composite: it holds for *every* common rate $p_e = p_c = p$, not
only at $p = 0.25$. Type I error control is a statement about the supremum
over that set, so a single-point evaluation cannot establish it.

Sweeping $p$ separates two questions that the single-point analysis conflates:
how the calibrated design behaves away from its calibration point, and what
enabling the adaptive components changes at that same threshold.  Note that the
comparison series holds $\gamma^*$ at the value calibrated for the *full*
design and merely disables SSR and RAR; it is a fixed-threshold ablation, not a
separately calibrated monitoring-only design.

Three caveats govern how the sweep should be read.

**The crossing point, if any, is not sharply identified.** Each row is a Monte
Carlo estimate and the grid is discrete. Locating a sign change in the
fixed-$\gamma^*$ contrast would require a denser grid and preferably paired
common-random-number contrasts; the table supports statements only about the
evaluated points.

**The SSR rule shares a threshold with the sweep.** SSR is a pooled
blinded nuisance-parameter rule under a working 1:1 Wald calculation, while the
operational allocation is response-adaptive and the final estimand is the
Bayesian two-arm comparison $p_e > p_c$. Any contrast involving SSR therefore
belongs to this prespecified working rule and should not be generalized to all
blinded SSR procedures.

**Extended trials accrue more drift.** With drift anchored to the planned
horizon $n_{\text{base}} = 90$, Rule B's primary $n=45$ mapping reaches at
most $N=100$. The largest possible terminal drift is therefore
$0.05 \times 99/89 \approx 0.0556$, rather than exactly $0.05$. The SSR
contrasts include this modest additional secular-trend exposure.

```{r composite-null-sweep}
#| cache: true
set.seed(600L)

sweep_rows <- lapply(p_null_grid, function(pn) {
  mon <- t1e_at(replicate(N_sim_sens,
           simulate_one_trial(mix = prior_mix, p_true_e = pn, p_c_data = pn,
                              delta = delta_trend,
                              use_ssr = FALSE, use_rar = FALSE)), gamma_star)
  all <- t1e_at(replicate(N_sim_sens,
           simulate_one_trial(mix = prior_mix, p_true_e = pn, p_c_data = pn,
                              delta = delta_trend,
                              use_ssr = TRUE,  use_rar = TRUE)), gamma_star)
  tibble(p = pn, `SSR + RAR disabled` = mon, `All mechanisms` = all,
         `Fixed-gamma* difference` = all - mon)
})
composite_null <- bind_rows(sweep_rows)

composite_null |>
  mutate(across(c(`SSR + RAR disabled`, `All mechanisms`),
                ~ sprintf("%.4f", .x)),
         `Fixed-gamma* difference` = sprintf("%+.4f", `Fixed-gamma* difference`)) |>
  kable(caption = paste("Type I error across the composite null",
                        "(p_e = p_c = p, trend active). Both columns hold the",
                        "full-design threshold gamma* fixed; 'SSR + RAR",
                        "disabled' is an ablation, not a separately calibrated",
                        "design. Read each row as a pointwise Monte Carlo",
                        "evaluation of the frozen Rule B pipeline."),
        align = c("r","r","r","r"))
```

```{r composite-null-plot, fig.cap="Type I error across the composite null under Rule B. Grey holds the full-design threshold gamma* with SSR and RAR disabled; green is the full design. Grey is a fixed-threshold ablation, not a separately calibrated monitoring-only design, so the gap between the series is not an attributable share."}
composite_null |>
  select(p, `SSR + RAR disabled`, `All mechanisms`) |>
  pivot_longer(-p, names_to = "Design", values_to = "T1E") |>
  ggplot(aes(p, T1E, color = Design, shape = Design)) +
  geom_hline(yintercept = alpha, linetype = "dashed",
             color = ZET$red, linewidth = 0.8) +
  # Put the nominal-alpha reference inside the panel to keep the figure compact.
  annotate("text", x = max(composite_null$p), y = alpha,
           label = "nominal α = 0.025", hjust = 1, vjust = -0.45,
           size = 4.3, fontface = "bold", color = ZET$red) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.8) +
  scale_color_manual(values = c(`SSR + RAR disabled` = ZET$neutralL,
                                `All mechanisms`  = ZET$green)) +
  # Five-point breaks make the low-error region readable.
  scale_y_continuous(breaks = seq(0, 0.25, by = 0.05),
                     labels = percent_format(accuracy = 1)) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = bquote("Common baseline rate " * p), y = "Type I error",
       color = NULL, shape = NULL) +
  theme_zetyra() +
  theme(axis.title = element_text(size = 15, face = "bold", color = ZET$ink,
                                  margin = margin(t = 4, r = 4)),
        axis.text  = element_text(size = 13, color = ZET$slate),
        legend.position             = "inside",
        legend.position.inside      = c(0.01, 0.99),
        legend.justification.inside = c(0, 1),
        legend.direction            = "vertical",
        legend.text  = element_text(size = 12, color = ZET$slate),
        legend.background = element_rect(fill = "white", color = "#D8DCE2"),
        legend.margin     = margin(4, 8, 4, 6))
```

---

## 5.7 Sensitivity Analysis


All three sweeps pass the relevant parameter directly into `simulate_one_trial()`.
No global variables are mutated.

```{r sens-prior-weight}
#| cache: true
# Swept quantity is the weight on the FLAT robustify() component: larger
# w_vague = less borrowing from the historical prior.
w_grid <- seq(0, 0.90, by = 0.10)
set.seed(300L)

t1e_w <- sapply(w_grid, function(w) {
  t1e_at(replicate(N_sim_sens,
                 simulate_one_trial(mix    = set_vague_weight(prior_mix, w),
                                    p_true_e = p_departure_strong,
                                    p_c_data = p_departure_strong,
                                    delta    = delta_trend,
                                    use_ssr  = TRUE,
                                    use_rar  = TRUE)), gamma_star)
})

p_w <- tibble(w = w_grid, t1e = t1e_w) |>
  ggplot(aes(w, t1e)) +
  geom_line(color = ZET$green, linewidth = 1.2) +
  geom_point(color = ZET$green, size = 2.5) +
  geom_hline(yintercept = alpha, linetype = "dashed", color = ZET$red) +
  scale_x_continuous(labels = percent_format()) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(title = "Vague-Component Weight", x = bquote(w[vague]), y = "T1E") +
  theme_zetyra()
```

```{r sens-time-trend}
#| cache: true
delta_grid <- seq(0, 0.15, by = 0.025)
set.seed(400L)

t1e_d <- sapply(delta_grid, function(d) {
  t1e_at(replicate(N_sim_sens,
                 simulate_one_trial(mix    = prior_mix,
                                    p_true_e = p_departure_strong,
                                    p_c_data = p_departure_strong,
                                    delta    = d,          # varied here
                                    use_ssr  = TRUE,
                                    use_rar  = TRUE)), gamma_star)
})

p_d <- tibble(d = delta_grid, t1e = t1e_d) |>
  ggplot(aes(d, t1e)) +
  geom_line(color = ZET$green, linewidth = 1.2) +
  geom_point(color = ZET$green, size = 2.5) +
  geom_hline(yintercept = alpha, linetype = "dashed", color = ZET$red) +
  scale_x_continuous(labels = percent_format()) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(title = "Time Trend Magnitude", x = bquote(δ), y = "T1E") +
  theme_zetyra()
```

```{r sens-ssr-timing}
#| cache: true
ssr_grid <- c(20L, 30L, 40L, 45L, 55L, 60L)
set.seed(500L)

t1e_ssr <- sapply(ssr_grid, function(nt) {
  t1e_at(replicate(N_sim_sens,
                 simulate_one_trial(mix      = prior_mix,
                                    p_true_e   = p_departure_strong,  # H0: p_e = p_c = p_departure_strong
                                    p_c_data   = p_departure_strong,
                                    delta      = delta_trend,
                                    n_ssr_trig = nt,       # varied here
                                    use_ssr    = TRUE,
                                    use_rar    = TRUE)), gamma_star)
})

p_ssr <- tibble(
  nt    = ssr_grid,
  t1e   = t1e_ssr,
  label = sprintf("n=%d\n(%d%%)", ssr_grid,
                  as.integer(round(ssr_grid / N_base * 100)))
) |>
  ggplot(aes(label, t1e)) +
  geom_col(fill = ZET$green, width = 0.55, alpha = 0.90) +
  geom_hline(yintercept = alpha, linetype = "dashed", color = ZET$red) +
  geom_text(aes(label = sprintf("%.4f", t1e)),
            vjust = -0.5, size = 3.5, color = ZET$ink) +
  scale_y_continuous(limits = c(0, max(t1e_ssr) * 1.25),
                     labels = percent_format(accuracy = 0.1)) +
  labs(title = "SSR Trigger Timing", x = bquote(n[SSR]), y = "T1E") +
  theme_zetyra()
```

```{r sens-panel, fig.width=13, fig.height=4.5, fig.cap="Sensitivity of Type I error to prior weight (left), time trend (centre), and Rule B timing (right). Dashed line = nominal alpha = 0.025."}
p_w | p_d | p_ssr
```

```{r sens-table}
tribble(
  ~Parameter,        ~Range,              ~`Min T1E`,             ~`Max T1E`,
  "Vague-component weight (w<sub>vague</sub>)", "0% – 90%",
    sprintf("%.4f", min(t1e_w)),   sprintf("%.4f", max(t1e_w)),
  "Time trend (δ)",  "0% – 15%",
    sprintf("%.4f", min(t1e_d)),   sprintf("%.4f", max(t1e_d)),
  "SSR timing (n<sub>SSR</sub>)",      "n = 20, 30, 40, 45, 55, 60",
    sprintf("%.4f", min(t1e_ssr)), sprintf("%.4f", max(t1e_ssr))
) |>
  kable(caption = "Sensitivity bounds across all parameter ranges",
        align   = c("l","l","r","r"),
        escape  = FALSE)
```

---

## 5.8 Reproducibility

```{r auto-summary}
#| code-fold: false
cat(sprintf("gamma*                = %.3f\n", gamma_star))
cat(sprintf("T1E (calibration)     = %.4f  (target: %.3f)\n", t1e_at_star, alpha))
cat(sprintf("T1E (fresh anchor)    = %.4f  (95%% SI [%.4f, %.4f])\n",
            anchor_validation_t1e,
            anchor_validation_ci["lower"], anchor_validation_ci["upper"]))
cat(sprintf("T1E (mild departure, p=0.27) = %.4f  (+%.1f%%)\n",
            t1e_mild, inflation_pct_mild))
cat(sprintf("T1E (strong departure, p=0.30) = %.4f  (+%.1f%%)\n",
            t1e_primary, inflation_pct))
cat(sprintf("T1E (earlier SSR, n_ssr=30)   = %.4f  (+%.1f%%)\n",
            t1e_neg, (t1e_neg - alpha)/alpha*100))
cat(sprintf("Earlier-vs-primary change = %+.4f (%+.1f%% of primary excess)\n",
            timing_change, timing_change_pct_excess))
cat(sprintf("\nFixed-threshold factorial ablation:\n"))
cat(sprintf("  Delta_SSR         = %+.4f\n", delta_f_ssr))
cat(sprintf("  Delta_RAR         = %+.4f\n", delta_f_rar))
cat(sprintf("  Delta_interaction = %+.4f\n", delta_f_int))
cat(sprintf("\nBaseline-rate departure x trend:\n"))
cat(sprintf("  Departure only    = +%.4f\n", t1e_departure_only  - t1e_neither))
cat(sprintf("  Trend only        = +%.4f\n", t1e_trend_only - t1e_neither))
cat(sprintf("  Both (primary)    = +%.4f\n", t1e_both_factors  - t1e_neither))
cat(sprintf("  Interaction term  = %+.4f\n", interaction_term))
```

```{r mcse-table}
mcse_tbl <- tribble(
  ~Estimate,                     ~T1E,           ~N_sim,
  "Calibration (at anchor)",     t1e_at_star,    N_sim_calib,
  "Independent anchor validation", anchor_validation_t1e, N_sim_sens,
  "Mild departure (p=0.27)",     t1e_mild,       N_sim_main,
  "Strong departure (p=0.30)",   t1e_primary,    N_sim_main,
  "Earlier SSR (n = 30)",        t1e_neg,        N_sim_main,
  "Factorial: baseline",         t1e_f_base,     N_sim_sens,
  "Factorial: SSR only",         t1e_f_ssr,      N_sim_sens,
  "Factorial: RAR only",         t1e_f_rar,      N_sim_sens,
  "Factorial: SSR + RAR",        t1e_f_both,     N_sim_sens,
  "Departure × trend: neither",  t1e_neither,    N_sim_sens,
  "Departure x trend: departure only", t1e_departure_only, N_sim_sens,
  "Departure × trend: trend only", t1e_trend_only, N_sim_sens,
  "Departure × trend: both",     t1e_both_factors, N_sim_sens
) |>
  mutate(
    MCSE   = mcse(T1E, N_sim),
    CI_lo  = T1E - 1.96 * MCSE,
    CI_hi  = T1E + 1.96 * MCSE,
    T1E    = sprintf("%.4f", T1E),
    MCSE   = sprintf("%.4f", MCSE),
    `95% SI` = sprintf("[%.4f, %.4f]", CI_lo, CI_hi),
    N_sim  = formatC(N_sim, format = "d", big.mark = ",")
  ) |>
  select(Estimate, T1E, MCSE, `95% SI`, N_sim)

kable(mcse_tbl,
      caption = "Monte Carlo standard errors and 95% simulation intervals for all key T1E estimates.",
      align = c("l", "r", "r", "c", "r"))

# ── Propagated MCSEs for derived interaction terms ──────────────────────────
# Interaction = T1E_both - T1E_A - T1E_B + T1E_neither
# All four simulations are independent, so Var(interaction) = sum of variances

# Departure x trend interaction
se_departure_trend <- sqrt(
  mcse(t1e_both_factors,  N_sim_sens)^2 +
  mcse(t1e_departure_only,  N_sim_sens)^2 +
  mcse(t1e_trend_only, N_sim_sens)^2 +
  mcse(t1e_neither,    N_sim_sens)^2
)

# Factorial interaction
se_fact <- sqrt(
  mcse(t1e_f_both, N_sim_sens)^2 +
  mcse(t1e_f_ssr,  N_sim_sens)^2 +
  mcse(t1e_f_rar,  N_sim_sens)^2 +
  mcse(t1e_f_base, N_sim_sens)^2
)

int_tbl <- tribble(
  ~`Interaction term`,                    ~Estimate,        ~MCSE,    ~`95% SI`,
  "Departure × trend", interaction_term, se_departure_trend,
    sprintf("[%+.4f, %+.4f]", interaction_term - 1.96*se_departure_trend, interaction_term + 1.96*se_departure_trend),
  "Factorial (SSR × RAR)",                  delta_f_int,     se_fact,
    sprintf("[%+.4f, %+.4f]", delta_f_int - 1.96*se_fact, delta_f_int + 1.96*se_fact)
) |>
  mutate(Estimate = sprintf("%+.4f", Estimate),
         MCSE    = sprintf("%.4f", MCSE))

kable(int_tbl,
      caption = paste("Propagated MCSEs for derived interaction terms.",
                      "Intervals are Monte Carlo simulation intervals, not",
                      "sampling-confidence intervals for clinical data."),
      align = c("l", "r", "r", "c"))
```

**Seeds:**

| Simulation | Seed |
|---|---|
| MAP prior MCMC | 42 (via `set.seed()` before `gMAP()`) |
| RAR validation | 42 |
| Threshold calibration | 1 |
| Independent anchor validation | 101 |
| Primary: strong departure (p=0.30) | 2026 |
| Primary: mild departure (p=0.27) | 2028 |
| Earlier-SSR timing sensitivity (n<sub>SSR</sub> = 30) | 2027 |
| Fixed-threshold factorial ablation | 3000 |
| Departure × trend (2x2) | 4000 |
| Sensitivity — prior weight | 300 |
| Sensitivity — time trend | 400 |
| Sensitivity — SSR timing | 500 |

```{r session-info}
#| code-fold: false
sessionInfo()
```

---

## References

::: {#refs}
:::
